---
title: "03.CG.fecundity.accession.models"
author: "Daniele Filiault"
date: "2/26/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Description and Motivation
Taking a closer look at fecundity in the common garden experiments.
Taking same approach as for fecundity
Look for local adaptation and try to assess which components of fecundity
vary across sites and years to cause fecundity differences.

Fit models to get fecundity for each site
First make data variable

## 1. Local adaptation with fecundity values - combined years
### 1a. prep data
```{r}
d11=readRDS(file="./data/d11_for_analysis.rds")
d12=readRDS(file="./data/d12_for_analysis.rds")

##restrict to fecundity
traits=c("fecundity")

d11$year=2011
d12$year=2012

d11=d11[, c("year", "exp", "block", "id", "fecundity")]
d12=d12[, c("year", "exp", "block", "id", "fecundity")]

d=rbind(d11, d12)
d$region[d$exp%in%c("ULL", "RAT")]="S"
d$region[d$exp%in%c("RAM", "ADA")]="N"
d$year=as.factor(d$year)
d=d[, c("year", "region", "exp","block", "id", "fecundity")]
d=na.omit(d)

hist(d$fecundity)
with(d, boxplot(fecundity~exp+year, ylab="fecundity",las=2))
```
So the distribution of the fecundity variable is highly skewed towards zero.  This is not a a case of zero inflation, but rather that there
simply are a lot of zero fecundity plants here.  Therefore, need a more complicated model.  Also should include block as a random effect.  Therefore...

So distribution of fecundity has zeros, so can't do gamma link.  But only 50 zero observations.


Fit glms with Tweedie distributions.
```{r tweedie_prep}
library(lme4)
library(statmod)
library(tweedie)
```

###1b. start with Ullstorp

```{r}
du <- d[d$exp=="ULL",]
hist(du$fecundity)
```

determine d for tweedie distribution
```{r eval=FALSE, include=TRUE}
### also need to test block in model, but to assess p, use easiest complete model - can't do random effects so just nest block within year

profile.u <- tweedie.profile(fecundity ~ id + year + id*year + year/block, data = du, p.vec = seq(1.1, 1.9, 0.1),do.plot=TRUE, fit.glm = TRUE)
print(profile.u$p.max)

```
Now fit models, assess model components

```{r ull_model}

#vp.up <- profile.u$p.max
vp.up <- 1.720408  ### doing this so i can knit the doc in a reasonable amount of time.  The Tweedie p assessment is slow...
dat.up <- du

glm.u1 <- glm(fecundity ~ id + year + id*year + year/block, family=tweedie(var.power=vp.up, link.power=0), data=dat.up)

glm.u2 <- glm(fecundity ~ id + year + year/block, family=tweedie(var.power=vp.up, link.power=0), data=dat.up)
anova(glm.u2, glm.u1, test="Chisq")

glm.u3 <- glm(fecundity ~ id + year/block, family=tweedie(var.power=vp.up, link.power=0), data=dat.up)
anova(glm.u3, glm.u1, test="Chisq")

glm.u4 <- glm(fecundity ~ id + year + id*year, family=tweedie(var.power=vp.up, link.power=0), data=dat.up)
anova(glm.u4, glm.u1, test="Chisq")

glm.u5 <- glm(fecundity ~ year + year/block, family=tweedie(var.power=vp.up, link.power=0), data=dat.up)
anova(glm.u5, glm.u1, test="Chisq")

#plot(glm.u1)

```
All terms in full model are significant, so use this one to get fitted fecundity values per line

```{r}
du.predict <- predict(glm.u1, type="response")
hist(du.predict, main="predicted values Ullstorp")

du.dat <- cbind(du,du.predict)
du.means <- aggregate(du.predict~id, data=du.dat, mean)
hist(du.means$du.predict, main="predicted means Ullstorp", xlab="fecundity")

```
continue

check local adaptation

```{r}
clim=read.table("./data/worldclim_swedish_acc.txt", sep="\t", h=T, stringsAsFactor=T)
clim_exp=read.table("./data/worldclim_swedish_exp.txt", sep="\t", h=T)
colnames(clim)[1]="id"

du.means <- merge(du.means, clim[,1:6], by="id")
lm.ufit <- lm(du.predict~long, data=du.means)
summary(lm.ufit)
plot(du.means$lat, du.means$du.predict, xlab="latitude of accession origin", ylab="fecundity in Ullstorp")
abline(lm(du.predict~lat, data=du.means), col="blue")
```
Yes, signs of local adaptation in Ullstorp

###1c. Adal East

```{r}
da <- d[d$exp=="ADA",]
hist(da$fecundity, main="Adal fecundity", xlab="fecundity")
```

determine d for tweedie distribution

```{r eval=FALSE, include=TRUE}
profile.a <- tweedie.profile(fecundity ~ id + year + id*year + year/block, data = da, p.vec = seq(1.1, 1.9, 0.1), fit.glm = TRUE, control=list(maxit=50), do.plot=TRUE)
print(profile.a$p.max)
```

Now fit models, assess model components

```{r adal_east_model}
#vp.up <-profile.a$p.max
vp.up <- 1.785714
dat.up <- da

glm.a1 <- glm(fecundity ~ id + year + id*year + year/block, family=tweedie(var.power=vp.up, link.power=0), data=dat.up)

glm.a2 <- glm(fecundity ~ id + year + year/block, family=tweedie(var.power=vp.up, link.power=0), data=dat.up)
anova(glm.a2, glm.a1, test="Chisq")

glm.a3 <- glm(fecundity ~ id + year/block, family=tweedie(var.power=vp.up, link.power=0), data=dat.up)
anova(glm.a3, glm.a1, test="Chisq")

glm.a4 <- glm(fecundity ~ id + year + id*year, family=tweedie(var.power=vp.up, link.power=0), data=dat.up)
anova(glm.a4, glm.a1, test="Chisq")

glm.a5 <- glm(fecundity ~ year + year/block, family=tweedie(var.power=vp.up, link.power=0), data=dat.up)
anova(glm.a5, glm.a1, test="Chisq")

#plot(glm.a1)
### wow, this is ugly!  the point messing everything up is row 1938

```

All terms in full model are significant, so use this one to get fitted fecundity values per line

```{r}
da.predict <- predict(glm.a1, type="response")
hist(da.predict)

da.dat <- cbind(da,da.predict)
da.means <- aggregate(da.predict~id, data=da.dat, mean)
hist(da.means$da.predict, main="hist of predicted means")
```

check local adaptation

```{r}
da.means <- merge(da.means, clim[,1:6], by="id")
lm.afit <- lm(da.predict~lat, data=da.means)
summary(lm.afit)
plot(da.means$lat, da.means$da.predict, xlab="latitude of accession origin", ylab="fecundity in Ådal East")
abline(lm(da.predict~lat, data=da.means), col="blue")
```

marginally sig local adaptation Adal East



###1d. Rathckegarden

```{r}
dr <- d[d$exp=="RAT",]
hist(dr$fecundity, main="histogram of Rat fecundity")
```

determine d for tweedie distribution

```{r eval=FALSE, include=TRUE}

profile.r <- tweedie.profile(fecundity ~ id + year + id*year + year/block, data = dr, p.vec = seq(1.1, 1.9, 0.1), fit.glm = TRUE, control=list(maxit=50), do.plot=TRUE)
print(profile.r$p.max)
### use this var.power for the tweedie distribution
```

Now fit models, assess model components

```{r rathckegarden_model}
#vp.up <-profile.r$p.max
vp.up <-  1.742857
dat.up <- dr

glm.r1 <- glm(fecundity ~ id + year + id*year + year/block, family=tweedie(var.power=vp.up, link.power=0), data=dat.up)

glm.r2 <- glm(fecundity ~ id + year + year/block, family=tweedie(var.power=vp.up, link.power=0), data=dat.up)
anova(glm.r2, glm.r1, test="Chisq")

glm.r3 <- glm(fecundity ~ id + year/block, family=tweedie(var.power=vp.up, link.power=0), data=dat.up)
anova(glm.r3, glm.r1, test="Chisq")

glm.r4 <- glm(fecundity ~ id + year + id*year, family=tweedie(var.power=vp.up, link.power=0), data=dat.up)
anova(glm.r4, glm.r1, test="Chisq")

glm.r5 <- glm(fecundity ~ year + year/block, family=tweedie(var.power=vp.up, link.power=0), data=dat.up)
anova(glm.r5, glm.r1, test="Chisq")

#plot(glm.r1)

```

All terms in full model are significant, so use this one to get fitted fecundity values per line

```{r}
dr.predict <- predict(glm.r1, type="response")
hist(dr.predict, main="hist of fitted values Rat")

dr.dat <- cbind(dr,dr.predict)
dr.means <- aggregate(dr.predict~id, data=dr.dat, mean)
hist(dr.means$dr.predict, main="hist of mean fitted values Rat")
```



check local adaptation

```{r}
dr.means <- merge(dr.means, clim[,1:6], by="id")
lm.rfit <- lm(dr.predict~lat, data=dr.means)
summary(lm.rfit)
plot(dr.means$lat, dr.means$dr.predict, xlab="latitude of accession origin", ylab="fecundity in Ratheckgården")
abline(lm(dr.predict~lat, data=dr.means), col="blue")
```
local adaptation Rathckegarden!

###1e. Ramsta

```{r}
dm <- d[d$exp=="RAM",]
hist(dm$fecundity, main="histogram of Ramsta fecundity")
```

```{r eval=FALSE, include=TRUE}

profile.m <- tweedie.profile(fecundity ~ id + year + id*year + year/block, data = dm, p.vec = seq(1.1, 1.9, 0.1), fit.glm = TRUE, control=list(maxit=50), do.plot=TRUE)
print(profile.m$p.max)
### use this var.power for the tweedie distribution
```

Now fit models, assess model components

```{r ramsta_model}
#vp.up <-profile.m$p.max
vp.up <- 1.571429
dat.up <- dm

glm.m1 <- glm(fecundity ~ id + year + id*year + year/block, family=tweedie(var.power=vp.up, link.power=0), data=dat.up, maxit=50)

glm.m2 <- glm(fecundity ~ id + year + year/block, family=tweedie(var.power=vp.up, link.power=0), data=dat.up)
anova(glm.m2, glm.m1, test="Chisq")

glm.m3 <- glm(fecundity ~ id + year/block, family=tweedie(var.power=vp.up, link.power=0), data=dat.up)
anova(glm.m3, glm.m1, test="Chisq")

glm.m4 <- glm(fecundity ~ id + year + id*year, family=tweedie(var.power=vp.up, link.power=0), data=dat.up, maxit=50)
anova(glm.m4, glm.m1, test="Chisq")

glm.m5 <- glm(fecundity ~ year + year/block, family=tweedie(var.power=vp.up, link.power=0), data=dat.up)
anova(glm.m5, glm.m1, test="Chisq")

#plot(glm.m1)

```


```{r}
dm.predict <- predict(glm.m1, type="response")
hist(dm.predict, main="hist of fitted values Ramsta")

dm.dat <- cbind(dm,dm.predict)
dm.means <- aggregate(dm.predict~id, data=dm.dat, mean)
hist(dm.means$dm.predict, main="hist of mean fitted values Ramsta")
```

check local adaptation

```{r}
dm.means <- merge(dm.means, clim[,1:6], by="id")
lm.mfit <- lm(dm.predict~lat, data=dm.means)
summary(lm.mfit)
plot(dm.means$lat, dm.means$dm.predict, xlab="latitude of accession origin", ylab="fecundity in Ramsta")
abline(lm(dm.predict~lat, data=dm.means), col="blue")
```
No local adaptation in Ramsta

###1f. Combining N/S comparisons

```{r}
all.means <- merge(da.means, dm.means[,1:2], by="id")
all.means <- merge(all.means, du.means[,1:2], by="id")
all.means <- merge(all.means, dr.means[,1:2], by="id")
all.means <- all.means[,c(1,3,4,5,6,7,2,8,9,10)]
pairs(all.means[,7:10])

all.means$n.means <- apply(all.means,1,function(x){(as.numeric(x[7])+as.numeric(x[8]))/2})
all.means$s.means <- apply(all.means,1,function(x){(as.numeric(x[9])+as.numeric(x[10]))/2})

lm.ns.mean <- lm(n.means~s.means, data=all.means)
lm.ns.mean.loc <- lm(n.means~s.means + region , data=all.means)
lm.ns.mean.int <- lm(n.means~s.means + region + s.means*region , data=all.means)

summary(lm.ns.mean)
summary(lm.ns.mean.loc)
summary(lm.ns.mean.int)

aov.ns.mean.loc <- aov(n.means~s.means + region, data=all.means)
ns.posthoc <- TukeyHSD(aov.ns.mean.loc, "region")
plot(ns.posthoc)

with(all.means, plot(s.means, n.means, xlab="mean fecundity in South", ylab="mean fecundity in North"))
#abline(lm.ns.mean)
abline(0,1, col="blue")

library(ggplot2)

ns.plot <- ggplot(all.means, aes(s.means, n.means)) + geom_point(aes(colour=region))
ns.plot + theme_bw() +   geom_smooth( aes(linetype = region, color=region), method = "lm") + labs(x="mean fecundity in Southern experiments", y="mean fecundity in Northern experiments")

lm.nfit <- lm(n.means~lat, data=all.means)
summary(lm.nfit)
plot(all.means$lat, all.means$n.means, xlab="latitude of accession origin", ylab=" mean fecundity in Northern sites")
abline(lm(n.means~lat, data=all.means), col="blue")

lm.sfit <- lm(s.means~lat, data=all.means)
summary(lm.sfit)
plot(all.means$lat, all.means$s.means, xlab="latitude of accession origin", ylab=" mean fecundity in Southern sites")
abline(lm(s.means~lat, data=all.means), col="blue")

```
1.  Joint years, local adaptation in S, not in N.
2.  generally higher fecundity in N

## 3. Local adaptation with fecundity values - 2011
###3a. start with Ullstorp

```{r}
du1 <- d[d$exp=="ULL" & d$year=="2011" ,]
hist(du1$fecundity)
```
determine d for tweedie distribution

```{r eval=FALSE, include=TRUE}

profile.u1 <- tweedie.profile(fecundity ~ id + block + block*id, data = du1, p.vec = seq(1.1, 1.9, 0.1), fit.glm = TRUE, control=list(maxit=50), do.plot=TRUE)
print(profile.u1$p.max)
### use this var.power for the tweedie distribution
```

Now fit models, assess model components

```{r ullstorp2011_model}
#vp.up <-profile.u1$p.max
vp.up <-1.7
dat.up <- du1

glm.u11 <- glm(fecundity ~ id + block + block*id, family=tweedie(var.power=vp.up, link.power=0), data=dat.up)

glm.u12 <- glm(fecundity ~ id + block, family=tweedie(var.power=vp.up, link.power=0), data=dat.up)
anova(glm.u12, glm.u11, test="Chisq")

glm.u13 <- glm(fecundity ~ id , family=tweedie(var.power=vp.up, link.power=0), data=dat.up)
anova(glm.u13, glm.u11, test="Chisq")


#plot(glm.u11)

```
All terms significant

 
```{r}
du1.predict <- predict(glm.u11, type="response")
hist(du1.predict, main="hist of fitted values Ullstorp 2011")

du1.dat <- cbind(du1,du1.predict)
du1.means <- aggregate(du1.predict~id, data=du1.dat, mean)
hist(du1.means$du1.predict, main="hist of mean fitted values Ullstorp 2011")
```

```{r}
du1.means <- merge(du1.means, clim[,1:6], by="id")
lm.u1fit <- lm(du1.predict~lat, data=du1.means)
summary(lm.u1fit)
plot(du1.means$lat, du1.means$du1.predict, xlab="latitude of accession origin", ylab="fecundity in Ullstorp 2011")
abline(lm(du1.predict~lat, data=du1.means), col="blue")
```

Yes, local adaptation Ull 2011, less effect than combined years

###3b. Adal East 2011

```{r}
da1 <- d[d$exp=="ADA" & d$year=="2011" ,]
hist(da1$fecundity)
```
determine d for tweedie distribution

```{r eval=FALSE, include=TRUE}

profile.a1 <- tweedie.profile(fecundity ~ id + block + block*id, data = da1, p.vec = seq(1.1, 1.9, 0.1), fit.glm = TRUE, control=list(maxit=50), do.plot=TRUE)
print(profile.a1$p.max)
### use this var.power for the tweedie distribution
```


```{r adaleast2011_model}
#vp.up <-profile.a1$p.max
vp.up <- 1.818367
dat.up <- da1

glm.a11 <- glm(fecundity ~ id + block + block*id, family=tweedie(var.power=vp.up, link.power=0), data=dat.up)

glm.a12 <- glm(fecundity ~ id + block, family=tweedie(var.power=vp.up, link.power=0), data=dat.up)
anova(glm.a12, glm.a11, test="Chisq")

glm.a13 <- glm(fecundity ~ id , family=tweedie(var.power=vp.up, link.power=0), data=dat.up)
anova(glm.a13, glm.a11, test="Chisq")


#plot(glm.a11)

```
All terms significant

 
```{r}
da1.predict <- predict(glm.a11, type="response")
hist(da1.predict, main="hist of fitted values Ådal East 2011")

da1.dat <- cbind(da1,da1.predict)
da1.means <- aggregate(da1.predict~id, data=da1.dat, mean)
hist(da1.means$da1.predict, main="hist of mean fitted values Ådal East 2011")
```

```{r}
da1.means <- merge(da1.means, clim[,1:6], by="id")
lm.a1fit <- lm(da1.predict~lat, data=da1.means)
summary(lm.a1fit)
plot(da1.means$lat, da1.means$da1.predict, xlab="latitude of accession origin", ylab="fecundity in Ådal East 2011")
abline(lm(da1.predict~lat, data=da1.means), col="blue")
```
Nothing Adal East 2011

###3c. Rat 2011

```{r}
dr1 <- d[d$exp=="RAT" & d$year=="2011" ,]
hist(dr1$fecundity)
```
determine d for tweedie distribution

```{r eval=FALSE, include=TRUE}

profile.r1 <- tweedie.profile(fecundity ~ id + block + block*id, data = dr1, p.vec = seq(1.1, 1.9, 0.1), fit.glm = TRUE, control=list(maxit=75), do.plot=TRUE)
print(profile.r1$p.max)
### use this var.power for the tweedie distribution
```

```{r rat2011_model}
#vp.up <-profile.r1$p.max
vp.up <- 1.491837
dat.up <- dr1

glm.r11 <- glm(fecundity ~ id + block + block*id, family=tweedie(var.power=vp.up, link.power=0), data=dat.up,maxit=100)

glm.r12 <- glm(fecundity ~ id + block, family=tweedie(var.power=vp.up, link.power=0), data=dat.up)
anova(glm.r12, glm.r11, test="Chisq")

glm.r13 <- glm(fecundity ~ id , family=tweedie(var.power=vp.up, link.power=0), data=dat.up)
anova(glm.r13, glm.r11, test="Chisq")


#plot(glm.r11)

```

all terms significant

```{r}
dr1.predict <- predict(glm.r11, type="response")
hist(dr1.predict, main="hist of fitted values Rathckegården 2011")

dr1.dat <- cbind(dr1,dr1.predict)
dr1.means <- aggregate(dr1.predict~id, data=dr1.dat, mean)
hist(dr1.means$dr1.predict, main="hist of mean fitted values Rathckegården 2011")
```



```{r}
dr1.means <- merge(dr1.means, clim[,1:6], by="id")
lm.r1fit <- lm(dr1.predict~lat, data=dr1.means)
summary(lm.r1fit)
plot(dr1.means$lat, dr1.means$dr1.predict, xlab="latitude of accession origin", ylab="fecundity in Rathckegården 2011")
abline(lm(dr1.predict~lat, data=dr1.means), col="blue")
```
Rat 2011  significant

###3d. Ramsta 2011
```{r}
dm1 <- d[d$exp=="RAM" & d$year=="2011" ,]
hist(dm1$fecundity)
```

```{r eval=FALSE, include=TRUE}

profile.m1 <- tweedie.profile(fecundity ~ id + block + block*id, data = dm1, p.vec = seq(1.1, 1.9, 0.1), fit.glm = TRUE, control=list(maxit=50), do.plot=TRUE)
print(profile.m1$p.max)
### use this var.power for the tweedie distribution
```

```{r ramsta2011_model}
#vp.up <-profile.m1$p.max
vp.up <- 1.638776  ### yes, exactly the same as the p.max for rat.  I double checked, and this seems correct.
dat.up <- dm1

glm.m11 <- glm(fecundity ~ id + block + block*id, family=tweedie(var.power=vp.up, link.power=0), data=dat.up,maxit=75)

glm.m12 <- glm(fecundity ~ id + block, family=tweedie(var.power=vp.up, link.power=0), data=dat.up, maxit=75)
anova(glm.m12, glm.m11, test="Chisq")

glm.m13 <- glm(fecundity ~ id , family=tweedie(var.power=vp.up, link.power=0), data=dat.up, maxit=75)
anova(glm.m13, glm.m11, test="Chisq")


#plot(glm.m11)

```

```{r}
dm1.predict <- predict(glm.m11, type="response")
hist(dm1.predict, main="hist of fitted values Ramsta 2011")

dm1.dat <- cbind(dm1,dm1.predict)
dm1.means <- aggregate(dm1.predict~id, data=dm1.dat, mean)
hist(dm1.means$dm1.predict, main="hist of mean fitted values Ramsta 2011")
```



```{r}
dm1.means <- merge(dm1.means, clim[,1:6], by="id")
lm.m1fit <- lm(dm1.predict~lat, data=dm1.means)
summary(lm.m1fit)
plot(dm1.means$lat, dm1.means$dm1.predict, xlab="latitude of accession origin", ylab="fecundity in Ramsta 2011")
abline(lm(dm1.predict~lat, data=dm1.means), col="blue")
```
Ramsta 2011 very significant!

###3e. N/S combined 2011

```{r}
all.means1 <- merge(da1.means, dm1.means[,1:2], by="id")
all.means1 <- merge(all.means1, du1.means[,1:2], by="id")
all.means1 <- merge(all.means1, dr1.means[,1:2], by="id")
all.means1 <- all.means1[,c(1,3,4,5,6,7,2,8,9,10)]
pairs(all.means1[,7:10])

all.means1$n.means <- apply(all.means1,1,function(x){(as.numeric(x[7])+as.numeric(x[8]))/2})
all.means1$s.means <- apply(all.means1,1,function(x){(as.numeric(x[9])+as.numeric(x[10]))/2})

lm.ns.mean1 <- lm(n.means~s.means, data=all.means1)
lm.ns.mean.loc1 <- lm(n.means~s.means + region , data=all.means1)
lm.ns.mean.int1 <- lm(n.means~s.means + region + s.means*region , data=all.means1)

anova(lm.ns.mean1)
anova(lm.ns.mean.loc1)
anova(lm.ns.mean.int1)

aov.ns.mean.loc1 <- aov(n.means~s.means + region, data=all.means1)
ns.posthoc1 <- TukeyHSD(aov.ns.mean.loc1, "region")
plot(ns.posthoc1)

with(all.means1, plot(s.means, n.means, xlab="mean fecundity in South", ylab="mean fecundity in North"))
#abline(lm.ns.mean)
abline(0,1, col="blue")

library(ggplot2)

ns.plot2011 <- ggplot(all.means1, aes(s.means, n.means)) + geom_point(aes(colour=region))
ns.plot2011 + theme_bw() +   geom_smooth( aes(linetype = region, color=region), method = "lm") + labs(x="mean fecundity in Southern experiments 2011", y="mean fecundity in Northern experiments 2011")
```

Do "classic" local adaptation on these means

```{r}

lm.n1 <- lm(n.means~lat, data=all.means1)
summary(lm.n1)
plot(all.means1$lat, all.means1$n.means, xlab="latitude of accession origin", ylab="mean fecundity in North 2011")
abline(lm(n.means~lat, data=all.means1), col="blue")

lm.s1 <- lm(s.means~lat, data=all.means1)
summary(lm.s1)
plot(all.means1$lat, all.means1$s.means, xlab="latitude of accession origin", ylab="mean fecundity in South 2011")
abline(lm(s.means~lat, data=all.means1), col="blue")

```
South is  significant, North is not


## 4. Local adaptation with fecundity values - 2012
###4a. start with Ullstorp

```{r}
du2 <- d[d$exp=="ULL" & d$year=="2012" ,]
hist(du2$fecundity)
```
determine d for tweedie distribution

```{r eval=FALSE, include=TRUE}

profile.u2 <- tweedie.profile(fecundity ~ id + block + block*id, data = du2, p.vec = seq(1.1, 1.9, 0.1), fit.glm = TRUE, control=list(maxit=50), do.plot=TRUE)
print(profile.u2$p.max)
### use this var.power for the tweedie distribution
```

Now fit models, assess model components

```{r ullstorp2012_model}
#vp.up <-profile.u2$p.max
vp.up <- 1.622449
dat.up <- du2

glm.u21 <- glm(fecundity ~ id + block + block*id, family=tweedie(var.power=vp.up, link.power=0), data=dat.up)

glm.u22 <- glm(fecundity ~ id + block, family=tweedie(var.power=vp.up, link.power=0), data=dat.up)
anova(glm.u22, glm.u21, test="Chisq")

glm.u23 <- glm(fecundity ~ id , family=tweedie(var.power=vp.up, link.power=0), data=dat.up)
anova(glm.u23, glm.u21, test="Chisq")


#plot(glm.u21)

```
all significant.  Use glm.u21


 
```{r}
du2.predict <- predict(glm.u21, type="response")
hist(du2.predict, main="hist of fitted values Ullstorp 2012")

du2.dat <- cbind(du2,du2.predict)
du2.means <- aggregate(du2.predict~id, data=du2.dat, mean)
hist(du2.means$du2.predict, main="hist of mean fitted values Ullstorp 2012")
```

```{r}
du2.means <- merge(du2.means, clim[,1:6], by="id")
lm.u2fit <- lm(du2.predict~lat, data=du2.means)
summary(lm.u2fit)
plot(du2.means$lat, du2.means$du2.predict, xlab="latitude of accession origin", ylab="fecundity in Ullstorp 2012")
abline(lm(du2.predict~lat, data=du2.means), col="blue")
```

Yes, strong local adaptation Ull 2012

###4b. Adal East
```{r}
da2 <- d[d$exp=="ADA" & d$year=="2012" ,]
hist(da2$fecundity)
```
determine d for tweedie distribution

```{r eval=FALSE, include=TRUE}

profile.a2 <- tweedie.profile(fecundity ~ id + block + block*id, data = da2, p.vec = seq(1.1, 1.9, 0.1), fit.glm = TRUE, control=list(maxit=50), do.plot=TRUE)
print(profile.a2$p.max)
### use this var.power for the tweedie distribution
```


```{r adaleast2012_model}
#vp.up <-profile.a2$p.max
vp.up <-  1.459184
dat.up <- da2

glm.a21 <- glm(fecundity ~ id + block + block*id, family=tweedie(var.power=vp.up, link.power=0), data=dat.up)

glm.a22 <- glm(fecundity ~ id + block, family=tweedie(var.power=vp.up, link.power=0), data=dat.up)
anova(glm.a22, glm.a21, test="Chisq")

glm.a23 <- glm(fecundity ~ id , family=tweedie(var.power=vp.up, link.power=0), data=dat.up)
anova(glm.a23, glm.a21, test="Chisq")


#plot(glm.a22)

```

Use model 1
 
```{r}
da2.predict <- predict(glm.a21, type="response")
hist(da2.predict, main="hist of fitted values Ådal East 2012")

da2.dat <- cbind(da2,da2.predict)
da2.means <- aggregate(da2.predict~id, data=da2.dat, mean)
hist(da2.means$da2.predict, main="hist of mean fitted values Ådal East 2012")
```

```{r}
da2.means <- merge(da2.means, clim[,1:6], by="id")
lm.a2fit <- lm(da2.predict~lat, data=da2.means)
summary(lm.a2fit)
plot(da2.means$lat, da2.means$da2.predict, xlab="latitude of accession origin", ylab="fecundity in Ådal East 2012")
abline(lm(da2.predict~lat, data=da2.means), col="blue")
```
Adal East 2012  significant 


### 4c. Rathckegarden


```{r}
dr2 <- d[d$exp=="RAT" & d$year=="2012" ,]
hist(dr2$fecundity)
```
determine d for tweedie distribution

```{r eval=FALSE, include=TRUE}

profile.r2 <- tweedie.profile(fecundity ~ id + block + block*id, data = dr2, p.vec = seq(1.1, 1.9, 0.1), fit.glm = TRUE, control=list(maxit=75), do.plot=TRUE)
print(profile.r2$p.max)
### use this var.power for the tweedie distribution
```


```{r rat2012_model}
#vp.up <-profile.r2$p.max
vp.up <- 1.638776
dat.up <- dr2

glm.r21 <- glm(fecundity ~ id + block + block*id, family=tweedie(var.power=vp.up, link.power=0), data=dat.up,maxit=100)

glm.r22 <- glm(fecundity ~ id + block, family=tweedie(var.power=vp.up, link.power=0), data=dat.up)
anova(glm.r22, glm.r21, test="Chisq")

glm.r23 <- glm(fecundity ~ id , family=tweedie(var.power=vp.up, link.power=0), data=dat.up)
anova(glm.r23, glm.r21, test="Chisq")


#plot(glm.r21)

```

all terms significant

```{r}
dr2.predict <- predict(glm.r21, type="response")
hist(dr2.predict, main="hist of fitted values Rathckegården 2012")

dr2.dat <- cbind(dr2,dr2.predict)
dr2.means <- aggregate(dr2.predict~id, data=dr2.dat, mean)
hist(dr2.means$dr2.predict, main="hist of mean fitted values Rathckegården 2012")
```



```{r}
dr2.means <- merge(dr2.means, clim[,1:6], by="id")
lm.r2fit <- lm(dr2.predict~lat, data=dr2.means)
summary(lm.r2fit)
plot(dr2.means$lat, dr2.means$dr2.predict, xlab="latitude of accession origin", ylab="fecundity in Rathckegården 2012")
abline(lm(dr2.predict~lat, data=dr2.means), col="blue")
```
### Rat 2012 v.significant local adaptation


### 4d. Ramsta

```{r}
dm2 <- d[d$exp=="RAM" & d$year=="2012" ,]
hist(dm2$fecundity)
```

```{r eval=FALSE, include=TRUE}

profile.m2 <- tweedie.profile(fecundity ~ id + block + block*id, data = dm2, p.vec = seq(1.1, 1.9, 0.1), fit.glm = TRUE, control=list(maxit=50), do.plot=TRUE)
print(profile.m2$p.max)
### use this var.power for the tweedie distribution
```

```{r ramsta2012_model}
#vp.up <-profile.m2$p.max
vp.up <- 1.457143
dat.up <- dm2

glm.m21 <- glm(fecundity ~ id + block + block*id, family=tweedie(var.power=vp.up, link.power=0), data=dat.up,maxit=75)

glm.m22 <- glm(fecundity ~ id + block, family=tweedie(var.power=vp.up, link.power=0), data=dat.up, maxit=75)
anova(glm.m22, glm.m21, test="Chisq")

glm.m23 <- glm(fecundity ~ id , family=tweedie(var.power=vp.up, link.power=0), data=dat.up, maxit=75)
anova(glm.m23, glm.m21, test="Chisq")

glm.m24 <- glm(fecundity~ 1 , family=tweedie(var.power=vp.up, link.power=0), data=dat.up, maxit=75)
anova(glm.m23, glm.m24, test="Chisq")

#plot(glm.m23)

```

## neither block nor block interaction is significant.  Use model 3


```{r}
dm2.predict <- predict(glm.m23, type="response")
hist(dm2.predict, main="hist of fitted values Ramsta 2012")

dm2.dat <- cbind(dm2,dm2.predict)
dm2.means <- aggregate(dm2.predict~id, data=dm2.dat, mean)
hist(dm2.means$dm2.predict, main="hist of mean fitted values Ramsta 2012")
```



```{r}
dm2.means <- merge(dm2.means, clim[,1:6], by="id")
lm.m2fit <- lm(dm2.predict~lat, data=dm2.means)
summary(lm.m2fit)
plot(dm2.means$lat, dm2.means$dm2.predict, xlab="latitude of accession origin", ylab="fecundity in Ramsta 2012")
abline(lm(dm2.predict~lat, data=dm2.means), col="blue")
```
Ramsta 2012 significant

### 4e. Combining N/S comparisons

```{r}
all.means2 <- merge(da2.means, dm2.means[,1:2], by="id")
all.means2 <- merge(all.means2, du2.means[,1:2], by="id")
all.means2 <- merge(all.means2, dr2.means[,1:2], by="id")
all.means2 <- all.means2[,c(1,3,4,5,6,7,2,8,9,10)]
pairs(all.means2[,7:10])

all.means2$n.means <- apply(all.means2,1,function(x){(as.numeric(x[7])+as.numeric(x[8]))/2})
all.means2$s.means <- apply(all.means2,1,function(x){(as.numeric(x[9])+as.numeric(x[10]))/2})

lm.ns.mean2 <- lm(n.means~s.means, data=all.means1)
lm.ns.mean.loc2 <- lm(n.means~s.means + region , data=all.means2)
lm.ns.mean.int2 <- lm(n.means~s.means + region + s.means*region , data=all.means2)

anova(lm.ns.mean2)
anova(lm.ns.mean.loc2)
anova(lm.ns.mean.int2)

aov.ns.mean.loc2 <- aov(n.means~s.means + region, data=all.means2)
ns.posthoc2 <- TukeyHSD(aov.ns.mean.loc2, "region")
plot(ns.posthoc2)

with(all.means2, plot(s.means, n.means, xlab="mean fecundity in South", ylab="mean fecundity in North"))
#abline(lm.ns.mean)
abline(0,1, col="blue")

library(ggplot2)

ns.plot2012 <- ggplot(all.means2, aes(s.means, n.means)) + geom_point(aes(colour=region))
ns.plot2012 + theme_bw() +   geom_smooth( aes(linetype = region, color=region), method = "lm") + labs(x="mean fecundity in Southern experiments 2012", y="mean fecundity in Northern experiments 2012")
```
fecundity much higher in the North than the South, N and S accessions fecundity the same



```{r}
lm.n2 <- lm(n.means~lat, data=all.means2)
summary(lm.n2)
plot(all.means2$lat, all.means2$n.means, xlab="latitude of accession origin", ylab="mean fecundity in North 2012")
abline(lm(n.means~lat, data=all.means2), col="blue")

lm.s2 <- lm(s.means~lat, data=all.means2)
summary(lm.s2)
plot(all.means2$lat, all.means2$s.means, xlab="latitude of accession origin", ylab="mean fecundity in South 2012")
abline(lm(s.means~lat, data=all.means2), col="blue")
```
So, relationship between latitude and fecundity is significant in both N and S, but in N, it goes the wrong direction!
Northern accessions don't make as many seeds in the North as the Southerns.
Does this mean that survival is more important in the Northern sites?



## 5. combining 2011 and 2012
From the above, it looks like fecundity patterns are different for the two years, as well as between sites so I want to look at this more closely.
First step is put variables together

```{r}

fit1 <- as.list(1:4)
means1 <-c("du1.means","dr1.means","da1.means","dm1.means")
sites <- c("ULL","RAT","ADA","RAM")

for(up in 1:4){
  up.m <- means1[[up]]
  up.dat <- get(up.m)
  colnames(up.dat)[2] <- "fecundity"
  up.dat$site <- sites[[up]]
  fit1[[up]] <- up.dat
}
fit1 <- do.call(rbind, fit1)
### no observations for line 9336 in Ullstorp in 2011
### no observations 6109 in Adal East in 2011
fit1$year <- 2011

fit2 <- as.list(1:4)
means2 <-c("du2.means","dr2.means","da2.means","dm2.means")
for(up in 1:4){
  up.m <- means2[[up]]
  up.dat <- get(up.m)
  colnames(up.dat)[2] <- "fecundity"
  up.dat$site <- sites[[up]]
  fit2[[up]] <- up.dat
}
fit2 <- do.call(rbind, fit2)
### no observations for line 9336 in Ullstorp in 2011
### no observations 6109 in Adal East in 2011
fit2$year <- 2012

### these observations are complete

acc.fit <- rbind(fit1, fit2)

hist(acc.fit$fecundity, main="predicted mean fecundity, all sites, both years")
with(acc.fit, boxplot(fecundity~year+site, xlab="year and site",ylab="fecundity", las=2))
```
This isn't totally normal, but I'll take it.
Next step is to fit some models to look at year and site effects and interactions

```{r}
lm.b1 <- lm(fecundity~id + year + site + year*site, data=acc.fit)
anova(lm.b1)

lm.b2 <- lm(fecundity~id + year + id*year + site + id*site + year*site, data=acc.fit)
anova(lm.b2)



```

So there are year difference, site differences, and interaction differences. This interaction is actually really, really big.


```{r}

all.means1$year <- "2011"
all.means2$year <- "2012"
all.means.both <- rbind(all.means1[,c(1:6,11:13)], all.means2[,c(1:6,11:13)])

ns.plot.both <- ggplot(all.means.both, aes(s.means, n.means)) + geom_point(aes(colour=region))
ns.plot.both + theme_bw() +   geom_smooth( aes(linetype = year, color= region), method = "lm") + labs(x="mean fecundity in Southern experiments", y="mean fecundity in Northern experiments")
```

## 6.  Output fitted values
Combine all.means, all.means1, all.means2
accessions in rows, phenotypes in columns

```{r}
all.min <- all.means[,c(1,7:12)]
colnames(all.min) <- c("id","ADA_both","RAM_both","ULL_both","RAT_both","N_both", "S_both")

all1.min <- all.means1[,c(1,7:12)]
colnames(all1.min) <- c("id","ADA_2011","RAM_2011","ULL_2011","RAT_2011","N_2011", "S_2011")

all2.min <- all.means2[,c(1,7:12)]
colnames(all2.min) <- c("id","ADA_2012","RAM_2012","ULL_2012","RAT_2012","N_2012", "S_2012")

fecund.fitted.values <- merge(all.min, all1.min, by="id", all=TRUE)
fecund.fitted.values <- merge(fecund.fitted.values, all2.min, by="id", all=TRUE)
write.table(fecund.fitted.values, file="./data/fecund.fitted.values.txt", quote=FALSE)
```


### 7. Output linear model summaries
output to generate summary figure
```{r}
#need pvalue, slope, adjr2 for all four sites for both years- easiest is to just output all models together

models <- c("lm.a1fit","lm.m1fit","lm.r1fit","lm.u1fit","lm.a2fit","lm.m2fit","lm.r2fit","lm.u2fit")
fecundity.adapt.models <- as.list(1:8)
for(up in 1:8){fecundity.adapt.models[[up]] <- get(models[up])}
save(fecundity.adapt.models, file="./data/fecundity.adapt.models.Rdat")
```



## 8.  Summary

Fecundity shows local adaptation at all sites and years in the South.
In the North, almost all nonsig or sig with the effect going in the opposite direction
(i.e. Southern accessions more fecund than Northern)
Only exception is Ramsta 2011.

So does this mean that survival is what impacts fitness in the North?
Is this overwinter survival or survival from winter to seed set?
And does this make sense with the year effects (i.e. stronger in 2012 in North)?

Evidence of a tradeoff fecundity/survival?