---
title: "05.CG.sss.accession.models.Rmd"
author: "Daniele Filiault"
date: "2/27/2019"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Description and Motivation
Taking a closer look at sss in the common garden experiments.
Taking same approach as for fitness and fecundity
Look for local adaptation and try to assess which components of sss
vary across sites and years to cause fitness differences.

Fit models to get sss for each site
First make data variable

## 1. Local adaptation with sss values - combined years
### 1a. prep data
```{r}
d11=readRDS(file="./data/d11_for_analysis.rds")
d12=readRDS(file="./data/d12_for_analysis.rds")

##restrict to sss
traits=c("sss")

d11$year=2011
d12$year=2012

d11=d11[, c("year", "exp", "block", "id", "sss")]
d12=d12[, c("year", "exp", "block", "id", "sss")]

d=rbind(d11, d12)
d$region[d$exp%in%c("ULL", "RAT")]="S"
d$region[d$exp%in%c("RAM", "ADA")]="N"
d$year=as.factor(d$year)
d=d[, c("year", "region", "exp","block", "id", "sss")]
d=na.omit(d)

hist(d$sss)
with(d, table(sss,year,exp))
```
So the distribution of the sss variable clearly binary - use glm with binomial/logit link function

```{r libraries}
library(lme4)

```

###1b. start with Ullstorp

```{r}
du <- d[d$exp=="ULL",]
hist(du$sss)
```


Fit models, assess model components

```{r ull_model}

dat.up <- du

glm.u6 <- glm(sss ~ id, family=binomial, data=dat.up)

glm.u7 <- glm(sss ~ 1, family=binomial, data=dat.up)
anova(glm.u6, glm.u7, test="Chisq")

glm.u8 <- glm(sss ~ id + year, family=binomial, data=dat.up)
anova(glm.u8, glm.u6, test="Chisq")

glm.u9 <- glm(sss ~ id + year + id* year, family=binomial, data=dat.up)
anova(glm.u8, glm.u9, test="Chisq")  # not significant

glm.u10 <- glm(sss ~ id + year +year/block, family=binomial, data=dat.up)
anova(glm.u8, glm.u10, test="Chisq")  # block effect significant

glm.u11 <- glm(sss ~ id + year + id*year +year/block, family=binomial, data=dat.up)
anova(glm.u11, glm.u10, test="Chisq")  # not significant
```

The best model is glm.u6, which simply has id in it.



```{r}
du.predict <- predict(glm.u6, type="response")
hist(du.predict, main="predicted values Ullstorp")

du.dat <- cbind(du,du.predict)
du.means <- aggregate(du.predict~id, data=du.dat, mean)
hist(du.means$du.predict, main="predicted means Ullstorp", xlab="sss")

```
continue

check local adaptation

```{r}
clim=read.table("./data/worldclim_swedish_acc.txt", sep="\t", h=T, stringsAsFactor=T)
clim_exp=read.table("./data/worldclim_swedish_exp.txt", sep="\t", h=T)
colnames(clim)[1]="id"

du.means <- merge(du.means, clim[,1:6], by="id")
lm.ufit <- lm(du.predict~long, data=du.means)
summary(lm.ufit)
plot(du.means$lat, du.means$du.predict, xlab="latitude of accession origin", ylab="sss in Ullstorp")
abline(lm(du.predict~lat, data=du.means), col="blue")
```
no patterns here - only a few accessions that had low survival


###1c. Adal East

```{r}
da <- d[d$exp=="ADA",]
hist(da$sss, main="Adal sss", xlab="sss")
```


Now fit models, assess model components

```{r adal_east_model}
dat.up <- da
  
glm.a6 <- glm(sss ~ id, family=binomial, data=dat.up)

glm.a7 <- glm(sss ~ 1, family=binomial, data=dat.up)
anova(glm.a6, glm.a7, test="Chisq")

glm.a8 <- glm(sss ~ id + year, family=binomial, data=dat.up)
anova(glm.a8, glm.a6, test="Chisq")

glm.a9 <- glm(sss ~ id + year + id* year, family=binomial, data=dat.up)
anova(glm.a8, glm.a9, test="Chisq")  # not significant

glm.a10 <- glm(sss ~ id + year +year/block, family=binomial, data=dat.up)
anova(glm.a8, glm.a10, test="Chisq")  # block effect significant

glm.a11 <- glm(sss ~ id + year + id*year +year/block, family=binomial, data=dat.up)
anova(glm.a11, glm.a10, test="Chisq")  # not significant
```

Take model 11

```{r}
da.predict <- predict(glm.a11, type="response")
hist(da.predict)

da.dat <- cbind(da,da.predict)
da.means <- aggregate(da.predict~id, data=da.dat, mean)
hist(da.means$da.predict, main="hist of predicted means")
```

check local adaptation

```{r}
da.means <- merge(da.means, clim[,1:6], by="id")
lm.afit <- lm(da.predict~lat, data=da.means)
summary(lm.afit)
plot(da.means$lat, da.means$da.predict, xlab="latitude of accession origin", ylab="sss in Ådal East")
abline(lm(da.predict~lat, data=da.means), col="blue")
```

marginally sig local adaptation Adal East



###1d. Rathckegarden

```{r}
dr <- d[d$exp=="RAT",]
hist(dr$sss, main="histogram of Rat sss")
```



Now fit models, assess model components

```{r rathckegarden_model}

dat.up <- dr

glm.r6 <- glm(sss ~ id, family=binomial, data=dat.up)

glm.r7 <- glm(sss ~ 1, family=binomial, data=dat.up)
anova(glm.r6, glm.r7, test="Chisq")

glm.r8 <- glm(sss ~ id + year, family=binomial, data=dat.up)
anova(glm.r8, glm.r6, test="Chisq")

glm.r9 <- glm(sss ~ id + year + id* year, family=binomial, data=dat.up)
anova(glm.r8, glm.r9, test="Chisq")  # not significant

glm.r10 <- glm(sss ~ id + year +year/block, family=binomial, data=dat.up)
anova(glm.r8, glm.r10, test="Chisq")  # block effect significant

glm.r11 <- glm(sss ~ id + year + id*year +year/block, family=binomial, data=dat.up)
anova(glm.r11, glm.r10, test="Chisq")  # not significant

#plot(glm.r1)

```

Again, model 11, full model with interaction

```{r}
dr.predict <- predict(glm.r11, type="response")
hist(dr.predict, main="hist of fitted values Rat")

dr.dat <- cbind(dr,dr.predict)
dr.means <- aggregate(dr.predict~id, data=dr.dat, mean)
hist(dr.means$dr.predict, main="hist of mean fitted values Rat")
```



check local adaptation

```{r}
dr.means <- merge(dr.means, clim[,1:6], by="id")
lm.rfit <- lm(dr.predict~lat, data=dr.means)
summary(lm.rfit)
plot(dr.means$lat, dr.means$dr.predict, xlab="latitude of accession origin", ylab="sss in Ratheckgården")
abline(lm(dr.predict~lat, data=dr.means), col="blue")
```
local adaptation Rathckegarden!

###1e. Ramsta

```{r}
dm <- d[d$exp=="RAM",]
hist(dm$sss, main="histogram of Ramsta sss")
```

Now fit models, assess model components


```{r ramsta_model}

dat.up <- dm

glm.m6 <- glm(sss ~ id, family=binomial, data=dat.up)

glm.m7 <- glm(sss ~ 1, family=binomial, data=dat.up)
anova(glm.m6, glm.m7, test="Chisq")

glm.m8 <- glm(sss ~ id + year, family=binomial, data=dat.up)
anova(glm.m8, glm.m6, test="Chisq")

glm.m9 <- glm(sss ~ id + year + id* year, family=binomial, data=dat.up)
anova(glm.m8, glm.m9, test="Chisq")  # not significant

glm.m10 <- glm(sss ~ id + year +year/block, family=binomial, data=dat.up)
anova(glm.m8, glm.m10, test="Chisq")  # block effect significant

glm.m11 <- glm(sss ~ id + year + id*year +year/block, family=binomial, data=dat.up)
anova(glm.m11, glm.m10, test="Chisq")  # not significant


#plot(glm.m1)

```

Here, taking the full model - glm.m11

```{r}
dm.predict <- predict(glm.m11, type="response")
hist(dm.predict, main="hist of fitted values Ramsta")

dm.dat <- cbind(dm,dm.predict)
dm.means <- aggregate(dm.predict~id, data=dm.dat, mean)
hist(dm.means$dm.predict, main="hist of mean fitted values Ramsta")
```

check local adaptation

```{r}
dm.means <- merge(dm.means, clim[,1:6], by="id")
lm.mfit <- lm(dm.predict~lat, data=dm.means)
summary(lm.mfit)
plot(dm.means$lat, dm.means$dm.predict, xlab="latitude of accession origin", ylab="sss in Ramsta")
abline(lm(dm.predict~lat, data=dm.means), col="blue")
```
local adaptation in Ramsta

###1f. Combining N/S comparisons

```{r}
all.means <- merge(da.means, dm.means[,1:2], by="id")
all.means <- merge(all.means, du.means[,1:2], by="id")
all.means <- merge(all.means, dr.means[,1:2], by="id")
all.means <- all.means[,c(1,3,4,5,6,7,2,8,9,10)]
pairs(all.means[,7:10])

all.means$n.means <- apply(all.means,1,function(x){(as.numeric(x[7])+as.numeric(x[8]))/2})
all.means$s.means <- apply(all.means,1,function(x){(as.numeric(x[9])+as.numeric(x[10]))/2})

lm.ns.mean <- lm(n.means~s.means, data=all.means)
lm.ns.mean.loc <- lm(n.means~s.means + region , data=all.means)
lm.ns.mean.int <- lm(n.means~s.means + region + s.means*region , data=all.means)

summary(lm.ns.mean)
summary(lm.ns.mean.loc)
summary(lm.ns.mean.int)

aov.ns.mean.loc <- aov(n.means~s.means + region, data=all.means)
ns.posthoc <- TukeyHSD(aov.ns.mean.loc, "region")
plot(ns.posthoc)

with(all.means, plot(s.means, n.means, xlab="mean sss in South", ylab="mean sss in North"))
#abline(lm.ns.mean)
abline(0,1, col="blue")

library(ggplot2)

ns.plot <- ggplot(all.means, aes(s.means, n.means)) + geom_point(aes(colour=region))
ns.plot + theme_bw() +   geom_smooth( aes(linetype = region, color=region), method = "lm") + labs(x="mean sss in Southern experiments", y="mean sss in Northern experiments")

lm.nfit <- lm(n.means~lat, data=all.means)
summary(lm.nfit)
plot(all.means$lat, all.means$n.means, xlab="latitude of accession origin", ylab=" mean sss in Northern sites")
abline(lm(n.means~lat, data=all.means), col="blue")

lm.sfit <- lm(s.means~lat, data=all.means)
summary(lm.sfit)
plot(all.means$lat, all.means$s.means, xlab="latitude of accession origin", ylab=" mean sss in Southern sites")
abline(lm(s.means~lat, data=all.means), col="blue")

```
1.  Joint years, local adaptation in S, not in N.
2.  generally higher sss in N

## 3. Local adaptation with sss values - 2011
###3a. start with Ullstorp

```{r}
du1 <- d[d$exp=="ULL" & d$year=="2011" ,]
hist(du1$sss)
```


Now fit models, assess model components

```{r ullstorp2011_model}

dat.up <- du1

glm.u11 <- glm(sss ~ 1, family=binomial, data=dat.up)

glm.u12 <- glm(sss ~ id, family=binomial, data=dat.up)
anova(glm.u12, glm.u11, test="Chisq")

glm.u13 <- glm(sss ~ id + block, family=binomial, data=dat.up)
anova(glm.u13, glm.u12, test="Chisq")


#plot(glm.a11)

```
No block effect, use model u12

 
```{r}
du1.predict <- predict(glm.u12, type="response")
hist(du1.predict, main="hist of fitted values Ullstorp 2011")

du1.dat <- cbind(du1,du1.predict)
du1.means <- aggregate(du1.predict~id, data=du1.dat, mean)
hist(du1.means$du1.predict, main="hist of mean fitted values Ullstorp 2011")
```

```{r}
du1.means <- merge(du1.means, clim[,1:6], by="id")
lm.u1fit <- lm(du1.predict~lat, data=du1.means)
summary(lm.u1fit)
plot(du1.means$lat, du1.means$du1.predict, xlab="latitude of accession origin", ylab="sss in Ullstorp 2011")
abline(lm(du1.predict~lat, data=du1.means), col="blue")
```

no local adaptation Ull 2011, again, no big survival differences


###3b. Adal East 2011

```{r}
da1 <- d[d$exp=="ADA" & d$year=="2011" ,]
hist(da1$sss)
```

model fitting

```{r adaleast2011_model}

dat.up <- da1

glm.a11 <- glm(sss ~ 1, family=binomial, data=dat.up)

glm.a12 <- glm(sss ~ id, family=binomial, data=dat.up)
anova(glm.a12, glm.a11, test="Chisq")

glm.a13 <- glm(sss ~ id + block, family=binomial, data=dat.up)
anova(glm.a13, glm.a12, test="Chisq")


#plot(glm.a11)

```
All terms significant

 
```{r}
da1.predict <- predict(glm.a13, type="response")
hist(da1.predict, main="hist of fitted values Ådal East 2011")

da1.dat <- cbind(da1,da1.predict)
da1.means <- aggregate(da1.predict~id, data=da1.dat, mean)
hist(da1.means$da1.predict, main="hist of mean fitted values Ådal East 2011")
```

```{r}
da1.means <- merge(da1.means, clim[,1:6], by="id")
lm.a1fit <- lm(da1.predict~lat, data=da1.means)
summary(lm.a1fit)
plot(da1.means$lat, da1.means$da1.predict, xlab="latitude of accession origin", ylab="sss in Ådal East 2011")
abline(lm(da1.predict~lat, data=da1.means), col="blue")
```
Adal East 2011 marginal

###3c. Rat 2011

```{r}
dr1 <- d[d$exp=="RAT" & d$year=="2011" ,]
hist(dr1$sss)
```


```{r rat2011_model}
dat.up <- dr1

glm.r11 <- glm(sss ~ 1, family=binomial, data=dat.up)

glm.r12 <- glm(sss ~ id, family=binomial, data=dat.up)
anova(glm.r12, glm.r11, test="Chisq")

glm.r13 <- glm(sss ~ id + block, family=binomial, data=dat.up)
anova(glm.r13, glm.r12, test="Chisq")


#plot(glm.r11)

```

all terms significant

```{r}
dr1.predict <- predict(glm.r13, type="response")
hist(dr1.predict, main="hist of fitted values Rathckegården 2011")

dr1.dat <- cbind(dr1,dr1.predict)
dr1.means <- aggregate(dr1.predict~id, data=dr1.dat, mean)
hist(dr1.means$dr1.predict, main="hist of mean fitted values Rathckegården 2011")
```



```{r}
dr1.means <- merge(dr1.means, clim[,1:6], by="id")
lm.r1fit <- lm(dr1.predict~lat, data=dr1.means)
summary(lm.r1fit)
plot(dr1.means$lat, dr1.means$dr1.predict, xlab="latitude of accession origin", ylab="sss in Rathckegården 2011")
abline(lm(dr1.predict~lat, data=dr1.means), col="blue")
```
Rat 2011  significant

###3d. Ramsta 2011
```{r}
dm1 <- d[d$exp=="RAM" & d$year=="2011" ,]
hist(dm1$sss)
```



```{r ramsta2011_model}
dat.up <- dm1

glm.m11 <- glm(sss ~ 1, family=binomial, data=dat.up)

glm.m12 <- glm(sss ~ id, family=binomial, data=dat.up)
anova(glm.m12, glm.m11, test="Chisq")

glm.m13 <- glm(sss ~ id + block, family=binomial, data=dat.up)
anova(glm.m13, glm.m12, test="Chisq")

#plot(glm.m11)

```

Use full model

```{r}
dm1.predict <- predict(glm.m13, type="response")
hist(dm1.predict, main="hist of fitted values Ramsta 2011")

dm1.dat <- cbind(dm1,dm1.predict)
dm1.means <- aggregate(dm1.predict~id, data=dm1.dat, mean)
hist(dm1.means$dm1.predict, main="hist of mean fitted values Ramsta 2011")
```



```{r}
dm1.means <- merge(dm1.means, clim[,1:6], by="id")
lm.m1fit <- lm(dm1.predict~lat, data=dm1.means)
summary(lm.m1fit)
plot(dm1.means$lat, dm1.means$dm1.predict, xlab="latitude of accession origin", ylab="sss in Ramsta 2011")
abline(lm(dm1.predict~lat, data=dm1.means), col="blue")
```
Ramsta 2011 very significant!

###3e. N/S combined 2011

```{r}
all.means1 <- merge(da1.means, dm1.means[,1:2], by="id")
all.means1 <- merge(all.means1, du1.means[,1:2], by="id")
all.means1 <- merge(all.means1, dr1.means[,1:2], by="id")
all.means1 <- all.means1[,c(1,3,4,5,6,7,2,8,9,10)]
pairs(all.means1[,7:10])

all.means1$n.means <- apply(all.means1,1,function(x){(as.numeric(x[7])+as.numeric(x[8]))/2})
all.means1$s.means <- apply(all.means1,1,function(x){(as.numeric(x[9])+as.numeric(x[10]))/2})

lm.ns.mean1 <- lm(n.means~s.means, data=all.means1)
lm.ns.mean.loc1 <- lm(n.means~s.means + region , data=all.means1)
lm.ns.mean.int1 <- lm(n.means~s.means + region + s.means*region , data=all.means1)

anova(lm.ns.mean1)
anova(lm.ns.mean.loc1)
anova(lm.ns.mean.int1)

aov.ns.mean.loc1 <- aov(n.means~s.means + region, data=all.means1)
ns.posthoc1 <- TukeyHSD(aov.ns.mean.loc1, "region")
plot(ns.posthoc1)

with(all.means1, plot(s.means, n.means, xlab="mean sss in South", ylab="mean sss in North"))
#abline(lm.ns.mean)
abline(0,1, col="blue")

library(ggplot2)

ns.plot2011 <- ggplot(all.means1, aes(s.means, n.means)) + geom_point(aes(colour=region))
ns.plot2011 + theme_bw() +   geom_smooth( aes(linetype = region, color=region), method = "lm") + labs(x="mean sss in Southern experiments 2011", y="mean sss in Northern experiments 2011")
```

Do "classic" local adaptation on these means

```{r}

lm.n1 <- lm(n.means~lat, data=all.means1)
summary(lm.n1)
plot(all.means1$lat, all.means1$n.means, xlab="latitude of accession origin", ylab="mean sss in North 2011")
abline(lm(n.means~lat, data=all.means1), col="blue")

lm.s1 <- lm(s.means~lat, data=all.means1)
summary(lm.s1)
plot(all.means1$lat, all.means1$s.means, xlab="latitude of accession origin", ylab="mean sss in South 2011")
abline(lm(s.means~lat, data=all.means1), col="blue")

```
South is  significant, North is not


## 4. Local adaptation with sss values - 2012
###4a. start with Ullstorp

```{r}
du2 <- d[d$exp=="ULL" & d$year=="2012" ,]
hist(du2$sss)
```
determine d for tweedie distribution


Now fit models, assess model components

```{r ullstorp2012_model}
dat.up <- du2

glm.u21 <- glm(sss ~ 1, family=binomial, data=dat.up)

glm.u22 <- glm(sss ~ id, family=binomial, data=dat.up)
anova(glm.u22, glm.u21, test="Chisq")

glm.u23 <- glm(sss ~ id + block, family=binomial, data=dat.up)
anova(glm.u23, glm.u22, test="Chisq")


#plot(glm.a21)

```
Use model 2

 
```{r}
du2.predict <- predict(glm.u22, type="response")
hist(du2.predict, main="hist of fitted values Ullstorp 2012")

du2.dat <- cbind(du2,du2.predict)
du2.means <- aggregate(du2.predict~id, data=du2.dat, mean)
hist(du2.means$du2.predict, main="hist of mean fitted values Ullstorp 2012")
```

```{r}
du2.means <- merge(du2.means, clim[,1:6], by="id")
lm.u2fit <- lm(du2.predict~lat, data=du2.means)
summary(lm.u2fit)
plot(du2.means$lat, du2.means$du2.predict, xlab="latitude of accession origin", ylab="sss in Ullstorp 2012")
abline(lm(du2.predict~lat, data=du2.means), col="blue")
```
Nothing Ullstorp 2012 - very high survival


###4b. Adal East
```{r}
da2 <- d[d$exp=="ADA" & d$year=="2012" ,]
hist(da2$sss)
```


```{r adaleast2012_model}
dat.up <- da2

glm.a21 <- glm(sss ~ 1, family=binomial, data=dat.up)

glm.a22 <- glm(sss ~ id, family=binomial, data=dat.up)
anova(glm.a22, glm.a21, test="Chisq")

glm.a23 <- glm(sss ~ id + block, family=binomial, data=dat.up)
anova(glm.a23, glm.a22, test="Chisq")



#plot(glm.a22)

```

Model 3

```{r}
da2.predict <- predict(glm.a23, type="response")
hist(da2.predict, main="hist of fitted values Ådal East 2012")

da2.dat <- cbind(da2,da2.predict)
da2.means <- aggregate(da2.predict~id, data=da2.dat, mean)
hist(da2.means$da2.predict, main="hist of mean fitted values Ådal East 2012")
```

```{r}
da2.means <- merge(da2.means, clim[,1:6], by="id")
lm.a2fit <- lm(da2.predict~lat, data=da2.means)
summary(lm.a2fit)
plot(da2.means$lat, da2.means$da2.predict, xlab="latitude of accession origin", ylab="sss in Ådal East 2012")
abline(lm(da2.predict~lat, data=da2.means), col="blue")
```
Adal East 2012 not significant 
very high survival

### 4c. Rathckegarden


```{r}
dr2 <- d[d$exp=="RAT" & d$year=="2012" ,]
hist(dr2$sss)
```


```{r rat2012_model}
dat.up <- dr2

glm.r21 <- glm(sss ~ 1, family=binomial, data=dat.up)

glm.r22 <- glm(sss ~ id, family=binomial, data=dat.up)
anova(glm.r22, glm.r21, test="Chisq")

glm.r23 <- glm(sss ~ id + block, family=binomial, data=dat.up)
anova(glm.r23, glm.r22, test="Chisq")


#plot(glm.r21)

```

Use model 3

```{r}
dr2.predict <- predict(glm.r22, type="response")
hist(dr2.predict, main="hist of fitted values Rathckegården 2012")

dr2.dat <- cbind(dr2,dr2.predict)
dr2.means <- aggregate(dr2.predict~id, data=dr2.dat, mean)
hist(dr2.means$dr2.predict, main="hist of mean fitted values Rathckegården 2012")
```



```{r}
dr2.means <- merge(dr2.means, clim[,1:6], by="id")
lm.r2fit <- lm(dr2.predict~lat, data=dr2.means)
summary(lm.r2fit)
plot(dr2.means$lat, dr2.means$dr2.predict, xlab="latitude of accession origin", ylab="sss in Rathckegården 2012")
abline(lm(dr2.predict~lat, data=dr2.means), col="blue")
```
### Rat 2012 v.significant local adaptation


### 4d. Ramsta

```{r}
dm2 <- d[d$exp=="RAM" & d$year=="2012" ,]
hist(dm2$sss)
```



```{r ramsta2012_model}
dat.up <- dm2

glm.m21 <- glm(sss ~ 1, family=binomial, data=dat.up)

glm.m22 <- glm(sss ~ id, family=binomial, data=dat.up)
anova(glm.m22, glm.m21, test="Chisq")

glm.m23 <- glm(sss ~ id + block, family=binomial, data=dat.up)
anova(glm.m23, glm.m22, test="Chisq")

glm.m24 <- glm(sss ~ block, family=binomial, data=dat.up)
anova(glm.m23, glm.m24, test="Chisq")


#plot(glm.m23)

```

Again, not a lot significant.  Use model 3 (which is better than model 2 - anova also shsss some line effects)


```{r}
dm2.predict <- predict(glm.m23, type="response")
hist(dm2.predict, main="hist of fitted values Ramsta 2012")

dm2.dat <- cbind(dm2,dm2.predict)
dm2.means <- aggregate(dm2.predict~id, data=dm2.dat, mean)
hist(dm2.means$dm2.predict, main="hist of mean fitted values Ramsta 2012")
```



```{r}
dm2.means <- merge(dm2.means, clim[,1:6], by="id")
lm.m2fit <- lm(dm2.predict~lat, data=dm2.means)
summary(lm.m2fit)
plot(dm2.means$lat, dm2.means$dm2.predict, xlab="latitude of accession origin", ylab="sss in Ramsta 2012")
abline(lm(dm2.predict~lat, data=dm2.means), col="blue")
```
Ramsta 2012 not significant

### 4e. Combining N/S comparisons

```{r}
all.means2 <- merge(da2.means, dm2.means[,1:2], by="id")
all.means2 <- merge(all.means2, du2.means[,1:2], by="id")
all.means2 <- merge(all.means2, dr2.means[,1:2], by="id")
all.means2 <- all.means2[,c(1,3,4,5,6,7,2,8,9,10)]
pairs(all.means2[,7:10])

all.means2$n.means <- apply(all.means2,1,function(x){(as.numeric(x[7])+as.numeric(x[8]))/2})
all.means2$s.means <- apply(all.means2,1,function(x){(as.numeric(x[9])+as.numeric(x[10]))/2})

lm.ns.mean2 <- lm(n.means~s.means, data=all.means1)
lm.ns.mean.loc2 <- lm(n.means~s.means + region , data=all.means2)
lm.ns.mean.int2 <- lm(n.means~s.means + region + s.means*region , data=all.means2)

anova(lm.ns.mean2)
anova(lm.ns.mean.loc2)
anova(lm.ns.mean.int2)

aov.ns.mean.loc2 <- aov(n.means~s.means + region, data=all.means2)
ns.posthoc2 <- TukeyHSD(aov.ns.mean.loc2, "region")
plot(ns.posthoc2)

with(all.means2, plot(s.means, n.means, xlab="mean sss in South", ylab="mean sss in North"))
#abline(lm.ns.mean)
abline(0,1, col="blue")

library(ggplot2)

ns.plot2012 <- ggplot(all.means2, aes(s.means, n.means)) + geom_point(aes(colour=region))
ns.plot2012 + theme_bw() +   geom_smooth( aes(linetype = region, color=region), method = "lm") + labs(x="mean sss in Southern experiments 2012", y="mean sss in Northern experiments 2012")
```



```{r}
lm.n2 <- lm(n.means~lat, data=all.means2)
summary(lm.n2)
plot(all.means2$lat, all.means2$n.means, xlab="latitude of accession origin", ylab="mean sss in North 2012")
abline(lm(n.means~lat, data=all.means2), col="blue")

lm.s2 <- lm(s.means~lat, data=all.means2)
summary(lm.s2)
plot(all.means2$lat, all.means2$s.means, xlab="latitude of accession origin", ylab="mean sss in South 2012")
abline(lm(s.means~lat, data=all.means2), col="blue")
```
So, relationship between latitude and sss is significant in both N and S, but in N, it goes the wrong direction!
Northern accessions don't make as many seeds in the North as the Southerns.
Does this mean that survival is more important in the Northern sites?



## 5. combining 2011 and 2012
From the above, it looks like sss patterns are different for the two years, as well as between sites so I want to look at this more closely.
First step is put variables together

```{r}

fit1 <- as.list(1:4)
means1 <-c("du1.means","dr1.means","da1.means","dm1.means")
sites <- c("ULL","RAT","ADA","RAM")

for(up in 1:4){
  up.m <- means1[[up]]
  up.dat <- get(up.m)
  colnames(up.dat)[2] <- "sss"
  up.dat$site <- sites[[up]]
  fit1[[up]] <- up.dat
}
fit1 <- do.call(rbind, fit1)
### no observations for line 9336 in Ullstorp in 2011
### no observations 6109 in Adal East in 2011
fit1$year <- 2011

fit2 <- as.list(1:4)
means2 <-c("du2.means","dr2.means","da2.means","dm2.means")
for(up in 1:4){
  up.m <- means2[[up]]
  up.dat <- get(up.m)
  colnames(up.dat)[2] <- "sss"
  up.dat$site <- sites[[up]]
  fit2[[up]] <- up.dat
}
fit2 <- do.call(rbind, fit2)

fit2$year <- 2012


acc.fit <- rbind(fit1, fit2)

hist(acc.fit$sss, main="predicted mean sss, all sites, both years")
with(acc.fit, boxplot(sss~year+site, xlab="year and site",ylab="sss", las=2))
```
This isn't totally normal, but I'll take it.
Next step is to fit some models to look at year and site effects and interactions

```{r}
lm.b1 <- lm(sss~id + year + site + year*site, data=acc.fit)
anova(lm.b1)

lm.b2 <- lm(sss~id + year + id*year + site + id*site + year*site, data=acc.fit)
anova(lm.b2)



```

So there are year difference, site differences, and interaction differences. This interaction is actually really, really big.


```{r}

all.means1$year <- "2011"
all.means2$year <- "2012"
all.means.both <- rbind(all.means1[,c(1:6,11:13)], all.means2[,c(1:6,11:13)])

ns.plot.both <- ggplot(all.means.both, aes(s.means, n.means)) + geom_point(aes(colour=region))
ns.plot.both + theme_bw() +   geom_smooth( aes(linetype = year, color= region), method = "lm") + labs(x="mean sss in Southern experiments", y="mean sss in Northern experiments")
```

sss patterns mimic ows.  Presumably most of this is due to ows.

## 6.  Output fitted values
Combine all.means, all.means1, all.means2
accessions in rows, phenotypes in columns

```{r}
all.min <- all.means[,c(1,7:12)]
colnames(all.min) <- c("id","ADA_both","RAM_both","ULL_both","RAT_both","N_both", "S_both")

all1.min <- all.means1[,c(1,7:12)]
colnames(all1.min) <- c("id","ADA_2011","RAM_2011","ULL_2011","RAT_2011","N_2011", "S_2011")

all2.min <- all.means2[,c(1,7:12)]
colnames(all2.min) <- c("id","ADA_2012","RAM_2012","ULL_2012","RAT_2012","N_2012", "S_2012")

sss.fitted.values <- merge(all.min, all1.min, by="id", all=TRUE)
sss.fitted.values <- merge(sss.fitted.values, all2.min, by="id", all=TRUE)
write.table(sss.fitted.values, file="./data/sss.fitted.values.txt", quote=FALSE)
```

### 7. Output linear model summaries
output to generate summary figure
```{r}
#need pvalue, slope, adjr2 for all four sites for both years- easiest is to just output all models together

models <- c("lm.a1fit","lm.m1fit","lm.r1fit","lm.u1fit","lm.a2fit","lm.m2fit","lm.r2fit","lm.u2fit")
sss.adapt.models <- as.list(1:8)
for(up in 1:8){sss.adapt.models[[up]] <- get(models[up])}
save(sss.adapt.models, file="./data/sss.adapt.models.Rdat")
```

## 8.  Summary
sss shows the same patterns as ows, suggesting the ows is the primary determinant of sss.
Look at correlations to check this

```{r}
ows <- read.table("./data/ows.fitted.values.txt")
sss <- read.table("./data/sss.fitted.values.txt")
for(up in c(8:11,14:17)){
  plot(sss[,up],ows[,up], ylab="overwinter survival", xlab="survival to seed set", main=colnames(sss)[up])
  fit  <- lm(ows[,up]~sss[,up])
  abline(fit, col="blue")
  abline(0,1,col="red")
  summary(fit)
}

```
