---
title: "12.climate.niche.bioclim.v2"
author: "Daniele Filiault"
date: "8/5/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library("RColorBrewer")
```

## 1. Rationale.

So it looks like N/S selection pressure is significant, but not 100% of differences in local adaptation
Can we identify others with climate parameters?  This is a redo/continuation of script 10, but using the newer v2.0 bioclim data
which was extracted in /Volumes/field_experiments/001.common.reference.files/008.climate.data/extract.climate.data.Rmd

```{r data prep}
# experimental accession data
c.acc <- read.table('./data/bioclim.v2.200.experimental.lines.txt',stringsAsFactors=FALSE, header=TRUE)  ### all accessions now have data
# experiment data
c.exp <- read.table('./data/bioclim.v2.200.experimental.sites.txt', stringsAsFactors=FALSE, header=TRUE)
# 1001g data
all.acc <- read.table("./data/bioclim.v2.200.1001g.lines.txt", stringsAsFactors=FALSE, header=TRUE)


### need to merge these so can use together
### columns: name, region, type, lat, long, climate.params....

### accession data first
# get region data
reg.dat <- read.table("./data/worldclim_swedish_acc.txt", stringsAsFactors=FALSE, header=TRUE, sep="\t")
c.acc <- merge(c.acc, reg.dat[,c(1,3)])
c.acc$type <- "exp.acc"
c.acc <- c.acc[, c(1,23,24,2:22)]
colnames(c.acc)[1] <- "name"
rownames(c.acc) <- c.acc$name

### experimental data
c.exp$region <- c("N Sweden", "N Sweden", "N Sweden","N Sweden", "N Sweden", "S Sweden","S Sweden", "N Sweden", "N Sweden", "S Sweden","S Sweden","S Sweden","S Sweden","S Sweden","S Sweden")
### only use experiments in this analysis
c.exp <- c.exp[c(1,2,4,9,10,11,12,15),]
c.exp$type <- substr(rownames(c.exp),1,1)
c.exp$type <- gsub("C", "common_garden", c.exp$type)
c.exp$type <- gsub("s", "snr", c.exp$type)
c.exp <- c.exp[,c(1,23,24,2:22)]

colnames(c.exp)[1] <- "name"
colnames(c.exp)==colnames(c.acc)

### OK, these look safe to rbind.

c.dat <- rbind(c.exp, c.acc)

## also do 1001g
all.acc$type <- "1001g"
all.acc$region <- "hap.filler"   ## put haplogroup here
all.acc <- all.acc[,c(1,23,24,2:22)]
colnames(all.acc)[1] <- "name"
all.acc <- all.acc[all.acc$name%in%c.dat$name == FALSE,]
all.dat <- rbind(all.acc, c.dat)

```


## 2. PCA analysis of accession data

can we define the localness of accessions by "climate space"?

```{r acc climate space PCA}
library(corrplot)
corrplot(cor(c.dat[,4:24]), method="color", type="upper", diag=FALSE)

pdf("./figures/PCA.bioclimv2.vars.pdf", width=10, height=10)
corrplot(cor(c.dat[,4:24]), method="ellipse", type="upper", diag=FALSE)
dev.off()


## clearly a LOT of correlation here.  Good data for PCA


library(ggfortify)

### only use 19 climate variables

up.dat <- c.dat[,6:24]
rownames(up.dat) <- rownames(c.dat)
mini.pca <- prcomp(up.dat, center=TRUE)
mini.pca
summary(mini.pca)

library(ggfortify)

autoplot(prcomp(up.dat, center=TRUE), data = c.dat, loadings = TRUE, colour="region",size="type")

### I like this analysis using only the 19 climate variables
### let's look if locally adaptive

pcs <- as.data.frame(mini.pca$x)
pcs$id <- rownames(pcs)

write.table(pcs, file="./data/experimental.pcas.txt",quote=FALSE)
```


As before first two PCs explain almost all the variation:
PC1 0.6711 BIO12 Annual Precipitation
PC2 0.3143 BIO4 Temperature Seasonality (standard deviation *100)


```{R mini PCA plot by hand for annotation}
pdf("./figures/PCA.accessions.19climvars.bioclimv2.pdf", width=8, height=6)
plot(pcs$PC1, pcs$PC2, xlab="PC1 (67.11%)", ylab="PC2 (31.43%)", type="n")
n.accs <- pcs[rownames(pcs)%in%rownames(c.dat[c.dat$region=="N Sweden",]),]
s.accs <- pcs[rownames(pcs)%in%rownames(c.dat[c.dat$region=="S Sweden",]),]
c.accs <- pcs[rownames(pcs)%in%rownames(c.dat[c.dat$region=="C Sweden",]),]
points(n.accs$PC1, n.accs$PC2, pch=19, col="dodgerblue")
points(c.accs$PC1, c.accs$PC2, pch=19, col="purple")
points(s.accs$PC1, s.accs$PC2, pch=19, col="green")
exp.sites <- pcs[rownames(pcs)%in%rownames(c.dat[c.dat$type%in%c("common_garden", "snr"),]),]
points(exp.sites$PC1, exp.sites$PC2, pch=23, cex=2.5, lwd=2)
legend("bottomleft", col=c("green", "purple","blue","black"),pch=c(19,19,19,23), legend=c("Southern lines", "Central lines", "Northern lines", "Experiment sites"))
text(exp.sites[,1], exp.sites[,2], labels=c("N2-C","N2-S","N3-S","N1-C","S2-S","S2-C","S3-S","S1-C"), pos=c(1,3,1,3), offset=1)
dev.off()

### technically, I have also plotted the experiment locations as accessions.  Fix this if using this plot for something real.


```

```{R PCA distribution in sweden data}
### look at distribution of BIO12 and BIO4
library(ggplot2)
library(ggmap)
library(maps)
library(mapdata)
library("sf")
library("rnaturalearth")
library("rnaturalearthdata")
library("rgeos")

theme_set(theme_bw())

world <- ne_countries(scale = "medium", returnclass = "sf")
class(world)

#sweden <- map_data("sweden")
sweden <- map('worldHires','Sweden')
sweden.map <- ggplot(data = world) +geom_sf() + coord_sf(xlim = c(sweden$range[1], sweden$range[2]), ylim = c(sweden$range[3], sweden$range[4]), expand = FALSE)
bio12.map <- sweden.map + geom_point(data=c.dat, mapping = aes(x = lon, y = lat, colour = BIO12))
bio4.map <- sweden.map + geom_point(data=c.dat, mapping = aes(x = lon, y = lat, colour = BIO4))
bio12.map
bio4.map
pdf("./figures/bioclim.maps.pdf")
print(bio12.map)
print(bio4.map)
dev.off()
```




## 3. PCA analysis of accession data within 1001g data

```{r 1001g climate space PCA}


library(corrplot)
corrplot(cor(all.dat[,4:24]), method="color", type="upper", diag=FALSE)

pdf("./figures/PCA.bioclimv2.1001g.pdf", width=10, height=10)
corrplot(cor(c.dat[,4:24]), method="ellipse", type="upper", diag=FALSE)
dev.off()


## clearly a LOT of correlation here.  Good data for PCA
## use 19 climate variables

up.dat <- all.dat[,6:24]
rownames(up.dat) <- rownames(all.dat)
all.pca <- prcomp(up.dat, center=TRUE)
all.pca
summary(all.pca)

library(ggfortify)

autoplot(prcomp(up.dat, center=TRUE), data = all.dat, loadings = TRUE, colour="region",size="type")

### I like this analysis using only the 19 climate variables
### let's look if locally adaptive

all.pcs <- as.data.frame(all.pca$x)
all.pcs$id <- rownames(all.pcs)
```



```{R all PCA plot by hand for annotation}
pdf("./figures/PCA.1001g.19climvars.bioclimv2.pdf", width=8, height=6)
plot(all.pcs$PC1, all.pcs$PC2, xlab="PC1 (67.11%)", ylab="PC2 (31.43%)", col="grey90", pch=19)
n.accs <- all.pcs[rownames(all.pcs)%in%rownames(c.dat[c.dat$region=="N Sweden",]),]
s.accs <- all.pcs[rownames(all.pcs)%in%rownames(c.dat[c.dat$region=="S Sweden",]),]
c.accs <- all.pcs[rownames(all.pcs)%in%rownames(c.dat[c.dat$region=="C Sweden",]),]
points(n.accs$PC1, n.accs$PC2, pch=19, col="dodgerblue")
points(c.accs$PC1, c.accs$PC2, pch=19, col="purple")
points(s.accs$PC1, s.accs$PC2, pch=19, col="green")
exp.sites <- all.pcs[rownames(all.pcs)%in%rownames(c.dat[c.dat$type%in%c("common_garden", "snr"),]),]
points(exp.sites$PC1, exp.sites$PC2, pch=23, cex=2.5, lwd=2)
legend("bottomleft", col=c("green", "purple","blue","black"),pch=c(19,19,19,23), legend=c("Southern lines", "Central lines", "Northern lines", "Experiment sites"))
text(exp.sites[,1], exp.sites[,2], labels=c("N2-C","N2-S","N3-S","N1-C","S2-S","S2-C","S3-S","S1-C"), pos=c(1,3,1,3), offset=1)
dev.off()

### technically, I have also plotted the experiment locations as accessions.  Fix this if using this plot for something real.


```

### repeat these plots with the K groups from 29.BLUP.heatmap.by.K.Rmd

```{r pca plots with kgroups}
## load K groups
kcl <- read.table("./data/K.groups.txt")
colnames(kcl) <- c("K.group")
c.dat$name <- rownames(c.dat)
c.dat <- merge(c.dat,kcl, by.x="name", by.y="row.names", all.x=TRUE)
rownames(c.dat) <- c.dat$name
ann.colors <- c("#A6CEE3","#1F78B4","#B2DF8A", "#33A02C", "#FB9A99", "#E31A1C")


pdf("./figures/PCA.accessions.19climvars.bioclimv2.K.groups.pdf", width=8, height=6)
plot(pcs$PC1, pcs$PC2, xlab="PC1 (67.11%)", ylab="PC2 (31.43%)", type="n")

k1 <- pcs[rownames(pcs)%in%rownames(c.dat[c.dat$K.group==1,]),]
points(k1$PC1, k1$PC2, pch=19, col=ann.colors[1])
k2 <- pcs[rownames(pcs)%in%rownames(c.dat[c.dat$K.group==2,]),]
points(k2$PC1, k2$PC2, pch=19, col=ann.colors[2])
k3 <- pcs[rownames(pcs)%in%rownames(c.dat[c.dat$K.group==3,]),]
points(k3$PC1, k3$PC2, pch=19, col=ann.colors[3])
k4 <- pcs[rownames(pcs)%in%rownames(c.dat[c.dat$K.group==4,]),]
points(k4$PC1, k4$PC2, pch=19, col=ann.colors[4])
k5 <- pcs[rownames(pcs)%in%rownames(c.dat[c.dat$K.group==5,]),]
points(k5$PC1, k5$PC2, pch=19, col=ann.colors[5])
k6 <- pcs[rownames(pcs)%in%rownames(c.dat[c.dat$K.group==6,]),]
points(k6$PC1, k6$PC2, pch=19, col=ann.colors[6])

exp.sites <- pcs[rownames(pcs)%in%rownames(c.dat[c.dat$type%in%c("common_garden", "snr"),]),]
points(exp.sites$PC1, exp.sites$PC2, pch=23, cex=2.5, lwd=2)
legend("bottomleft", col=c(ann.colors,"black"),pch=c(rep(19,6),23), legend=c("K1", "K2", "K3","K4","K5","K6", "Experiments"))
text(exp.sites[,1], exp.sites[,2], labels=c("N2-C","N2-S","N3-S","N1-C","S2-S","S2-C","S3-S","S1-C"), pos=c(1,3,1,3), offset=1)
dev.off()

```

### repeat these plots yet again with the K groups from 29.BLUP.heatmap.by.K.Rmd and the most up to date nomenclature 04Aug20

```{r pca plots with kgroups new annotation}
## get new K group names

kcol <- brewer.pal(6, "Paired")[c(6,1,2,5,3,4)]
ki <- data.frame(matrix(c("1","S1","2","S2","3","C","4","N1","5","N2","6","B"),ncol=2,byrow=TRUE))
colnames(ki) <- c("K.group","K.name")
c.dat <- merge(c.dat, ki, by="K.group", all.x=TRUE)
c.dat$K.name <- factor(c.dat$K.name, levels=c("B","S1","S2","C","N1","N2"))
rownames(c.dat) <- c.dat$name

pdf("./figures/PCA.accessions.19climvars.bioclimv2.K.groups.annotationAug2020.pdf", width=8, height=6)
plot(pcs$PC1, pcs$PC2, xlab="PC1 (67.11%)", ylab="PC2 (31.43%)", type="n")

for(up in 1:6){
  up.f <- levels(c.dat$K.name)[up]
  print(up.f)
  up.dat <- pcs[rownames(pcs)%in%rownames(c.dat[c.dat$K.name==up.f,]),]
  points(up.dat$PC1, up.dat$PC2, pch=19, col=kcol[up])
}

exp.sites <- pcs[rownames(pcs)%in%rownames(c.dat[c.dat$type%in%c("common_garden"),]),]
points(exp.sites$PC1, exp.sites$PC2, pch=23, cex=2.5, lwd=2)
legend("bottomleft", col=c(kcol,"black"),pch=c(rep(19,6),23), legend=c(levels(c.dat$K.name), "Experiments"), title="K groups")
text(exp.sites[,1], exp.sites[,2], labels=c("NA","NM","SR","SU"), pos=c(1,3,1,3), offset=1)
dev.off()

```



