---
title: "19.fecundity.by.admixture"
author: "Daniele Filiault"
date: "9/19/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(RColorBrewer)
library(reshape)
library(ggplot2)
```

## Introduction

running script 16 with survival to seed set to look for relationship between proportion of genome and overwinter survival

```{r prep admixture data}
### admixture groups for 1001g data - from the paper:
#We defined nine groups based on these clusters and assigned each individual to a group if more than 60% of its genome
#derived from the corresponding cluster. The 135 individuals not
#matching this criterion were labeled ‘‘Admixed.’’
admix <- read.csv(file="./data/1001genomes-accessionsand1001genomes-admixture.csv", stringsAsFactors=FALSE)
admix.m <- admix[,grep("K9", colnames(admix))]
rownames(admix.m) <- admix$id
#admix.m$id <- rownames(admix.m)

max.a <- apply(admix.m,1, function(x)(max(x, na.rm=TRUE)))
a.group <- rep(NA, nrow(admix.m))
for(up in 1:nrow(admix.m)){
  up.dat <- admix.m[up,]
  a.group[up] <- which(up.dat==max.a[up])-1
}
names(a.group)<- rownames(admix.m)
a.group <- cbind(admix[,c(1:4)], max.a,a.group, admix.m)


## subset for experiment accessions.  Some will be missing.  Should rerun admixture with expanded data?
fit.dat <- read.table("./data/fecund.fitted.values.txt", stringsAsFactors=FALSE)
fit.dat <- fit.dat[,c(1,8:11,14:17)]

a.fit <- merge(a.group,fit.dat)  ### only 186 out of 200 accessions have admixture info

## get info about N/C/S accession
acc.list <- read.table("./data/acc_list.txt", sep="\t", header=TRUE)
colnames(acc.list)[1] <- "id"
a.fit <- merge(a.fit, acc.list[,c(1,3)])
a.fit$region <- droplevels(a.fit$region)

group.names <- c("W.Europe","Relicts","Germany","N.Sweden","Iberia","Asia","S.Sweden","Italy/E.Europe","C.Europe")

```



### Effect of Proportion of admixture group
```{R fitness by admixture proportion}
fit.columns <- colnames(a.fit)[16:23]

plot.fit.prop <- function(fit.col){
  par(mfcol=c(2,2))
  with(a.fit, plot(K9.ChrAll.3, get(fit.col), type="n", xlab="proportion genome N.Swedish", ylab=paste("fecundity",fit.col, sep=" ")))
  with(a.fit[a.fit$region=="N Sweden",], points(K9.ChrAll.3, get(fit.col), col="cadetblue", pch=19))
  with(a.fit[a.fit$region=="S Sweden",], points(K9.ChrAll.3, get(fit.col), col="green", pch=19))
  with(a.fit[a.fit$region=="C Sweden",], points(K9.ChrAll.3, get(fit.col), col="purple", pch=19))
  abline(lm(get(fit.col)~K9.ChrAll.3, data=a.fit))
  abline(lm(get(fit.col)~K9.ChrAll.3, data=a.fit[a.fit$region=="C Sweden",]), col="purple", lty=4)

  with(a.fit, plot(K9.ChrAll.6, get(fit.col), type="n", xlab="proportion of genome S.Swedish", ylab=paste("fecundity",fit.col, sep=" ")))
  with(a.fit[a.fit$region=="S Sweden",], points(K9.ChrAll.6, get(fit.col), col="green", pch=19))
  with(a.fit[a.fit$region=="N Sweden",], points(K9.ChrAll.6, get(fit.col), col="cadetblue", pch=19))
  with(a.fit[a.fit$region=="C Sweden",], points(K9.ChrAll.6, get(fit.col), col="purple", pch=19))
  abline(lm(get(fit.col)~K9.ChrAll.6, data=a.fit))
  abline(lm(get(fit.col)~K9.ChrAll.6, data=a.fit[a.fit$region=="C Sweden",]), col="purple", lty=4)

  with(a.fit, plot(K9.ChrAll.2, get(fit.col),type="n", xlab="proportion of genome German", ylab=paste("fecundity",fit.col, sep=" ")))
  with(a.fit[a.fit$region=="S Sweden",], points(K9.ChrAll.2, get(fit.col), col="green", pch=19))
  with(a.fit[a.fit$region=="N Sweden",], points(K9.ChrAll.2, get(fit.col), col="cadetblue", pch=19))
  with(a.fit[a.fit$region=="C Sweden",], points(K9.ChrAll.2, get(fit.col), col="purple", pch=19))
  abline(lm(get(fit.col)~K9.ChrAll.2, data=a.fit))
  abline(lm(get(fit.col)~K9.ChrAll.2, data=a.fit[a.fit$region=="C Sweden",]), col="purple", lty=4)

  with(a.fit, plot(1-(K9.ChrAll.3+K9.ChrAll.6), get(fit.col), type="n", xlab="proportion of genome other groups", ylab=paste("fecundity",fit.col, sep=" ")))
  with(a.fit[a.fit$region=="N Sweden",], points(1-(K9.ChrAll.3+K9.ChrAll.6+K9.ChrAll.2), get(fit.col), col="cadetblue", pch=19))
  with(a.fit[a.fit$region=="S Sweden",], points(1-(K9.ChrAll.3+K9.ChrAll.6+K9.ChrAll.2), get(fit.col), col="green", pch=19))
  with(a.fit[a.fit$region=="C Sweden",], points(1-(K9.ChrAll.3+K9.ChrAll.6+K9.ChrAll.2), get(fit.col), col="purple", pch=19))
  p.nsw <- 1-(a.fit$K9.ChrAll.3+a.fit$K9.ChrAll.6+a.fit$K9.ChrAll.2)
  abline(lm(get(fit.col)~p.nsw, data=a.fit))
  p.nsw.c <- 1-(a.fit[a.fit$region=="C Sweden",]$K9.ChrAll.3+a.fit[a.fit$region=="C Sweden",]$K9.ChrAll.6+a.fit[a.fit$region=="C Sweden",]$K9.ChrAll.2)
  abline(lm(get(fit.col)~p.nsw.c, data=a.fit[a.fit$region=="C Sweden",]), col="purple", lty=4)
  legend(x="topright", legend=c("North","Central", "South"), col=c("cadetblue","purple","green"), pch=19)
}

for(up.c in fit.columns){
  pdf(file=paste("./figures/fecundity.by.admix.prop.", up.c, ".pdf", sep=""), width=8, height=6)
  par(mar=c(5,4,1,1))
  plot.fit.prop(fit.col=up.c)
  dev.off()
}

```

What's most important in determining fitness, prop N, prop S, or prop non-Swedish?

These proportions are weird variables (high correlation), and there is probably a better way to do this, but let's just start with plain old linear models like in the above figures

```{R lm fit by prop}

prop.lm <- function(fit.col){
  out.dat <- as.list(1:4)
  out.dat[[1]] <- lm(get(fit.col)~K9.ChrAll.3, data=a.fit)
  out.dat[[2]] <- lm(get(fit.col)~K9.ChrAll.6, data=a.fit)
  out.dat[[3]] <- lm(get(fit.col)~K9.ChrAll.2, data=a.fit)
  p.nsw <- 1-(a.fit$K9.ChrAll.3+a.fit$K9.ChrAll.6+a.fit$K9.ChrAll.2)
  out.dat[[4]] <- lm(get(fit.col)~p.nsw, data=a.fit)
  return(out.dat)
}

### 2011 first
fit.columns <- colnames(a.fit)[grep("2011",colnames(a.fit))]
ada1 <- prop.lm(fit.col=fit.columns[1])
ram1 <- prop.lm(fit.col=fit.columns[2])
ull1 <- prop.lm(fit.col=fit.columns[3])
rat1 <- prop.lm(fit.col=fit.columns[4])

### 2012 second
fit.columns <- colnames(a.fit)[grep("2012",colnames(a.fit))]
ada2 <- prop.lm(fit.col=fit.columns[1])
ram2 <- prop.lm(fit.col=fit.columns[2])
ull2 <- prop.lm(fit.col=fit.columns[3])
rat2 <- prop.lm(fit.col=fit.columns[4])

###### slopes ######
### 2011
ada1.s <- unlist(lapply(ada1, function(x){summary(x)$coefficients[2,1]}))
ram1.s <- unlist(lapply(ram1, function(x){summary(x)$coefficients[2,1]}))
ull1.s <- unlist(lapply(ull1, function(x){summary(x)$coefficients[2,1]}))
rat1.s <- unlist(lapply(rat1, function(x){summary(x)$coefficients[2,1]}))
slope1 <- rbind(ada1.s, ram1.s, ull1.s, rat1.s)
colnames(slope1) <- c("n.swedish", "s.swedish", "german", "other")

### 2012
ada2.s <- unlist(lapply(ada2, function(x){summary(x)$coefficients[2,1]}))
ram2.s <- unlist(lapply(ram2, function(x){summary(x)$coefficients[2,1]}))
ull2.s <- unlist(lapply(ull2, function(x){summary(x)$coefficients[2,1]}))
rat2.s <- unlist(lapply(rat2, function(x){summary(x)$coefficients[2,1]}))
slope2 <- rbind(ada2.s, ram2.s, ull2.s, rat2.s)
colnames(slope2) <- c("n.swedish", "s.swedish", "german", "other")

######### pvals #############
### 2011
ada1.p <- unlist(lapply(ada1, function(x){summary(x)$coefficients[2,4]}))
ram1.p <- unlist(lapply(ram1, function(x){summary(x)$coefficients[2,4]}))
ull1.p <- unlist(lapply(ull1, function(x){summary(x)$coefficients[2,4]}))
rat1.p <- unlist(lapply(rat1, function(x){summary(x)$coefficients[2,4]}))
pval1 <- rbind(ada1.p, ram1.p, ull1.p, rat1.p)
colnames(pval1) <- c("n.swedish", "s.swedish", "german", "other")

### 2012
ada2.p <- unlist(lapply(ada2, function(x){summary(x)$coefficients[2,4]}))
ram2.p <- unlist(lapply(ram2, function(x){summary(x)$coefficients[2,4]}))
ull2.p <- unlist(lapply(ull2, function(x){summary(x)$coefficients[2,4]}))
rat2.p <- unlist(lapply(rat2, function(x){summary(x)$coefficients[2,4]}))
pval2 <- rbind(ada2.p, ram2.p, ull2.p, rat2.p)
colnames(pval2) <- c("n.swedish", "s.swedish", "german", "other")



```


```{R plot slopes}
### plot these slopes
ad.colors <- brewer.pal(9, "Paired")[c(4,7,3,6)]  ### numbers are one higher than admixture group because groups start at 0!
ad.names <- c("N.Swedish", "S.Swedish", "German", "Other")
p.pch <-c(17, 19)

plot.slope <- function(up.s, up.p, up.t){
  up.s <- up.s[4:1,]
  up.p <- up.p[4:1,]
  par(mar=c(4,4,2,2))
  plot(0,0, xlim=c(min(up.s), max(up.s)), ylim=c(0.5,nrow(up.s)), yaxt="n", ylab="", xlab="slope fecundity~admix proportion", main=up.t)
  abline(v=0, col="grey70")
  for(up.r in c(1:4)){
    up.dat <- up.s[up.r,]
    up.pv <- up.p[up.r,]
    up.pv <- sapply(up.pv, function(x){x<0.05})
    up.pv <- as.numeric(up.pv)+1
    up.pv <- p.pch[up.pv]
    points(x=up.dat, y=rep(up.r,length(up.dat)), col=ad.colors, pch=up.pv, cex=1.7)
  }
  axis(2, at=1:4, labels=substr(rownames(up.s),1,3), las=2)
  legend(x="bottom",legend=c(ad.names, "sig", "not sig"), col=c(ad.colors, "black", "black"), pch=c(rep(19,4),19,17), ncol=3, bg="white")
}

par(mfcol=c(1,2))
plot.slope(up.s=slope1, up.p=pval1, up.t="2011")
plot.slope(up.s=slope2, up.p=pval2, up.t="2012")

pdf("./figures/slope.fecundity.admixture.pdf", width=10, height=6)
par(mfcol=c(1,2))
plot.slope(up.s=slope1, up.p=pval1, up.t="2011")
plot.slope(up.s=slope2, up.p=pval2, up.t="2012")
dev.off()
```


```{R fitness rank}

#Is a pattern in proportion when accessions are ranked by beta?

pdf(file="./figures/fecundity.rank.by.admix.prop.pdf", width=18, height=8)
par(mfcol=c(4,3), mar=c(0.2,4,1.5,0.2))
for(up.c in 16:23){
  p.dat <- a.fit[order(a.fit[,up.c]),]
  barplot(t(as.matrix(p.dat[,7:15])), col=brewer.pal(9, "Paired"), xaxt="n",space=0, ylab="genome proportion")
  mtext(colnames(a.fit)[up.c],side=3, line=0)
}
plot(0,type='n',axes=FALSE,ann=FALSE)
legend(legend=group.names, fill=brewer.pal(9, "Paired"), title="admixture group",bg="white",x="left")
#barplot(0, plot=FALSE, axes=FALSE, ann=FALSE,col=brewer.pal(9, "Paired"), legend.text=group.names, xaxt="n",space=0, args.legend=list(title="admixture group",bg="white",x="left") )

dev.off()
```
```{R admixture group fecundity diffs}
ga.fit <- a.fit[a.fit$max.a>=0.60,]
ga.fit <- ga.fit[ga.fit$a.group%in%c(2,3,6),]
table(ga.fit$a.group)


#### 2011 first
ag1.aov <- as.list(1:4)
ag1.tukey <- as.list(1:4
                     )
col1 <- colnames(ga.fit)[grep("2011",colnames(ga.fit))]

for(up in 1:4){
  aov.up <- aov(get(col1[up])~as.factor(a.group), data=ga.fit)
  tt.up <- TukeyHSD(aov.up)
  ag1.aov[[up]] <- aov.up
  ag1.tukey[[up]] <- tt.up
}

par(mfrow=c(2,2))
lapply(ag1.tukey, plot)

lapply(ag1.aov,summary)


#### 2012 second
ag2.aov <- as.list(1:4)
ag2.tukey <- as.list(1:4
                     )
col2 <- colnames(ga.fit)[grep("2012",colnames(ga.fit))]

for(up in 1:4){
  aov.up <- aov(get(col2[up])~as.factor(a.group), data=ga.fit)
  tt.up <- TukeyHSD(aov.up)
  ag2.aov[[up]] <- aov.up
  ag2.tukey[[up]] <- tt.up
}

par(mfrow=c(2,2))
lapply(ag2.tukey, plot)

lapply(ag2.aov,summary)

```


```{R plot by means rxn norm}
##reshape data
ga.long <- ga.fit[,c(1,6,grep("2011", colnames(ga.fit)), grep("2012", colnames(ga.fit)))]
mdata <- melt(ga.long, id=c("id","a.group"))
var.s <- strsplit(as.character(mdata$variable), split="_")
var.s <- do.call(rbind, var.s)
colnames(var.s) <- c("site", "year")
ga.long <- cbind(mdata, var.s)
colnames(ga.long)[4] <- "fecundity"
ga.long$site = factor(ga.long$site,levels(ga.long$site)[c(1,2,4,3)])


ag.colors <- brewer.pal(9, "Paired")[c(3,4,7)]

ad.group.boxplot <- ggplot(ga.long, aes(x=as.factor(a.group), y=fecundity, fill=as.factor(a.group))) +
  geom_violin() +
  geom_boxplot(width=0.1, show.legend=FALSE)+
  xlab("admixture group")+ 
  facet_grid(year~site) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_manual(name = "admixture group", labels = c("Germany", "N.Sweden", "S.Sweden"), values=ag.colors) +
  scale_x_discrete(labels=c("Germany", "N.Sweden", "S.Sweden"))

print(ad.group.boxplot)

pdf(file="./figures/fecundity.by.admix.group.pdf", width=8, height=6)
print(ad.group.boxplot)
dev.off()

```