---
title: "22.compare.betas.across.exp.by.type"
author: "Daniele Filiault"
date: "11/4/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

The goal here is to compare the betas of the most common allele in each haplotype group at each site/year combination.  Questions:

1.  common versus opposing effects between admixture groups that "match" with local adaptation and/or general fitness.  This can be local adaptation to N/S or to Sweden in general.
2.  Patterns when same or different allele is most common in N and S admixture group?
3.  Is there any difference by the evolutionary history type (ancestral, derived, standing)?

## Data prep
```{r prep betas}
## individual betas
## summary data from 07.geo.betas.Rmd - betas of alt alleles
load("./data/geo.beta.2011.2012.Rdat")
betas <- geo.beta.2011.2012[,c(1:7,23:26)]
## composite betas
## output from 06.modeled.fitness.GWAS.Rmd
load("./data/composite.betas.Rdat")
colnames(composite.betas)[1:3] <- c("Chromosome", "rs", "Position")

betas <- merge(betas, composite.betas)
colnames(betas) <- gsub("beta_1", "N2",colnames(betas))
colnames(betas) <- gsub("beta_2", "N1",colnames(betas))
colnames(betas) <- gsub("beta_3", "S1",colnames(betas))
colnames(betas) <- gsub("beta_4", "S2",colnames(betas))
```

```{r prep admix data}
### done in script ./scripts/admix.count.all.snps.R
### ran this on the cluster with ./scripts/submit.admix.count.all.snps.sh
### output is admixture group of reference and alt alleles for the 1001g data
### so a few of our experimental lines are missing from this.  I think that's OK, but I will need to think about it more...
### this gives the number of acessions determined to be in each admixture group that have the reference(R) and alt(A) allele.

load("./data/admix.count.Rdat")

### subset to SNPs that were used for GWAS
admix.count <- admix.count[row.names(admix.count)%in%betas$rs,]  #735745

### remove non-polymorphic (1 or 2 calls)
a.sums <- apply(admix.count, 1, sum)
admix.prop.pol <- admix.count[a.sums>1,]

## which admixture group is which (from 0-8)
group.names <- c("W.Europe","Relicts","Germany","N.Sweden","Iberia","Asia","S.Sweden","Italy/E.Europe","C.Europe")
```

```{r prep histories}
## allele histories
### ancestral and derived alleles were determined in /projects/field_experiments/001.common.reference.files/004.genus.vcfs

p.snps <- read.table("./data/Polarized.snps.txt",colClasses="character")
rownames(p.snps) <- gsub("Chr", "", rownames(p.snps))
p.snps$rs <- rownames(p.snps)
p.snps <- p.snps[p.snps$rs%in%betas$rs,]

### 000 is cases without calls in either ref or alt species 71950
### 100 is reference ancestral  271785 (ie alt is derived allele)
### 010 is alt ancestral  60617 
### 001 is both ref and alt alleles occur in species 155840
```

## get most frequent alleles in N and S admixture groups
```{R common allele in admix group}

pos.line <- admix.prop.pol[100,]
ad.group <- "3"

com.ad.group <- function(pos.line,ad.group){
  up.dat <- pos.line[grep(ad.group,names(pos.line))]
  if(up.dat[1]==up.dat[2]){
    #print(up.dat)
    out.dat <- c(2,0.5)
    }else{
      max.allele <- names(up.dat)[which(up.dat==max(up.dat))]
      if(substr(max.allele,1,1)=="A"){bi.allele <- 1}
      if(substr(max.allele,1,1)=="R"){bi.allele <- -1}
      max.prop <- up.dat[which(up.dat==max(up.dat))]/sum(up.dat)
      out.dat <- c(bi.allele, max.prop)
    }
  return(out.dat)
}

ger.allele <- apply(admix.prop.pol, 1, function(x){com.ad.group(pos.line=x, ad.group="2")})
ger.allele <- t(ger.allele)
colnames(ger.allele) <- c("mc.allele","prop")

ss.allele <- apply(admix.prop.pol, 1, function(x){com.ad.group(pos.line=x, ad.group="6")})
ss.allele <- t(ss.allele)
colnames(ss.allele) <- c("mc.allele","prop")

ns.allele <- apply(admix.prop.pol, 1, function(x){com.ad.group(pos.line=x, ad.group="3")})
ns.allele <- t(ns.allele)
colnames(ns.allele) <- c("mc.allele","prop")
```


## change sign of betas appropriately
```{r change beta signs}
##remove SNPs that are 50% in both
##could also randomly assign?
##but very few SNPs, comparatively speaking.
ga.common <- ger.allele[ger.allele[,1]%in%c(-1,1),] #99.4 percent still in
betas.common <- betas[as.character(betas$rs)%in%rownames(ga.common),]
ga.common <- ga.common[order(rownames(ga.common)),]
#rownames(ga.common)==betas.common$rs -> check.me
#summary(check.me)
ger.betas <- apply(betas.common[,4:ncol(betas)],2,function(x){x*ga.common[,1]})
ger.betas <- cbind(ger.betas, betas.common[,1:3])

na.common <- ns.allele[ns.allele[,1]%in%c(-1,1),] #99.4 percent still in
betas.common <- betas[as.character(betas$rs)%in%rownames(na.common),]
na.common <- na.common[order(rownames(na.common)),]
#rownames(na.common)==betas.common$rs -> check.me
#summary(check.me)
ns.betas <- apply(betas.common[,4:ncol(betas)],2,function(x){x*na.common[,1]})
ns.betas <- cbind(ns.betas, betas.common[,1:3])

sa.common <- ss.allele[ss.allele[,1]%in%c(-1,1),] #99.4 percent still in
betas.common <- betas[as.character(betas$rs)%in%rownames(sa.common),]
sa.common <- sa.common[order(rownames(sa.common)),]
#rownames(sa.common)==betas.common$rs -> check.me
#summary(check.me)
ss.betas <- apply(betas.common[,4:ncol(betas)],2,function(x){x*sa.common[,1]})
ss.betas <- cbind(ss.betas, betas.common[,1:3])
```
## plot betas by site
```{r betas by site NS, eval=FALSE}
ssb.c <- ss.betas[ss.betas$rs%in%ns.betas$rs,]
nsb.c <- ns.betas[ns.betas$rs%in%ssb.c$rs,]
#nsb.c$rs==ssb.c$rs -> blah
#table(blah)
# now these can be easily plotted together

pdf(file="./figures/most.commmon.beta.NvsS.pdf", width=16, height=12)
par(mfrow=c(3,4))
for(up.beta in 1:12){
  par(pty="s")
  max.val <- abs(max(c(ssb.c[,up.beta],nsb.c[,up.beta])))
  smoothScatter(ssb.c[,up.beta],nsb.c[,up.beta],xlab="effect of most common allele S.Swedish admix group", ylab="effect of most common allele N.Swedish admix group", main=colnames(ssb.c[up.beta]),xlim=c(-max.val, max.val), ylim=c(-max.val, max.val))
  abline(h=0,lty=3)
  abline(v=0,lty=3)
  
  ##add numbers per quadrant
  same <- ssb.c[,up.beta]==nsb.c[,up.beta]
  ss.pos <- ssb.c[, up.beta]>0
  quad.nos <- table(paste(same, ss.pos))
  text(-max.val, max.val, quad.nos[1], pos=4) #not same, S negative
  text(max.val, -max.val, quad.nos[2], pos=2) #not same, S positive
  text(-max.val, -max.val, quad.nos[3], pos=4) #same, S negative
  text(max.val, max.val, quad.nos[4], pos=2) #same, S positive
}
dev.off()
```

```{r betas by site NG, eval=FALSE}
gb.c <- ger.betas[ger.betas$rs%in%ns.betas$rs,]
nsb.c <- ns.betas[ns.betas$rs%in%gb.c$rs,]
#nsb.c$rs==ssb.c$rs -> blah
#table(blah)
# now these can be easily plotted together

pdf(file="./figures/most.commmon.beta.NvsG.pdf", width=16, height=12)
par(mfrow=c(3,4))
for(up.beta in 1:12){
  par(pty="s")
  max.val <- abs(max(c(gb.c[,up.beta],nsb.c[,up.beta])))
  smoothScatter(gb.c[,up.beta],nsb.c[,up.beta],xlab="effect of most common allele German admix group", ylab="effect of most common allele N.Swedish admix group", main=colnames(gb.c[up.beta]),xlim=c(-max.val, max.val), ylim=c(-max.val, max.val))
  abline(h=0,lty=3)
  abline(v=0,lty=3)
  
  ##add numbers per quadrant
  same <- gb.c[,up.beta]==nsb.c[,up.beta]
  g.pos <-gb.c[, up.beta]>0
  quad.nos <- table(paste(same, g.pos))
  text(-max.val, max.val, quad.nos[1], pos=4) #not same, G negative
  text(max.val, -max.val, quad.nos[2], pos=2) #not same, G positive
  text(-max.val, -max.val, quad.nos[3], pos=4) #same, G negative
  text(max.val, max.val, quad.nos[4], pos=2) #same, G positive
}
dev.off()
```

```{r betas by site SG, eval=FALSE}
ssb.c <- ss.betas[ss.betas$rs%in%ger.betas$rs,]
gb.c <- ger.betas[ger.betas$rs%in%ssb.c$rs,]
#gb.c$rs==ssb.c$rs -> blah
#table(blah)
# now these can be easily plotted together

pdf(file="./figures/most.commmon.beta.SvsG.pdf", width=16, height=12)
par(mfrow=c(3,4))
for(up.beta in 1:12){
  par(pty="s")
  max.val <- abs(max(c(ssb.c[,up.beta],gb.c[,up.beta])))
  smoothScatter(ssb.c[,up.beta],gb.c[,up.beta],xlab="effect of most common allele S.Swedish admix group", ylab="effect of most common allele Ger admix group", main=colnames(ssb.c[up.beta]),xlim=c(-max.val, max.val), ylim=c(-max.val, max.val))
  abline(h=0,lty=3)
  abline(v=0,lty=3)
  
  ##add numbers per quadrant
  same <- ssb.c[,up.beta]==gb.c[,up.beta]
  ss.pos <- ssb.c[, up.beta]>0
  quad.nos <- table(paste(same, ss.pos))
  text(-max.val, max.val, quad.nos[1], pos=4) #not same, S negative
  text(max.val, -max.val, quad.nos[2], pos=2) #not same, S positive
  text(-max.val, -max.val, quad.nos[3], pos=4) #same, S negative
  text(max.val, max.val, quad.nos[4], pos=2) #same, S positive
}
dev.off()
```



## What are evol history of alleles??

```{r hist most common allele by admix group}
## Northern Swedish admixture group
ns.hist <- merge(ns.allele, p.snps, by="row.names")
ns.hist <- ns.hist[ns.hist$mc.allele%in%c(-1,1),]
ns.hist.s <- split(ns.hist, ns.hist$x)
ns.hist.s[[1]]$a.hist <- 0 #can't tell code 000
ns.hist.s[[2]]$a.hist <- 1 #standing code 001
aa <- split(ns.hist.s[[3]], ns.hist.s[[3]]$mc.allele) # alt ancestral code 010
aa[[1]]$a.hist <- 3 #alt ancestral, most common is ref, most common derived
aa[[2]]$a.hist <- 2 #alt ancestral, most common is alt, most common ancestral
ra <- split(ns.hist.s[[4]], ns.hist.s[[4]]$mc.allele) # ref ancestral code 100
ra[[1]]$a.hist <- 2 #ref ancestral, most common is ref, most common ancestral
ra[[2]]$a.hist <- 3 #ref ancestral, most common is alt, most common derived
ns.hist <- rbind(ns.hist.s[[1]], ns.hist.s[[2]], aa[[1]], aa[[2]], ra[[1]], ra[[2]])
ns.hist[,1] <- as.character(ns.hist[,1])
ns.hist <- ns.hist[order(ns.hist[,1]),]

##a.hist:
##0 = can't tell
##1 = standing
##2 = most common ancestral
##3 = most common derived
a.hist.levels <- c("unknown","standing", "ancestral", "derived")

## Southern Swedish admixture group
ss.hist <- merge(ss.allele, p.snps, by="row.names")
ss.hist <- ss.hist[ss.hist$mc.allele%in%c(-1,1),]
ss.hist.s <- split(ss.hist, ss.hist$x)
ss.hist.s[[1]]$a.hist <- 0 #can't tell code 000
ss.hist.s[[2]]$a.hist <- 1 #standing code 001
aa <- split(ss.hist.s[[3]], ss.hist.s[[3]]$mc.allele) # alt ancestral code 010
aa[[1]]$a.hist <- 3 #alt ancestral, most common is ref, most common derived
aa[[2]]$a.hist <- 2 #alt ancestral, most common is alt, most common ancestral
ra <- split(ss.hist.s[[4]], ss.hist.s[[4]]$mc.allele) # ref ancestral code 100
ra[[1]]$a.hist <- 2 #ref ancestral, most common is ref, most common ancestral
ra[[2]]$a.hist <- 3 #ref ancestral, most common is alt, most common derived
ss.hist <- rbind(ss.hist.s[[1]], ss.hist.s[[2]], aa[[1]], aa[[2]], ra[[1]], ra[[2]])
ss.hist[,1] <- as.character(ss.hist[,1])
ss.hist <- ss.hist[order(ss.hist[,1]),]

## german admixture group
ger.hist <- merge(ger.allele, p.snps, by="row.names")
ger.hist <- ger.hist[ger.hist$mc.allele%in%c(-1,1),]
ger.hist.s <- split(ger.hist, ger.hist$x)
ger.hist.s[[1]]$a.hist <- 0 #can't tell code 000
ger.hist.s[[2]]$a.hist <- 1 #standing code 001
aa <- split(ger.hist.s[[3]], ger.hist.s[[3]]$mc.allele) # alt ancestral code 010
aa[[1]]$a.hist <- 3 #alt ancestral, most common is ref, most common derived
aa[[2]]$a.hist <- 2 #alt ancestral, most common is alt, most common ancestral
ra <- split(ger.hist.s[[4]], ger.hist.s[[4]]$mc.allele) # ref ancestral code 100
ra[[1]]$a.hist <- 2 #ref ancestral, most common is ref, most common ancestral
ra[[2]]$a.hist <- 3 #ref ancestral, most common is alt, most common derived
ger.hist <- rbind(ger.hist.s[[1]], ger.hist.s[[2]], aa[[1]], aa[[2]], ra[[1]], ra[[2]])
ger.hist[,1] <- as.character(ger.hist[,1])
ger.hist <- ss.hist[order(ger.hist[,1]),]

```

## difference in betas correlation patterns by evolutionary histories?

```{r N.Swedish admix group correlations between plot}
ns.dat <- merge(ns.betas, ns.hist, by="rs", all=TRUE)
par(pty="s")
max.val <- abs(max(c(ns.dat$S1_comp,ns.dat$S2_comp)))
smoothScatter(ns.dat$S1_comp, ns.dat$S2_comp, ylim=c(-max.val, max.val), xlim=c(-max.val, max.val))
abline(h=0,lty=3)
abline(v=0,lty=3)
```

So I think that this is going a bit far down the rabbit hole.  Let's stick with the "X plots" and do evolutionary history in
In the plots of weighted fitness.

## permutation test for X plots

So the trends in X plots suggest local adaptation, but are these numbers significant?
Test this by permutation.
1. Keep common/opposing most common allele structure.
2. Rotate the genomes (i.e. the betas)
3. Repolarize betas using (1).
4. Calculate # in each quadrant
5. Return
6. do 1000 or 10000x.

```{r permutation function}
al1 <- ns.allele
al2 <- ss.allele
betas <- betas
n.perm <- 10

permute.quadrants <- function(al1, al2, betas, n.perm){
## identify common positions
  pos <- table(c(as.character(rownames(al1)),as.character(rownames(al2)), as.character(betas$rs)))
  pos <- pos[pos==3]
  pos <- names(pos)
## remove positions where most common allele is 2 (i.e. a tie)
  t1 <- al1[al1[,1]==2,]
  t2 <- al2[al2[,1]==2,]
  t.pos <- unique(c(rownames(t1), rownames(t2)))
  pos <- pos[pos%in%t.pos==FALSE]

## subset to same commonpositions for all data
  al1 <- al1[rownames(al1)%in%pos,]
  al2 <- al2[rownames(al2)%in%pos,]
  colnames(al1)[1] <- "mc1"
  colnames(al2)[1] <- "mc2"
#rownames(al1)==rownames(al2) -> ugh
#table(ugh)
  al <- cbind(al1, al2[,1])
  al <- al[,-2]
  colnames(al)[2] <- "mc2"
  al <- as.data.frame(al)
  
  bs <- betas[betas$rs%in%pos,]
  bs <- bs[order(bs$Chromosome, bs$Position),]
  #bs$rs==rownames(al) -> ugh
  #table(ugh)

### this sets up variables.
### now rotate alleles

  quad.perm <- function(){
    rot <- sample(1:(length(pos)-1), size=1)
    rot.pos <- c(rot:length(pos),1:(rot-1))
    al.up <- al[rot.pos,]

    b1 <- bs[,4:ncol(bs)]*al.up$mc1
    b2 <- bs[,4:ncol(bs)]*al.up$mc2

### get counts in each quadrant for each site/year combo
    counts <- lapply(as.list(1:ncol(b1)), function(up.beta){
      same <- b1[,up.beta]==b2[,up.beta]
      ss.pos <- b1[, up.beta]>0
      quad.nos <- table(paste(same, ss.pos))
      return(quad.nos)
    })
    counts <- do.call(rbind, counts)
    return(counts)
  }
### this does one rotation/permutation
#n.perm <- 10
  out.perm <- array(NA, dim=c(ncol(b1),4,n.perm))
  for(up in 1:n.perm){
    print(up)
    up.counts <- quad.perm()
    out.perm[,,up] <- up.counts
  }
  colnames(out.perm) <- colnames(up.counts)
  return(out.perm)
}
```

```{r do permutation, eval=FALSE}
#test.perm <- permute.quadrants(al1=al1, al2=al2, betas=betas, n.perm=10)

#al1 <- ns.allele
#al2 <- ss.allele
#betas <- betas
#n.perm <- 10
ns.perm <- permute.quadrants(al1=ns.allele, al2=ss.allele, betas=betas,n.perm=10000)
save(ns.perm, file="./data/ns.perm.Rdat")
ng.perm <- permute.quadrants(al1=ns.allele, al2=ger.allele, betas=betas,n.perm=10000)
save(ng.perm, file="./data/ng.perm.Rdat")
sg.perm <- permute.quadrants(al1=ss.allele, al2=ger.allele, betas=betas,n.perm=10000)
### need to finish this one!!!!
```

```{r observed quadrant values fxn}
quadrant.nos <- function(al1, al2, betas){
## identify common positions
  pos <- table(c(as.character(rownames(al1)),as.character(rownames(al2)), as.character(betas$rs)))
  pos <- pos[pos==3]
  pos <- names(pos)
## remove positions where most common allele is 2 (i.e. a tie)
  t1 <- al1[al1[,1]==2,]
  t2 <- al2[al2[,1]==2,]
  t.pos <- unique(c(rownames(t1), rownames(t2)))
  pos <- pos[pos%in%t.pos==FALSE]

## subset to same commonpositions for all data
  al1 <- al1[rownames(al1)%in%pos,]
  al2 <- al2[rownames(al2)%in%pos,]
  colnames(al1)[1] <- "mc1"
  colnames(al2)[1] <- "mc2"
#rownames(al1)==rownames(al2) -> ugh
#table(ugh)
  al <- cbind(al1, al2[,1])
  al <- al[,-2]
  colnames(al)[2] <- "mc2"
  al <- as.data.frame(al)
  
  bs <- betas[betas$rs%in%pos,]
  bs <- bs[order(bs$Chromosome, bs$Position),]
  #bs$rs==rownames(al) -> ugh
  #table(ugh)

### this sets up variables.
  ### get counts in each quadrant for each site/year combo
    b1 <- bs[,4:ncol(bs)]*al$mc1
    b2 <- bs[,4:ncol(bs)]*al$mc2
    counts <- lapply(as.list(1:ncol(b1)), function(up.beta){
      same <- b1[,up.beta]==b2[,up.beta]
      ss.pos <- b1[, up.beta]>0
      quad.nos <- table(paste(same, ss.pos))
      return(quad.nos)
    })
    counts <- do.call(rbind, counts)
    rownames(counts) <- colnames(b1)
    return(counts)
}
```

```{r get observed quadrant values}

ns.qn <- quadrant.nos(al1=ns.allele, al2=ss.allele, betas=betas)
ng.qn <- quadrant.nos(al1=ns.allele, al2=ger.allele, betas=betas) 
sg.qn <- quadrant.nos(al1=ss.allele, al2=ger.allele, betas=betas)
```

```{r sig of quadrant values}
up.perm <- ns.perm
up.obs <- ns.qn
col1 <- 1
col2 <- 3

quad.sig <- function(up.perm, up.obs, col1, col2){
  all.counts <- matrix(NA, ncol=4, nrow=nrow(up.obs))
  for(up in 1:nrow(up.obs)){
    exp.row <- up
    obs1 <- up.obs[exp.row, col1]
    obs2 <- up.obs[exp.row, col2]

  perm1 <- up.perm[exp.row, col1,]
  less.p1 <- obs1<=perm1
  less.p1 <- length(less.p1[less.p1==TRUE])
  more.p1 <- obs1>=perm1
  more.p1 <- length(more.p1[more.p1==TRUE])

  perm2 <- up.perm[exp.row, col2,]
  less.p2 <- obs2<=perm2
  less.p2 <- length(less.p2[less.p2==TRUE])
  more.p2 <- obs2>=perm2
  more.p2 <- length(more.p2[more.p2==TRUE])

  out.counts <- c(less.p1, more.p1, less.p2, more.p2)
  all.counts[up,] <- out.counts
  }
  rownames(all.counts) <- rownames(up.obs)
  colnames(all.counts)[3:4] <- paste(c("less", "more"),colnames(up.obs)[col2],sep="_")
  colnames(all.counts)[1:2] <- paste(c("less", "more"),colnames(up.obs)[col1],sep="_")

  return(all.counts)
}

ns.sig <- quad.sig(up.perm=ns.perm, up.obs=ns.qn, col1=1, col2=3)
ng.sig <- quad.sig(up.perm=ng.perm, up.obs=ng.qn, col1=1, col2=3)
sg.sig <- quad.sig(up.perm=sg.perm, up.obs=sg.qn, col1=1, col2=3)
#need to finish permutations for this one!
# the column here are the number of permutations for which:
#1 - opposing effects, beta of first group is less than 0 -> observed is less than permutation (bottom right of plots)
#2 - opposing effects, beta of first group is less than 0 -> observed is greater than permutation
#3 - common effects, beta of first group is less than 0 -> observed is less than permutation (bottom left of plots)
#4 - common effects, beta of first group is less than 0 -> observed is greater than permutation
```
 
```{r means of perm counts}
up.perm <- ns.perm
get.perm.mean <- function(up.perm){
  out.mean <- matrix(NA, nrow=nrow(up.perm), ncol=ncol(up.perm))
  for(up.c in 1:ncol(up.perm)){
    for(up.r in 1:nrow(up.perm)){ß
      up.mean <- mean(up.perm[up.r, up.c,])
      out.mean[up.r, up.c] <- up.mean
    }
  }
  return(out.mean)
}

ns.pm <- get.perm.mean(ns.perm)
ng.pm <- get.perm.mean(ng.perm)
```




