---
title: "26.parse.redone.admixture"
author: "Daniele Filiault"
date: "2/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(maps)
library(mapdata)
library("sf")
library("RColorBrewer")
library("rnaturalearth")
library("rnaturalearthdata")
library("rgeos")
```

## Introduction

Since 14 of our experimental accessions were not included in the 1001g admixture analysis, I redid admixture with Fernando's recall of the 1001g data (including all our 200 accessions).  I did this in /Volumes/field_experiments/001.common.reference.files/011.admixture.field.samples.

```{r load data}
admix.file <- "/Volumes/field_experiments/001.common.reference.files/011.admixture.field.samples/prunedData.9.Q"
ad.n.file <- "/Volumes/field_experiments/001.common.reference.files/011.admixture.field.samples/prunedData.fam"
geo.file <- "./data/accs_2029_ids_lat_long.modified_MN_SH.csv"
genomes.paper.file <- "./data/1001g.accessions.csv"
```


### Prepare data

```{r assemble admixture results}

admix <- read.table(admix.file, stringsAsFactors=FALSE)
colnames(admix) <- paste("g",c(1:9),sep="")
ad.n <- read.table(ad.n.file, stringsAsFactors=FALSE)
admix$id <- ad.n[,2]

admix.m <- admix
rownames(admix.m) <- admix$id
#admix.m$id <- rownames(admix.m)

max.a <- apply(admix.m[,1:9],1, function(x)(max(x, na.rm=TRUE)))
a.group <- rep(NA, nrow(admix.m))
for(up in 1:nrow(admix.m)){
  up.dat <- admix.m[up,]
  a.group[up] <- which(up.dat==max.a[up])
}
names(a.group)<- rownames(admix.m)
a.group <- cbind(max.a,a.group, admix.m)

geo <- read.csv(geo.file,stringsAsFactors=FALSE)
a.group <- merge(a.group,geo[,2:6], by.x="id", by.y="tg_ecotypeid",all.x=TRUE)

with(a.group, table(a.group,country))
ag.names <- c("N.Sweden","Germany","S.Sweden","Iberia","Italy/E.Europe","Russia/Asia","C.Europe","USA","W.Europe")  ### these are my admixture group names based on the table
a.group$group.names <- ag.names[a.group$a.group]
# identify admixed
a.group$group.names[which(a.group$max<0.65)] <- "admixed"

### output this data to use in next scripts
#write.csv(a.group, file="./data/complete.admixture.groups.csv", quote=FALSE, row.names=FALSE)
```

Data is assembled.  Let's check these groups out against the ones in the 1001g paper.  Especially concerned about all the Italians in the S. Sweden admixture group?

```{r versus.1001g}
paper.dat <- read.csv(genomes.paper.file, stringsAsFactors=FALSE,header=FALSE)
colnames(paper.dat) <- c("id","seq.by","name","country","col.loc","latitude","longitude","collector", "date","CS.no", "old.ad.group","who.knows", "ad.group.color")

tog.dat <- merge(a.group,paper.dat[,c(1:2,11)], all.x=TRUE)

with(tog.dat, table(group.names, old.ad.group))
```

There are some pretty big differences in grouping here.  For the most part, I can still group Swedish accessions into N.Sweden, S.Sweden, others, but there are some results that are quite strange.
1. The Italians in the N. Swedish admixture group.
2. The old "german" group is split up very differently
3. Spanish accessions are split up quite differently, and relicts are within one of the Spanish admixture groups.

But geographically-speaking, both groupings make sense (except for the Italians in N. Sweden - dunno what that's about.  MN said that the sequencing of those guys was strange, somehow?  Badly done, low coverage, etc.)
Let's just look at them a bit more closely:

```{R check italians in N.Swedish group}
ugh <- a.group[a.group$a.group==1 & a.group$country=="ITA",]
ital.old.dat <- tog.dat[tog.dat$id%in%ugh$id,]
```
Just mysterious.  Lots of these are the same accessions, though (sequenced at Monsanto and MPI).





### Still, I would like to make some plots of these admixture groups for the paper.
Subset data for experimental accessions
```{r experiment admixture groups}
## get info about N/C/S accession
acc.list <- read.table("./data/acc_list.txt", sep="\t", header=TRUE)
acc.list <- acc.list[grep("Sweden",acc.list$region),] ## removes caroline accessions
colnames(acc.list)[1] <- "id"
a.exp <- merge(tog.dat, acc.list[,c(1,3)], all.y=TRUE)
a.exp$region <- droplevels(a.exp$region)
with(a.exp, table(old.ad.group, group.names))
table(a.exp$group.names)
```

```{r barplot admixture results,  fig.height = 6, fig.width =9}
group.colors <- brewer.pal(9, "Paired")[c(4,7,3,2,6,5,8,1,9)]

hist(a.exp$max.a)
a.exp$region <- factor(a.exp$region, levels=c("S Sweden", "C Sweden", "N Sweden"))
a.exp <- a.exp[order(a.exp$region, a.exp$a.group, a.exp$max.a),]

c.lim <- (which(a.exp$region=="C Sweden"))
s.lim <- (which(a.exp$region=="S Sweden"))
n.lim <- (which(a.exp$region=="N Sweden"))

pdf("./figures/admixture.groups.all.experimental.lines.pdf", width=14, height=7)
par(xpd=TRUE)
barplot(t(as.matrix(a.exp[,4:12])), col=group.colors, main="All experimental lines", legend.text=ag.names, xaxt="n",space=0, args.legend=list(title="admixture group",bg="white",x="bottomright"), ylab="genome proportion", xlab="collection location in Sweden")
rect(min(c.lim)-1,-0.04,max(c.lim),-0.01, col="purple")
rect(min(s.lim)-1,-0.04,max(s.lim),-0.01, col="green")
rect(min(n.lim)-1,-0.04,max(n.lim),-0.01, col="blue")
mtext("Central",1,line=1,at=mean(c.lim)-0.5)
mtext("Southern",1,line=1,at=mean(s.lim)-0.5)
mtext("Northern",1,line=1,at=mean(n.lim)-0.5)
dev.off()

### also output in Rmd
par(xpd=TRUE)
barplot(t(as.matrix(a.exp[,4:12])), col=group.colors, main="All experimental lines", legend.text=ag.names, xaxt="n",space=0, args.legend=list(title="admixture group",bg="white",x="bottomright"), ylab="genome proportion", xlab="collection location in Sweden")
rect(min(c.lim)-1,-0.04,max(c.lim),-0.01, col="purple")
rect(min(s.lim)-1,-0.04,max(s.lim),-0.01, col="green")
rect(min(n.lim)-1,-0.04,max(n.lim),-0.01, col="blue")
mtext("Central",1,line=1,at=mean(c.lim)-0.5)
mtext("Southern",1,line=1,at=mean(s.lim)-0.5)
mtext("Northern",1,line=1,at=mean(n.lim)-0.5)
```


```{r map admixture results}

world <- ne_countries(scale = "medium", returnclass = "sf")
europe.map <- ggplot(data = world) +geom_sf() + coord_sf(xlim = c(-14, 35), ylim = c(35, 65), expand = FALSE) +   scale_x_continuous(breaks=c(0,10,20))
e.map <- europe.map + 
  geom_point(dat=tog.dat,mapping = aes(x = longitude, y = latitude, color = as.factor(a.group))) +
  theme_bw() +
  scale_color_manual(values=group.colors, name="admixture \ngroup",labels=ag.names)


pdf(file="./figures/expanded.admixture.groups.europe.map.pdf", width=5, height=6)
print(e.map)
dev.off()

print(e.map)
```






