---
title: "43.blup.slopesNS"
author: "Daniele Filiault"
date: "8/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(ggplot2)
library(gridExtra)
library(multcompView)
library(corrplot)
library(heatmaply)
library(egg)
library(gplots)
library(RColorBrewer)
#setwd("/Volumes/field_experiments/adaptation_sweden/common.gardens")
setwd("/groups/nordborg/projects/field_experiments/adaptation_sweden/common.gardens/")
```

## Introduction

Part of "local adaptation" would be having different slopes in the reaction norms - i.e. differential fitness
Let's look at this, broken down by N/S and by year using normalized BLUPs.
Idea is to use these values in a GWAS to map loci responsible for differential fitness.  
If seems rational in some way, can go back and try to do GxE GWAS specifically for experiment pairs.  This is just an easy conceptual way to start.

### 1.Prepare data
```{r load data}
blups <- read.table("./data/normalized.BLUPs.txt", stringsAsFactors=FALSE)

K.groups <- read.table("./data/K.groups.txt",stringsAsFactors=FALSE)
colnames(K.groups) <- "K.group"
K.groups$id <- rownames(K.groups)

kcol <- brewer.pal(6, "Paired")[c(6,1,2,5,3,4)]
ki <- data.frame(matrix(c("1","S1","2","S2","3","C","4","N1","5","N2","6","B"),ncol=2,byrow=TRUE))
colnames(ki) <- c("K.group","K.name")
K.groups <- merge(K.groups, ki, by="K.group")
K.groups$K.name <- factor(K.groups$K.name, levels=c("B","S1","S2","C","N1","N2"))

sshort <- data.frame(matrix(c("ADA","NA","RAM","NM","RAT","SR","ULL","SU"),ncol=2,byrow=TRUE))
colnames(sshort) <- c("site","site.s")

blups$id <- rownames(blups)
blups <- merge(blups, K.groups, by="id")

b11 <- blups[,c(1,10,11, grep("2011",colnames(blups)))]
b12 <- blups[,c(1,10,11, grep("2012",colnames(blups)))]

blups.l <- blups %>% gather(exp,blup, ADA_2011:ULL_2012)
sy <- do.call(rbind, strsplit(blups.l$exp, "_"))
sy <- as.data.frame(sy)
colnames(sy)=c("site","year")
blups.l <- cbind(blups.l, sy)
blups.l <- merge(blups.l, sshort)
```

### 2. Reaction norms 2011
```{r reaction norms 2011}
mr1.rn <- ggplot(blups.l[blups.l$year=="2011" & blups.l$site%in%c("RAM","RAT"),], aes(site.s,blup, group=id)) + geom_line(aes(colour=K.name)) + scale_color_manual(values=kcol,name = "admixture group") + xlab("Site") +ylab("normalized fitness BLUPs") + scale_x_discrete(expand = c(0.05, 0.05)) + theme_light()

mu1.rn <- ggplot(blups.l[blups.l$year=="2011" & blups.l$site%in%c("RAM","ULL"),], aes(site.s,blup, group=id)) + geom_line(aes(colour=K.name)) + scale_color_manual(values=kcol,name = "admixture group") + xlab("Site") +ylab("normalized fitness BLUPs") + scale_x_discrete(expand = c(0.05, 0.05)) + theme_light()

ar1.rn <- ggplot(blups.l[blups.l$year=="2011" & blups.l$site%in%c("ADA","RAT"),], aes(site.s,blup, group=id)) + geom_line(aes(colour=K.name)) + scale_color_manual(values=kcol,name = "admixture group") + xlab("Site") +ylab("normalized fitness BLUPs") + scale_x_discrete(expand = c(0.05, 0.05)) + theme_light()

au1.rn <- ggplot(blups.l[blups.l$year=="2011" & blups.l$site%in%c("ADA","ULL"),], aes(site.s,blup, group=id)) + geom_line(aes(colour=K.name)) + scale_color_manual(values=kcol,name = "admixture group") + xlab("Site") +ylab("normalized fitness BLUPs") + scale_x_discrete(expand = c(0.05, 0.05)) + theme_light()

grid.arrange(mr1.rn, mu1.rn, ar1.rn, au1.rn, nrow = 2)

pdf(file="./figures/rxn.norms.2011NS.normalized.pdf", width=11, height=7)
grid.arrange(mr1.rn, mu1.rn, ar1.rn, au1.rn, nrow = 2)
dev.off()
```

### 3. Reaction norms 2012
```{r reaction norms 2012}
mr2.rn <- ggplot(blups.l[blups.l$year=="2012" & blups.l$site%in%c("RAM","RAT"),], aes(site.s,blup, group=id)) + geom_line(aes(colour=K.name)) + scale_color_manual(values=kcol,name = "admixture group") + xlab("Site") +ylab("normalized fitness BLUPs") + scale_x_discrete(expand = c(0.05, 0.05)) + theme_light()

mu2.rn <- ggplot(blups.l[blups.l$year=="2012" & blups.l$site%in%c("RAM","ULL"),], aes(site.s,blup, group=id)) + geom_line(aes(colour=K.name)) + scale_color_manual(values=kcol,name = "admixture group") + xlab("Site") +ylab("normalized fitness BLUPs") + scale_x_discrete(expand = c(0.05, 0.05)) + theme_light()

ar2.rn <- ggplot(blups.l[blups.l$year=="2012" & blups.l$site%in%c("ADA","RAT"),], aes(site.s,blup, group=id)) + geom_line(aes(colour=K.name)) + scale_color_manual(values=kcol,name = "admixture group") + xlab("Site") +ylab("normalized fitness BLUPs") + scale_x_discrete(expand = c(0.05, 0.05)) + theme_light()

au2.rn <- ggplot(blups.l[blups.l$year=="2012" & blups.l$site%in%c("ADA","ULL"),], aes(site.s,blup, group=id)) + geom_line(aes(colour=K.name)) + scale_color_manual(values=kcol,name = "admixture group") + xlab("Site") +ylab("normalized fitness BLUPs") + scale_x_discrete(expand = c(0.05, 0.05)) + theme_light()

grid.arrange(mr2.rn, mu2.rn, ar2.rn, au2.rn, nrow = 2)

pdf(file="./figures/rxn.norms.2012NS.normalized.pdf", width=11, height=7)
grid.arrange(mr2.rn, mu2.rn, ar2.rn, au2.rn, nrow = 2)
dev.off()
```

### 4. Get slopes between N-S blups
```{r blup slopes}
blups$ar1.sl <- blups$RAT_2011-blups$ADA_2011
blups$ar2.sl <- blups$RAT_2012-blups$ADA_2012
blups$au1.sl <- blups$ULL_2011-blups$ADA_2011
blups$au2.sl <- blups$ULL_2012-blups$ADA_2012
blups$mr1.sl <- blups$RAT_2011-blups$RAM_2011
blups$mr2.sl <- blups$RAT_2012-blups$RAM_2012
blups$mu1.sl <- blups$ULL_2011-blups$RAM_2011
blups$mu2.sl <- blups$ULL_2012-blups$RAM_2012


```

### 5. Nonparametric tests of N/S slope by k group

```{R nonpara tests K group}

### get labels for multiple testing
get.tuk.lab <- function(up.kmc){
  tlab <- up.kmc[[3]][,3]
  names(tlab) <- rownames(up.kmc[[3]])
  tlab <- data.frame(multcompLetters(tlab)['Letters'])
  return(tlab)
}


## do tests
nblup <- blups[,c(grep("sl",colnames(blups)),1,10,11)]
kw.normfit <- as.list(rep(NA,8)) 
ph.normfit <- as.list(rep(NA,8))

for(up in 1:8){
  up.exp <- colnames(nblup[up])
  up.kt <- kruskal.test(get(up.exp)~K.name, data=nblup)
  up.ph <- kruskalmc(get(up.exp)~K.name, data=nblup)
  out.ph <- get.tuk.lab(up.ph)
  kw.normfit[[up]] <- up.kt
  ph.normfit[[up]] <- out.ph
}  

ph.normfit <- do.call(cbind,ph.normfit)
colnames(ph.normfit) <- colnames(nblup)[1:8]
ph.normfit$K.group <- rownames(ph.normfit)
```

### 6. Plot boxplots with nonparametric letters 2011

```{R nonparam plots 2011}
############
### ar1.sl
max.vals <- aggregate(nblup$ar1.sl,by=list(nblup$K.name), max, na.rm=TRUE)
max.buffer <- data.frame(max.vals[,2] + (0.1*max(max.vals[,2])))
Tukey_test <- cbind(max.buffer,ph.normfit$ar1.sl, max.vals[,1])
colnames(Tukey_test) <- c("ar1.sl","Tukey_labels","K.name")
tt.ar1 <- Tukey_test

ar1.plot <- ggplot(data=nblup, aes(x=K.name, y=ar1.sl,fill=K.name)) +
  geom_boxplot() +
  geom_jitter(width=0.1) +
  scale_fill_manual(values=kcol) +
  theme_bw() +
  labs(y="normalized fitness slope",x="K group name") +
  theme(legend.position = "none") +
  geom_text(data=tt.ar1, aes(label=tt.ar1$Tukey_labels)) + ggtitle("NA-SR slope 2011")

############
### au1.sl
max.vals <- aggregate(nblup$au1.sl,by=list(nblup$K.name), max, na.rm=TRUE)
max.buffer <- data.frame(max.vals[,2] + (0.1*max(max.vals[,2])))
Tukey_test <- cbind(max.buffer,ph.normfit$au1.sl, max.vals[,1])
colnames(Tukey_test) <- c("au1.sl","Tukey_labels","K.name")
tt.au1 <- Tukey_test

au1.plot <- ggplot(data=nblup, aes(x=K.name, y=au1.sl,fill=K.name)) +
  geom_boxplot() +
  geom_jitter(width=0.1) +
  scale_fill_manual(values=kcol) +
  theme_bw() +
  labs(y="normalized fitness slope",x="K group name") +
  theme(legend.position = "none") +
  geom_text(data=tt.au1, aes(label=tt.au1$Tukey_labels)) + ggtitle("NA-SU slope 2011")

############
### mr1.sl
max.vals <- aggregate(nblup$mr1.sl,by=list(nblup$K.name), max, na.rm=TRUE)
max.buffer <- data.frame(max.vals[,2] + (0.1*max(max.vals[,2])))
Tukey_test <- cbind(max.buffer,ph.normfit$mr1.sl, max.vals[,1])
colnames(Tukey_test) <- c("mr1.sl","Tukey_labels","K.name")
tt.mr1 <- Tukey_test

mr1.plot <- ggplot(data=nblup, aes(x=K.name, y=mr1.sl,fill=K.name)) +
  geom_boxplot() +
  geom_jitter(width=0.1) +
  scale_fill_manual(values=kcol) +
  theme_bw() +
  labs(y="normalized fitness slope",x="K group name") +
  theme(legend.position = "none") +
  geom_text(data=tt.mr1, aes(label=tt.mr1$Tukey_labels)) + ggtitle("NM-SR slope 2011")

############
### mu1.sl
max.vals <- aggregate(nblup$mu1.sl,by=list(nblup$K.name), max, na.rm=TRUE)
max.buffer <- data.frame(max.vals[,2] + (0.1*max(max.vals[,2])))
Tukey_test <- cbind(max.buffer,ph.normfit$mu1.sl, max.vals[,1])
colnames(Tukey_test) <- c("mu1.sl","Tukey_labels","K.name")
tt.mu1 <- Tukey_test

mu1.plot <- ggplot(data=nblup, aes(x=K.name, y=mu1.sl,fill=K.name)) +
  geom_boxplot() +
  geom_jitter(width=0.1) +
  scale_fill_manual(values=kcol) +
  theme_bw() +
  labs(y="normalized fitness slope",x="K group name") +
  theme(legend.position = "none") +
  geom_text(data=tt.mu1, aes(label=tt.mu1$Tukey_labels)) + ggtitle("NM-SU slope 2011")
```

```{R generate 2011 slope plot}
pdf("./figures/normalized.fitness.blup.NSslope.nonparametric.2011.by.K.pdf",width=10, height=8)
figure.nonparam <- ggarrange(ar1.plot, au1.plot, mr1.plot, mu1.plot,
          ncol = 2, nrow = 2)
annotate_figure(figure.nonparam, top = text_grob("normalized fitness BLUP slopes - KW tests", face = "bold", size = 20))
dev.off()
```

### 6. Plot boxplots with nonparametric letters 2011

```{R nonparam plots 2012}
############
### ar2.sl
max.vals <- aggregate(nblup$ar2.sl,by=list(nblup$K.name), max, na.rm=TRUE)
max.buffer <- data.frame(max.vals[,2] + (0.1*max(max.vals[,2])))
Tukey_test <- cbind(max.buffer,ph.normfit$ar2.sl, max.vals[,1])
colnames(Tukey_test) <- c("ar2.sl","Tukey_labels","K.name")
tt.ar2 <- Tukey_test

ar2.plot <- ggplot(data=nblup, aes(x=K.name, y=ar2.sl,fill=K.name)) +
  geom_boxplot() +
  geom_jitter(width=0.1) +
  scale_fill_manual(values=kcol) +
  theme_bw() +
  labs(y="normalized fitness slope",x="K group name") +
  theme(legend.position = "none") +
  geom_text(data=tt.ar2, aes(label=tt.ar2$Tukey_labels)) + ggtitle("NA-SR slope 2012")

############
### au2.sl
max.vals <- aggregate(nblup$au2.sl,by=list(nblup$K.name), max, na.rm=TRUE)
max.buffer <- data.frame(max.vals[,2] + (0.1*max(max.vals[,2])))
Tukey_test <- cbind(max.buffer,ph.normfit$au2.sl, max.vals[,1])
colnames(Tukey_test) <- c("au2.sl","Tukey_labels","K.name")
tt.au2 <- Tukey_test

au2.plot <- ggplot(data=nblup, aes(x=K.name, y=au2.sl,fill=K.name)) +
  geom_boxplot() +
  geom_jitter(width=0.1) +
  scale_fill_manual(values=kcol) +
  theme_bw() +
  labs(y="normalized fitness slope",x="K group name") +
  theme(legend.position = "none") +
  geom_text(data=tt.au2, aes(label=tt.au2$Tukey_labels)) + ggtitle("NA-SU slope 2012")

############
### mr2.sl
max.vals <- aggregate(nblup$mr2.sl,by=list(nblup$K.name), max, na.rm=TRUE)
max.buffer <- data.frame(max.vals[,2] + (0.1*max(max.vals[,2])))
Tukey_test <- cbind(max.buffer,ph.normfit$mr2.sl, max.vals[,1])
colnames(Tukey_test) <- c("mr2.sl","Tukey_labels","K.name")
tt.mr2 <- Tukey_test

mr2.plot <- ggplot(data=nblup, aes(x=K.name, y=mr2.sl,fill=K.name)) +
  geom_boxplot() +
  geom_jitter(width=0.1) +
  scale_fill_manual(values=kcol) +
  theme_bw() +
  labs(y="normalized fitness slope",x="K group name") +
  theme(legend.position = "none") +
  geom_text(data=tt.mr2, aes(label=tt.mr2$Tukey_labels)) + ggtitle("NM-SR slope 2012")

############
### mu2.sl
max.vals <- aggregate(nblup$mu2.sl,by=list(nblup$K.name), max, na.rm=TRUE)
max.buffer <- data.frame(max.vals[,2] + (0.1*max(max.vals[,2])))
Tukey_test <- cbind(max.buffer,ph.normfit$mu2.sl, max.vals[,1])
colnames(Tukey_test) <- c("mu2.sl","Tukey_labels","K.name")
tt.mu2 <- Tukey_test

mu2.plot <- ggplot(data=nblup, aes(x=K.name, y=mu2.sl,fill=K.name)) +
  geom_boxplot() +
  geom_jitter(width=0.1) +
  scale_fill_manual(values=kcol) +
  theme_bw() +
  labs(y="normalized fitness slope",x="K group name") +
  theme(legend.position = "none") +
  geom_text(data=tt.mu2, aes(label=tt.mu2$Tukey_labels)) + ggtitle("NM-SU slope 2012")
```

```{R generate 2012 slope plot}
pdf("./figures/normalized.fitness.blup.NSslope.nonparametric.2012.by.K.pdf",width=10, height=8)
figure.nonparam <- ggarrange(ar2.plot, au2.plot, mr2.plot, mu2.plot,
          ncol = 2, nrow = 2)
annotate_figure(figure.nonparam, top = text_grob("normalized fitness BLUP slopes - KW tests", face = "bold", size = 20))
dev.off()
```
### 7. correlations between slope phenotypes
```{R slope correlations}
my_cor <- cor(nblup[,1:8], use="na.or.complete")
colfunc<-colorRampPalette(c("red","white","royalblue"))
heatmaply_cor(my_cor, colors= colfunc(100),cellnote=my_cor,cellnote_textposition="middle center",k_col = 4, k_row = 4, file="./figures/pairwise.correlations.normalizes.blups.slopesNS.pdf")

```

### 8. output slopes for GWAS
```{R output slopes}
out.blups <- nblup[,c(9,1:8)]
write.csv(out.blups, file="./data/slopeNS.blups.csv", row.names=FALSE,quote=FALSE)
write.table(out.blups, file="./data/slopeNS.blups.txt", quote=FALSE)
```

### 9. GWAS done in limix with standard marginal GWAS script

### 10. Parse GWAS to get top significant SNPs 2011 and 2012
```{R top slope candidates}

setwd("/groups/nordborg/projects/field_experiments/adaptation_sweden/common.gardens/data/limix")
fs <- dir()
fs <- fs[grep("sl.limix.results.csv",fs)]
fs <- fs[-grep("png",fs)]

sig.snps <- as.list(rep(NA,length(fs)))
for(up in 1:length(fs)){
  up.f <- fs[up]
  up.d <- read.csv(file=up.f, stringsAsFactors=FALSE)
  od4 <- up.d[-log10(up.d$pv)>4,]
  #od5 <- up.d[-log10(up.d$pv)>5,] very few SNPs pass this criteria...
  up.s <- strsplit(up.f,".", fixed=TRUE)[[1]][1]
  od4$pheno <- up.s
  sig.snps[[up]] <- od4
}

sig.snps <- do.call(rbind, sig.snps)
sig.snps$year <- substr(sig.snps$pheno,3,3)
sig.snps$exps <- substr(sig.snps$pheno,1,2)
colnames(sig.snps)[1] <- "chr"

write.csv(sig.snps, file="../slope.snps.sig5.all.csv", quote=FALSE,row.names=FALSE)

### 2011 SNPS
sig1 <- sig.snps[sig.snps$year=="1",]
sig1 <- sig1[,1:2]
sig1 <- unique(sig1)
write.csv(sig1, file="../slope.snps.sig5.2011.csv", quote=FALSE, row.names=FALSE)

### 2012 SNPS
sig2 <- sig.snps[sig.snps$year=="2",]
sig2 <- sig2[,1:2]
sig2 <- unique(sig2)
write.csv(sig2, file="../slope.snps.sig5.2012.csv", quote=FALSE, row.names=FALSE)
```

### 11.  Plot mean rxn norms per K group
```{R plot mean rxn norms 2011}
km <- blups %>% 
        group_by(K.name) %>% 
        summarise(ADA_2011 = mean(ADA_2011, na.rm=TRUE),
                  ADA_2012 = mean(ADA_2012, na.rm=TRUE),
                  RAM_2011 = mean(RAM_2011, na.rm=TRUE),
                  RAM_2012 = mean(RAM_2012, na.rm=TRUE),
                  ULL_2011 = mean(ULL_2011, na.rm=TRUE),
                  ULL_2012 = mean(ULL_2012, na.rm=TRUE),
                  RAT_2011 = mean(RAT_2011, na.rm=TRUE),
                  RAT_2012 = mean(RAT_2012, na.rm=TRUE)
        )

km.l <- km %>% gather(exp,km, ADA_2011:RAT_2012)
sy <- do.call(rbind, strsplit(km.l$exp, "_"))
sy <- as.data.frame(sy)
colnames(sy)=c("site","year")
km.l <- cbind(km.l, sy)
km.l <- merge(km.l, sshort)
colnames(km.l)[4] <- "blup"

mr1.rn <- ggplot(km.l[km.l$year=="2011" & km.l$site%in%c("RAM","RAT"),], aes(site.s,blup,group=K.name)) + geom_line(aes(colour=K.name)) + scale_color_manual(values=kcol,name = "K group") + xlab("Site") +ylab("mean normalized fitness BLUPs") + scale_x_discrete(expand = c(0.05, 0.05)) + theme_light()

mu1.rn <- ggplot(km.l[km.l$year=="2011" & km.l$site%in%c("RAM","ULL"),], aes(site.s,blup,group=K.name)) + geom_line(aes(colour=K.name)) + scale_color_manual(values=kcol,name = "K group") + xlab("Site") +ylab("mean normalized fitness BLUPs") + scale_x_discrete(expand = c(0.05, 0.05)) + theme_light()

ar1.rn <- ggplot(km.l[km.l$year=="2011" & km.l$site%in%c("ADA","RAT"),], aes(site.s,blup,group=K.name)) + geom_line(aes(colour=K.name)) + scale_color_manual(values=kcol,name = "K group") + xlab("Site") +ylab("mean normalized fitness BLUPs") + scale_x_discrete(expand = c(0.05, 0.05)) + theme_light()

au1.rn <- ggplot(km.l[km.l$year=="2011" & km.l$site%in%c("ADA","ULL"),], aes(site.s,blup,group=K.name)) + geom_line(aes(colour=K.name)) + scale_color_manual(values=kcol,name = "K group") + xlab("Site") +ylab("mean normalized fitness BLUPs") + scale_x_discrete(expand = c(0.05, 0.05)) + theme_light()

grid.arrange(mr1.rn, mu1.rn, ar1.rn, au1.rn, nrow = 2)

pdf(file="./figures/rxn.norms.2011NS.normalized.K.means.pdf", width=11, height=7)
grid.arrange(mr1.rn, mu1.rn, ar1.rn, au1.rn, nrow = 2)
dev.off()
```

```{R plot mean rxn norms 2012}
mr2.rn <- ggplot(km.l[km.l$year=="2012" & km.l$site%in%c("RAM","RAT"),], aes(site.s,blup,group=K.name)) + geom_line(aes(colour=K.name)) + scale_color_manual(values=kcol,name = "K group") + xlab("Site") +ylab("mean normalized fitness BLUPs") + scale_x_discrete(expand = c(0.05, 0.05)) + theme_light()

mu2.rn <- ggplot(km.l[km.l$year=="2012" & km.l$site%in%c("RAM","ULL"),], aes(site.s,blup,group=K.name)) + geom_line(aes(colour=K.name)) + scale_color_manual(values=kcol,name = "K group") + xlab("Site") +ylab("mean normalized fitness BLUPs") + scale_x_discrete(expand = c(0.05, 0.05)) + theme_light()

ar2.rn <- ggplot(km.l[km.l$year=="2012" & km.l$site%in%c("ADA","RAT"),], aes(site.s,blup,group=K.name)) + geom_line(aes(colour=K.name)) + scale_color_manual(values=kcol,name = "K group") + xlab("Site") +ylab("mean normalized fitness BLUPs") + scale_x_discrete(expand = c(0.05, 0.05)) + theme_light()

au2.rn <- ggplot(km.l[km.l$year=="2012" & km.l$site%in%c("ADA","ULL"),], aes(site.s,blup,group=K.name)) + geom_line(aes(colour=K.name)) + scale_color_manual(values=kcol,name = "K group") + xlab("Site") +ylab("mean normalized fitness BLUPs") + scale_x_discrete(expand = c(0.05, 0.05)) + theme_light()

grid.arrange(mr2.rn, mu2.rn, ar2.rn, au2.rn, nrow = 2)

pdf(file="./figures/rxn.norms.2012NS.normalized.K.means.pdf", width=11, height=7)
grid.arrange(mr2.rn, mu2.rn, ar2.rn, au2.rn, nrow = 2)
dev.off()
```

### 12.  Plot fitness versus slope for each experiment by K group means
```{r plot fit vs slope}
# get mean slopes by K group
ksm <- blups %>% 
        group_by(K.name) %>% 
        summarise(ar1.sl = mean(ar1.sl, na.rm=TRUE),
                  ar2.sl = mean(ar2.sl, na.rm=TRUE),
                  au1.sl = mean( au1.sl, na.rm=TRUE),
                  au2.sl = mean(au2.sl, na.rm=TRUE),
                  mr1.sl = mean(mr1.sl, na.rm=TRUE),
                  mr2.sl = mean(mr2.sl, na.rm=TRUE),
                  mu1.sl = mean(mu1.sl, na.rm=TRUE),
                  mu2.sl = mean(mu2.sl, na.rm=TRUE)
        )
k.means <- merge(ksm, km)
k.means <- k.means[order(k.means$K.name),]

kssd <- blups %>% 
        group_by(K.name) %>% 
        summarise(ar1.sl = sd(ar1.sl, na.rm=TRUE),
                  ar2.sl = sd(ar2.sl, na.rm=TRUE),
                  au1.sl = sd(au1.sl, na.rm=TRUE),
                  au2.sl = sd(au2.sl, na.rm=TRUE),
                  mr1.sl = sd(mr1.sl, na.rm=TRUE),
                  mr2.sl = sd(mr2.sl, na.rm=TRUE),
                  mu1.sl = sd(mu1.sl, na.rm=TRUE),
                  mu2.sl = sd(mu2.sl, na.rm=TRUE))

ksd <- blups %>% 
        group_by(K.name) %>% 
        summarise(ADA_2011 = sd(ADA_2011, na.rm=TRUE),
                  ADA_2012 = sd(ADA_2012, na.rm=TRUE),
                  RAM_2011 = sd(RAM_2011, na.rm=TRUE),
                  RAM_2012 = sd(RAM_2012, na.rm=TRUE),
                  ULL_2011 = sd(ULL_2011, na.rm=TRUE),
                  ULL_2012 = sd(ULL_2012, na.rm=TRUE),
                  RAT_2011 = sd(RAT_2011, na.rm=TRUE),
                  RAT_2012 = sd(RAT_2012, na.rm=TRUE)
        )


## step one: scatter plot colored by K group of all obvs
up.exp <- "ULL_2011"
up.s <- "u"
up.y <- 1

up.slopes <- colnames(blups)[grep(up.s, colnames(blups))]
up.slopes <- up.slopes[grep("sl", up.slopes)]

up.dat <- blups[,c(1,10,11, which(colnames(blups)%in%c(up.exp,up.slopes)))]

up <- 1
up.s <- up.slopes[up]

scat.p <- ggplot(up.dat, aes(x=get(up.s), y=get(up.exp), color=K.name)) +
  geom_point() +
  scale_color_manual(values=kcol,name = "K group")+
  xlab(up.s) +
  ylab(up.exp) + 
  geom_point(data=k.means,colour=kcol,size=10)


## step three: add error bars

