---
title: "01.Common.Gardens.DLF.initial"
author: "Daniele Filiault"
date: "1/16/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Description and Motivation
Redoing Ben's single trait GWAS to include slugs as well
Also learning how gemma GWAS works.

Here I am starting a new markdown for the purpose of taking over Ben's analyses of the Swedish Common Garden experiments.  His old .Rmd is: /projects/field_experiments/adaptation_sweden/main.Rmd  
Before I start, I need to regenerate genotype files so I can use gemma.  Modified Ben's script prep_SNPs.sh to make prep_SNPs_DLF.sh and ran on mendel.
SNPs and other files for gemma are in:
/projects/field_experiments/001.common.reference.files/006.field.SNPs.Fernando

## 1. GWAS for all marginal traits

### 1a. Add slug damage to phenotypes

* 0 = no plant
* 1 = no damage
* 2 = light damage
* 3 = heavy damage  
Scored visually in the field in Rathckegården, Fall 2011 by DLF

```{r}
data <- read.csv("./data/rat.snail.2011.csv")
data <- data[,1:5]
data$rep <-c(rep("A",nrow(data)/3),rep("B", nrow(data)/3),rep("C",nrow(data)/3))
glm.data.t2 <- unique(data[,c(4,6)])
glm.data.t2$no.plant<-99
glm.data.t2$s1 <- 99
glm.data.t2$s2<- 99
glm.data.t2$s3<- 99

for (i in 1:nrow(glm.data.t2)){
  glm.data.t2$no.plant[i]<-nrow(data[data$line.no==glm.data.t2[i,1] & data$rep==glm.data.t2[i,2] & data$slug=="0",])
  glm.data.t2$s1[i]<-nrow(data[data$line.no==glm.data.t2[i,1] & data$rep==glm.data.t2[i,2] & data$slug=="1",])
  glm.data.t2$s2[i]<-nrow(data[data$line.no==glm.data.t2[i,1] & data$rep==glm.data.t2[i,2] & data$slug=="2",])
  glm.data.t2$s3[i]<-nrow(data[data$line.no==glm.data.t2[i,1] & data$rep==glm.data.t2[i,2] & data$slug=="3",])
}

total.plant <- apply(glm.data.t2,1,function(x) sum(as.numeric(x[4:6])))
glm.data.t2$total.plant <- total.plant
glm.data.t2$ps3 <- glm.data.t2$s3/glm.data.t2$total.plant
glm.data.t2$ps23 <- (glm.data.t2$s2+glm.data.t2$s3)/glm.data.t2$total.plant
glm.data.t2$s12 <- glm.data.t2$total.plant-glm.data.t2$s3
glm.data.t2$s23 <- glm.data.t2$total.plant-glm.data.t2$s1

glm.data.t2$rep <- as.factor(glm.data.t2$rep)
glm.data.t2$line.no <- as.factor(glm.data.t2$line.no)

to.remove <- c("201","202","203","error", "Eden-1","Ull-1","empty")
glm.data.t2 <- glm.data.t2[glm.data.t2$line.no%in%to.remove==FALSE,,drop=TRUE]
glm.data.t2$line.no <- as.character(glm.data.t2$line.no)
glm.data.t2$line.no <- as.factor(glm.data.t2$line.no)
```

Test two sets of models.  First is first is proportion s3 (heavy damage) vs s2 s1

```{r}
glm1 <- glm(cbind(s3,s12) ~ rep, data=glm.data.t2, family=binomial(link="logit")) # rep sig
glm2 <- glm(cbind(s3,s12) ~ line.no, data=glm.data.t2, family=binomial(link="logit")) # line.no sig
glm3 <- glm(cbind(s3,s12) ~ rep + line.no, data=glm.data.t2, family=binomial(link="logit")) # both sig 

```

So, both block and line are significant together.  Include both in model.

Second model is proportion s1 (not touched) vs s2 s3

```{r}
glm4 <- glm(cbind(s1,s23) ~ rep, data=glm.data.t2, family=binomial(link="logit")) # rep sig
glm5 <- glm(cbind(s1,s23) ~ line.no, data=glm.data.t2, family=binomial(link="logit")) # line.no sig
glm6 <- glm(cbind(s1,s23) ~ rep + line.no, data=glm.data.t2, family=binomial(link="logit")) # both sig together
```
Again, both block and line are significant together.  Include both in model.

Get fitted values for mapping.  These are the mean modeled proportion of plants per line with either no or severe slug damage.

```{r}
glm.data.t2$fitted.s3 <- glm3$fitted.values
s3.means <- aggregate(fitted.s3~line.no, data=glm.data.t2, mean)
hist(s3.means$fitted)

glm.data.t2$fitted.s1 <- glm6$fitted.values
s1.means <- aggregate(fitted.s1~line.no, data=glm.data.t2, mean)
hist(s1.means$fitted)

id=read.table("./data/acc_list.txt", sep="\t", h=T, na.strings=".", encoding="utf-8")
s1.means <- merge(s1.means, id, all.x=TRUE, by.x="line.no", by.y="tubes")
slug.no.damage.GWAS <- s1.means[,c(3,2)]
s3.means <- merge(s3.means, id, all.x=TRUE, by.x="line.no", by.y="tubes")
slug.severe.damage.GWAS <- s3.means[,c(3,2)]

slug.phenos <- merge(slug.no.damage.GWAS, slug.severe.damage.GWAS)
colnames(slug.phenos) <- c("id", "X2011_RAT_noslug", "X2011_RAT_severeslug")

save(slug.phenos, file="./data/slug.phenos.fitted.Rdat")

```

### 1b. Combine with BLUPs for all other phenotypes, too
These are by year and by site, so the most "marginal" possible.  
Think about sense of combining years, combining N/S for phenos?  Ben only did for fitness.  Should explore year-to-year differences more?  Or do N/S
by year for comparing with SnR estimates?

```{r}
dat <- read.table("./res/blups.txt", stringsAsFactors=FALSE, header=TRUE)
dat <- merge(dat, slug.phenos)
row.names(dat)=paste(dat$id)
dat=dat[,-1]

pref="./GWA/snps/sweden_200_MAF10.bimbam"

##make a phenotype file
acclist=scan(paste(pref, ".acclist", sep=""), sep=",")

##make a table of blups in the same order as the genotype
phen=dat[paste(acclist), ]
phenfile=paste(pref, ".all.marginal.phen.txt", sep="")
#write.table(phen, phenfile, sep="\t", row.names=F, col.names=F)
```


### 1c. Run GWAS with gemma
```{r eval=FALSE, results='hide'}

##run gemma using gemma-wrapper
## to get system() to work in .Rstudio, you need to open .RStudio in the commmand line to inhert setup.  
## command is:
## open -a RStudio

system("mkdir ./GWA/output_gemma_marginal/")
system(paste("/Users/daniele.filiault/.gem/ruby/2.0.0/bio-gemma-wrapper-0.98.1/bin/gemma-wrapper --loco --json --cache-dir ./GWA/.gemma-cache -- -gk -g ", pref, ".geno.gz -a ", pref, ".map -p ", phenfile, " > ./GWA/output_gemma_marginal/K.json", sep=""))

for(n in 1:ncol(phen)){
#for(n in 39:ncol(phen)){
    pheno=colnames(phen)[n]
    system(paste("/Users/daniele.filiault/.gem/ruby/2.0.0/bio-gemma-wrapper-0.98.1/bin/gemma-wrapper --loco --json --cache-dir ./GWA/.gemma-cache --input ./GWA/output_gemma_marginal/K.json -- -lmm 4 -g ", pref, ".geno.gz -a ", pref, ".map -p ", phenfile, " -n ", n , " > ./GWA/output_gemma_marginal/out_marginal_", pheno, sep=""))
}
```


Make manhattan plots and output 

```{r eval=FALSE, include=TRUE}
ph2=data.frame(phen=colnames(phen), ph2=NA)
gwsignif=list()
#for(n in 1:ncol(phen)){

### column 38 (X2012_ADAL_FT) didn't work - all NaN and NA in output file
### because all the data is NAs!

for(n in 39:ncol(phen)){
    pheno=colnames(phen)[n]
    print(pheno)
    x=unlist(strsplit(scan(paste("./GWA/output_gemma_marginal/out_marginal_", pheno,sep=""), what="character")[10], ","))[7]  ### had to play with this a bit to deal with a warning message that made my output different from Ben's
    f=substring(gsub("]]", "", x), 2, (nchar(x)-3))
    f2=gsub("assoc", "log", f)
    ph2$ph2[n]=ph2gemma(f2)
    p=read.delim(f, header=T, sep="\t")
    saveRDS(p, paste("./res/gemma_marginal/gemma_lmm_", pheno,".rds", sep=""))
    ##make a manhattan plot
    colnames(p)[3]="pos"
    colnames(p)=gsub("p_lrt", "pval", colnames(p))
    p$score=-log10(p$pval)
    p=p[p$af>=0.1,]
    p$fdr_pval=p.adjust(p$pval, "fdr")
    gwsignif[[pheno]]=p[p$fdr_pval<=0.05,]
    jpeg(paste("./GWA/manhattans_marginal/lmm_gwa_", pheno, ".jpeg", sep=""),res=600, unit="in", width=8, height=4, pointsize=12)
    manhattan(p)
    dev.off()
    if(n==1){
        beta=p[, c("rs","beta")]
        se=p[, c("rs", "se")]
    }else{
        beta=merge(p[, c("rs","beta")], beta, by="rs", all=T)
        se=merge(p[, c("rs","se")], se, by="rs", all=T)
    }
}
```
