---
title: "27.fitness.BLUPs"
author: "Daniele Filiault"
date: "12/4/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#setwd("/Volumes/field_experiments/adaptation_sweden/common.gardens/")

library(lme4) # mixed models
library(statmod) 
library(tweedie) # tweedie model specification
library(merTools)
library(MuMIn) # for calculating R2 of mixed models
library(RColorBrewer)
library(pander)
library(xtable)
library(dplyr)
library(tidyr)
library(ggplot2)
library(corrplot)
library(heatmaply)
library(gridExtra)
library(superheat) #extended heatmap functions
library(pheatmap)
library(gplots)
##for maps
library(maps)
library(mapdata)
library("sf")
library("rnaturalearth")
library("rnaturalearthdata")
library("rgeos")
library("egg")
library("gplots")


```

## Want to estimate BLUPs per accessions (i.e. use accession as a random effect in a mixed effects model)

### 1. prep data

```{r data prep}
d11=readRDS(file="./data/d11_for_analysis.rds")
d12=readRDS(file="./data/d12_for_analysis.rds")

##restrict to fitness
traits=c("fitness")

d11$year=2011
d12$year=2012

d11=d11[, c("year", "exp", "block", "id", "fitness")]
d12=d12[, c("year", "exp", "block", "id", "fitness")]

d=rbind(d11, d12)
d$region[d$exp%in%c("ULL", "RAT")]="S"
d$region[d$exp%in%c("RAM", "ADA")]="N"
d$year=as.factor(d$year)
d=d[, c("year", "region", "exp","block", "id", "fitness")]
d=na.omit(d)

hist(d$fitness)
hist(sqrt(d$fitness))
with(d, boxplot(fitness~exp+year))

pdf(file="./figures/27.figures/raw.fitness.boxplot.by.year.site.pdf", width=10, height=6)
with(d, boxplot(fitness~year+exp, col=c(rep(brewer.pal(9, "Paired")[4],4), rep(brewer.pal(9, "Paired")[3],4))))
dev.off()

## will also look at background effects captured as genetic or K groups in subsection 4

kg <- read.table("./data/29.data/all.clusters.and normalized.BLUPs.txt", fill=TRUE, header=TRUE)
```



### 2. fixed effect linear models
```{r fixed effect linear models}

### do as normal lm first to understand specification, then move to glm so can use tweedie
lm1 <- lm(fitness~id*year*exp, data=d)
lm3 <- lm(fitness~id*year*exp + exp:year/block, data=d)

plot(lm3)
anova(lm1,lm3) ### yes, lm3 is much better.  The block helps.
anova(lm3)

### test model fits by removing some components

lm4 <- lm(fitness~id + year*exp + exp:year/block, data=d)
lm5 <- lm(fitness~id, data=d)
lm6 <- lm(fitness~year*exp + exp:year/block, data=d)
anova(lm3, lm4)
anova(lm3, lm5)
anova(lm3, lm6)


lm.pve <- function(up.lm){
  ssq <- anova(up.lm)[[2]]
  pve <- ssq/sum(ssq) *100
  names(pve) <- rownames(anova(up.lm))
  return(pve)
}

pve3 <- lm.pve(lm3)  
pve1 <- lm.pve(lm1)  

#(summary(lm3))
pander(anova(lm3))

# make anova table with PVE
lm3.tab <- as.data.frame(anova(lm3))
lm3.tab$PVE <- pve3
rownames(lm3.tab)[8] <- "year:exp/block"
xtab.lm3 <- xtable(lm3.tab)
print(xtab.lm3, file="./figures/27.figures/fixed.effects.anova.table.txt")
```

```{r heritability per site/year}
combos <- expand.grid(unique(d$year), unique(d$exp))

check.pvals <- function(d,y,e){
  up.d <- d[d$year==y & d$exp==e,]
  up.lm <- lm(fitness~id+block, data=up.d)
  return(anova(up.lm))
  }
exp.anovas <- apply(combos, 1, function(x){check.pvals(d=d, y=x[1], e=x[2])}) # all genotypes significant, block sig in all but RAM_2012

get.pve <- function(d, y, e){
  up.d <- d[d$year==y & d$exp==e,]
  up.lm <- lm(fitness~id+block, data=up.d)
  pve <- lm.pve(up.lm)
  return(pve)
}

exp.pves <- apply(combos, 1, function(x){get.pve(d=d, y=x[1], e=x[2])})

exp.pves <- as.data.frame(t(exp.pves))
exp.pves <- cbind(combos, exp.pves)
colnames(exp.pves) <- c("year", "site", "genotype", "block", "residuals")
xtab.pve <- xtable(exp.pves)
print(xtab.pve, file="./figures/27.figures/pve.table.txt")


```

So the full model with triple interaction plus block is the best one (lm3). Plots aren't the best - you can clearly see the effect of the truncated 0 distribution. Can try this as a tweedie model?

### 3. fixed effect linear models - tweedie
```{r test tweedie models, eval=FALSE}

#profile.all.dat <- tweedie.profile(fitness~id*year*exp + exp:year/block, data = d, p.vec = seq(1.1, 1.9, 0.1),do.plot=TRUE, fit.glm = TRUE)
#print(profile.all.dat$p.max)
#vp.up <- 1.385714
#dat.up <- d

#glm.a11 <- glm(fitness~id*year*exp + exp:year/block, family=tweedie(var.power=vp.up, link.power=0), data=dat.up)

#Analysis of Deviance Table
#Model: Tweedie, link: mu^0
#Response: fitness
#Terms added sequentially (first to last)
#                Df Deviance Resid. Df Resid. Dev
#NULL                            25753     883.49
#id             199   49.732     25554     833.75
#year             1    0.385     25553     833.37
#exp              3   12.883     25550     820.48
#id:year        199   22.459     25351     798.03
#id:exp         597   35.043     24754     762.98
#year:exp         3   75.799     24751     687.18
#id:year:exp    595   29.557     24156     657.63
#year:exp:block  16   22.163     24140     635.46

### so the plots aren't a lot better on this model and the PVE (or PDE) by the variables is about the same as in the normal linear model.

### other thing to try might be a mixed effects model with block as a random effect...
```

Plots for tweedie model aren't much better.  Also, estimates of effect sizes are really similar.  Just stick with the regular linear model despite its shortcomings.

Probably more natural choice is to specify block as a random effect in a mixed effect model.  I'm going that direction to estimate BLUPs anyway...

### 4. fixed effect linear models incorporating genetic groups
Genetic groups determined in script 29
```{r full linear models with Kgroups}
dk <-merge(d, kg, all.x=TRUE)

#check models and get pve for all 7 ks
ks <- colnames(dk)[grep("k.", colnames(dk))]
ks <- ks[order(ks)]

k.anovas <- lapply(as.list(ks), function(up.k){
  up.lm <- lm(fitness~as.factor(get(up.k))*year*exp + exp:year/block, data=dk)
  return(anova(up.lm))
})

k.pve <- sapply(ks, function(up.k){
  up.lm <- lm(fitness~as.factor(get(up.k))*year*exp + exp:year/block, data=dk)
  return(lm.pve(up.lm))
})
k.pve <- as.data.frame(k.pve)
k.pve$terms <- gsub("as.factor(get(up.k))", "k.group", rownames(k.pve), fixed=TRUE)

k.pve.l <- gather(k.pve, kgroup, pve, k.2:k.8)

kterms <- k.pve[grep("k.group", k.pve$terms),"terms"]
ggplot(k.pve.l[k.pve.l$terms%in%kterms,], aes(x=kgroup, y=pve, group=terms, color=terms)) + geom_line()


## output model with k=6
lm.k6 <-  lm(fitness~k.6*year*exp + exp:year/block, data=dk)

lm.k6.tab <- as.data.frame(anova(lm.k6))
lm.k6.tab$PVE <- lm.pve(lm.k6)
rownames(lm.k6.tab)[8] <- "year:exp/block"
xtab.lm.k6 <- xtable(lm.k6.tab)
print(xtab.lm.k6, file="./figures/27.figures/fixed.effects.k6.anova.table.txt")
```
There is a big jump in pve for kgroup between k=5 and k=6.  This reinforces the idea that k=6 is appropriate

```{r by experiment linear models with 6 kgroups}
combos <- expand.grid(unique(d$year), unique(d$exp))

up.kn <- "k.6"

check.pvals.k <- function(d,y,e, up.kn){
  up.d <- d[d$year==y & d$exp==e,]
  up.lm <- lm(fitness~as.factor(get(up.kn))+block, data=up.d)
  return(anova(up.lm)$"Pr(>F)")
}

k6.ps <- apply(combos, 1, function(x){check.pvals.k(d=dk, y=x[1], e=x[2], up.kn=up.kn)}) # all genotypes significant, block sig in all but RAM_2012
k6.ps <-k6.ps[-3,]
k6.ps <- t(k6.ps)
colnames(k6.ps) <- c("p.kgroup", "p.block")

get.pve.k <- function(d, y, e, up.kn){
  up.d <- d[d$year==y & d$exp==e,]
  up.lm <- lm(fitness~as.factor(get(up.kn))+block, data=up.d)
  pve <- lm.pve(up.lm)
  return(pve)
}

k6.pve <- apply(combos, 1, function(x){get.pve.k(d=dk, y=x[1], e=x[2], up.kn="k.6")}) 
k6.pve <- as.data.frame(t(k6.pve))
k6.pve <- cbind(combos, k6.pve)
colnames(k6.pve) <- c("year", "site", "kgroup", "block", "residuals")
xtab.k6.pve <- xtable(k6.pve)
print(xtab.k6.pve, file="./figures/27.figures/pve.table.kgroup.txt")

```
Experiments with the highest kgroup PVEs are the ones that show locally-adaptive kgroup patterns (see script 29).  Expected, but still good to confirm.

### 5.  Mixed effect model to get BLUPS
```{r mixed model specifications}
mm1 <- lmer(fitness ~ 1 + (1 | block), data=d)
mm2 <- lmer(fitness ~ 1 + (1 | id), data=d)
mm3 <- lmer(fitness~exp + (1|exp:block), data=d) # so this is correctly specified (without year effect, at least)
mm4 <- lmer(fitness~exp + year + (1|exp:year:block), data=d)  ## I think this is also correctly specified!
ranef(mm4)
fixef(mm4)
mm5 <- lmer(fitness~exp*year + (1|exp:year:block), data=d)
anova(mm4, mm5)
ranef(mm5)
fixef(mm5)

mm6 <- lmer(fitness~exp*year + (1|exp:year:block) + (1|id), data=d)
anova(mm5,mm6)

mm7 <- lmer(fitness~exp*year + (1|exp:year:block) + (1|exp:year:id), data=d)
anova(mm6,mm7)

mm8 <- lmer(fitness~exp*year  + (1|exp:year:id), data=d)

c2 <- predictInterval(mm7)
int.mm7 <- REsim(mm7)
plotREsim(REsim(mm7))

ranks <- expectedRank(mm7)

VarCorr(mm7)
anova(mm7)
summary(mm7)

# random effect variances
var.mm7 <- VarCorr(mm7)
print(var.mm7,comp=c("Variance","Std.Dev."),digits=4)
re.var <- as.data.frame(var.mm7, row.names = NULL,optional = FALSE)

r.squaredGLMM(mm5)
r.squaredGLMM(mm6)
r.squaredGLMM(mm7)
r.squaredGLMM(mm8)
## R2m is from marginal(fixed effects)
## R2c is from entire model
```

So now let's think about these BLUPs per accession per experiment.
Can we use them to demonstrate local adaptation?
Can they be used for GWAS?

## 5. explore BLUP relationships

```{r explore blups}
blups <- ranef(mm7)[[1]]
blups <- as.data.frame(blups)
desc <- do.call(rbind,strsplit(rownames(blups),":"))
colnames(desc) <- c("exp","year","id")
blups <- cbind(blups, desc)
colnames(blups)[1] <- "fb"

hist(blups$fb)
pdf(file="./figures/27.figures/blup.fitness.boxplot.by.year.site.pdf", width=10, height=6)
with(blups, boxplot(fb~year+exp, col=c(rep(brewer.pal(9, "Paired")[4],4), rep(brewer.pal(9, "Paired")[3],4))))
dev.off()

## write BLUPS to file
write.csv(blups, file="./data/27.data/marginal.blups.csv", quote=FALSE,row.names=FALSE)

## save model to file
save(mm7, file="./data/27.data/mm7.final.mixed.model.Rdat")
```


```{r load admixture results, eval=FALSE}
# parsed in 26.parse.redone.admixture.Rmd
# depreciated
a.group <- read.csv("./data/complete.admixture.groups.csv", stringsAsFactors = FALSE)
```


```{r add fitness blups plot rxn norms}
acc.fit.rxn <- blups
acc.fit.rxn <- merge(acc.fit.rxn, a.group[,c(1:3,17)], all.x=TRUE)
acc.fit.rxn$a.group <- as.factor(acc.fit.rxn$a.group)
acc.fit.rxn$site.complex <- paste(acc.fit.rxn$exp, acc.fit.rxn$year, sep="_")
acc.fit.rxn$group.names <- as.factor(acc.fit.rxn$group.names)
acc.fit.rxn$group.names = factor(acc.fit.rxn$group.names,levels(acc.fit.rxn$group.names)[c(5,3,6,4,2,7,1)])

pdf(file="./figures/27.figures/reaction.norms.blups.pdf", width=8, height=5)
rxn.norms <- ggplot(acc.fit.rxn, aes(site.complex,fb, group=id)) + 
  geom_line(aes(colour=group.names)) + 
  scale_color_manual(values=group.colors[c(1,2,3,4,7,9,8)],name = "admixture group") +
  xlab("Site_year") +
  ylab("fitness BLUPS")+
  guides(color = guide_legend(override.aes = list(size = 3)))
print(rxn.norms)
dev.off()
```

```{r pairwise correlations between blups, eval=FALSE}
## need to reformat blups to long format
blups <- read.csv(file="./data/27.data/marginal.blups.csv", stringsAsFactors=FALSE)
blups.l <- blups %>% pivot_wider(id_cols=id,names_from=c(exp,year),values_from=fb)%>%as.data.frame()
colnames(blups.l) <- c("id","NA_2011","NA_2012","NM_2011", "NM_2012","SR_2011", "SR_2012","SU_2011","SU_2012")

pdf("./figures/27.figures/correlation.fitness.blups.pdf", width=8, height=8)
corrplot(cor(blups.l[,2:9], use="na.or.complete"), method="ellipse", type="upper", diag=FALSE)
dev.off()

my_cor <- cor(blups.l[,2:9], use="na.or.complete")
colfunc<-colorRampPalette(c("red","white","royalblue"))
#heatmaply_cor(my_cor, colors= colfunc(100),cellnote=my_cor,cellnote_textposition="middle center")
### since these are interactive plots in plotly, they are hard to save statically, so I saved this one manually from the window
#file="./figures/correlation.dendrogram.fitness.blups.jpeg")

#try using egg package
##ggheatmap(
  #mtcars,
  #scale = "column",
  #row_side_colors = mtcars[, c("cyl", "gear")]
#)

ns.site <- as.data.frame(c(rep("North",4),rep("South",4)))
ns.palette = c("North" = "green2", "South" = "dodgerblue2")
colnames(ns.site) <- "location"
ns.cols <- c(rep("green2",4), rep("dodgerblue2", 4))

mycolors <- colorRampPalette(c("blue", "lightyellow1", "red"))
heatmap.2(my_cor, trace = "none", col = mycolors)

my_cor_na <- my_cor
rna <- which(my_cor==1)
my_cor_na[rna] <- NA

pdf(file="./figures/27.figures/blup.correlation.heatmap.pdf", width=10, height=9)
heatmap.2(my_cor, trace = "none", col = mycolors, margins=c(6.5,6.5), colsep=c(3,5), rowsep=c(3,5),ColSideColors=ns.cols, RowSideColors=ns.cols)
dev.off()

```

```{r blup values by experiment and admixture group}
ga.long.short <- acc.fit.rxn
ga.long.short$exp <- gsub("ADA","NA", ga.long.short$exp)
ga.long.short$exp <- gsub("RAM","NM", ga.long.short$exp)
ga.long.short$exp <- gsub("RAT","SR", ga.long.short$exp)
ga.long.short$exp <- gsub("ULL","SU", ga.long.short$exp)
### remove admixture groups with only one accession
acc.tab <- table(ga.long.short$group.names)
ga.long.short <- ga.long.short[ga.long.short$group.names%in%names(acc.tab[acc.tab>8]),]
ga.long.short$group.names = droplevels(ga.long.short$group.names)
ga.long.short$group.names <- factor(ga.long.short$group.names, level=c("C.Europe","S.Sweden", "N.Sweden","admixed"))

ad.group.short.boxplot <- ggplot(ga.long.short, aes(x=group.names, y=fb, fill=as.factor(group.names))) +
  geom_violin() +
  geom_boxplot(width=0.1, show.legend=FALSE)+
  xlab("admixture group")+ 
  facet_grid(year~exp) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_manual(name = "admixture group", values=brewer.pal(9, "Paired")[c(8,4,3,1)]) +
  ylab("fitness BLUPS")

pdf(file="./figures/27.figures/fitness.blups.by.admix.group.short.names.pdf", width=10, height=6)
print(ad.group.short.boxplot)
dev.off()

```

### reaction norms between years, same site

```{r ind rxn norms}
#test.sub <- acc.fit.rxn[acc.fit.rxn$year=="2011" & acc.fit.rxn$exp%in%c("RAT","ULL"),]
ada.sub <- acc.fit.rxn[acc.fit.rxn$exp%in%c("ADA"),]
ram.sub <- acc.fit.rxn[acc.fit.rxn$exp%in%c("RAM"),]
ull.sub <- acc.fit.rxn[acc.fit.rxn$exp%in%c("ULL"),]
rat.sub <- acc.fit.rxn[acc.fit.rxn$exp%in%c("RAT"),]


ada.rn <- ggplot(ada.sub, aes(site.complex,fb, group=id)) + geom_line(aes(colour=group.names)) + scale_color_manual(values=group.colors[c(1,2,3,4,7,9,8)],name = "admixture group") + xlab("Site_year") +ylab("fitness BLUPS") + scale_x_discrete(expand = c(0.05, 0.05)) + theme_light()

ram.rn <- ggplot(ram.sub, aes(site.complex,fb, group=id)) + geom_line(aes(colour=group.names)) + scale_color_manual(values=group.colors[c(1,2,3,4,7,9,8)],name = "admixture group") + xlab("Site_year") +ylab("fitness BLUPS") + scale_x_discrete(expand = c(0.05, 0.05)) + theme_light()


ull.rn <- ggplot(ull.sub, aes(site.complex,fb, group=id)) + geom_line(aes(colour=group.names)) + scale_color_manual(values=group.colors[c(1,2,3,4,7,9,8)],name = "admixture group") + xlab("Site_year") +ylab("fitness BLUPS") + scale_x_discrete(expand = c(0.05, 0.05)) + theme_light()


rat.rn <- ggplot(rat.sub, aes(site.complex,fb, group=id)) + geom_line(aes(colour=group.names)) + scale_color_manual(values=group.colors[c(1,2,3,4,7,9,8)],name = "admixture group") + xlab("Site_year") +ylab("fitness BLUPS") + scale_x_discrete(expand = c(0.05, 0.05)) + theme_light()

grid.arrange(ada.rn, ram.rn, ull.rn, rat.rn, nrow = 2)

pdf(file="./figures/27.figures/site.rxn.norms.pdf", width=11, height=7)
grid.arrange(ada.rn, ram.rn, ull.rn, rat.rn, nrow = 2)
dev.off()
```

### reaction norms between sites, 2011

```{r 2011 rxn norms}
up.year <- "2011"
au.sub <- acc.fit.rxn[acc.fit.rxn$year==up.year & acc.fit.rxn$exp%in%c("ADA","ULL"),]
at.sub <- acc.fit.rxn[acc.fit.rxn$year==up.year & acc.fit.rxn$exp%in%c("ADA","RAT"),]
ru.sub <- acc.fit.rxn[acc.fit.rxn$year==up.year & acc.fit.rxn$exp%in%c("RAM","ULL"),]
rt.sub <- acc.fit.rxn[acc.fit.rxn$year==up.year & acc.fit.rxn$exp%in%c("RAM","RAT"),]
ar.sub <- acc.fit.rxn[acc.fit.rxn$year==up.year & acc.fit.rxn$exp%in%c("ADA","RAM"),]
ut.sub <- acc.fit.rxn[acc.fit.rxn$year==up.year & acc.fit.rxn$exp%in%c("ULL","RAT"),]


au.rn <- ggplot(au.sub, aes(site.complex,fb, group=id)) + geom_line(aes(colour=group.names)) + scale_color_manual(values=group.colors[c(1,2,3,4,7,9,8)],name = "admixture group") + xlab("Site_year") +ylab("fitness BLUPs") + scale_x_discrete(expand = c(0.05, 0.05)) + theme_light()

at.rn <- ggplot(at.sub, aes(site.complex,fb, group=id)) + geom_line(aes(colour=group.names)) + scale_color_manual(values=group.colors[c(1,2,3,4,7,9,8)],name = "admixture group") + xlab("Site_year") +ylab("fitness BLUPs") + scale_x_discrete(expand = c(0.05, 0.05)) + theme_light()

ru.rn <- ggplot(ru.sub, aes(site.complex,fb, group=id)) + geom_line(aes(colour=group.names)) + scale_color_manual(values=group.colors[c(1,2,3,4,7,9,8)],name = "admixture group") + xlab("Site_year") +ylab("fitness BLUPs") + scale_x_discrete(expand = c(0.05, 0.05)) + theme_light()

rt.rn <- ggplot(rt.sub, aes(site.complex,fb, group=id)) + geom_line(aes(colour=group.names)) + scale_color_manual(values=group.colors[c(1,2,3,4,7,9,8)],name = "admixture group") + xlab("Site_year") +ylab("fitness BLUPs") + scale_x_discrete(expand = c(0.05, 0.05)) + theme_light()

ar.rn <- ggplot(ar.sub, aes(site.complex,fb, group=id)) + geom_line(aes(colour=group.names)) + scale_color_manual(values=group.colors[c(1,2,3,4,7,9,8)],name = "admixture group") + xlab("Site_year") +ylab("fitness BLUPs") + scale_x_discrete(expand = c(0.05, 0.05)) + theme_light()

ut.rn <- ggplot(ut.sub, aes(site.complex,fb, group=id)) + geom_line(aes(colour=group.names)) + scale_color_manual(values=group.colors[c(1,2,3,4,7,9,8)],name = "admixture group") + xlab("Site_year") +ylab("fitness BLUPs") + scale_x_discrete(expand = c(0.05, 0.05)) + theme_light()

grid.arrange(au.rn, at.rn, ru.rn, rt.rn, ar.rn, ut.rn, nrow = 3)

pdf(file="./figures/27.figures/2011.rxn.norms.pdf", width=9, height=11)
grid.arrange(au.rn, at.rn, ru.rn, rt.rn, ar.rn, ut.rn, nrow = 3)
dev.off()

```
### reaction norms between sites, 2012

```{r 2012 rxn norms}
up.year <- "2012"
au.sub <- acc.fit.rxn[acc.fit.rxn$year==up.year & acc.fit.rxn$exp%in%c("ADA","ULL"),]
at.sub <- acc.fit.rxn[acc.fit.rxn$year==up.year & acc.fit.rxn$exp%in%c("ADA","RAT"),]
ru.sub <- acc.fit.rxn[acc.fit.rxn$year==up.year & acc.fit.rxn$exp%in%c("RAM","ULL"),]
rt.sub <- acc.fit.rxn[acc.fit.rxn$year==up.year & acc.fit.rxn$exp%in%c("RAM","RAT"),]
ar.sub <- acc.fit.rxn[acc.fit.rxn$year==up.year & acc.fit.rxn$exp%in%c("ADA","RAM"),]
ut.sub <- acc.fit.rxn[acc.fit.rxn$year==up.year & acc.fit.rxn$exp%in%c("ULL","RAT"),]


au.rn <- ggplot(au.sub, aes(site.complex,fb, group=id)) + geom_line(aes(colour=group.names)) + scale_color_manual(values=group.colors[c(1,2,3,4,7,9,8)],name = "admixture group") + xlab("Site_year") +ylab("fitness BLUPs") + scale_x_discrete(expand = c(0.05, 0.05)) + theme_light()

at.rn <- ggplot(at.sub, aes(site.complex,fb, group=id)) + geom_line(aes(colour=group.names)) + scale_color_manual(values=group.colors[c(1,2,3,4,7,9,8)],name = "admixture group") + xlab("Site_year") +ylab("fitness BLUPs") + scale_x_discrete(expand = c(0.05, 0.05)) + theme_light()

ru.rn <- ggplot(ru.sub, aes(site.complex,fb, group=id)) + geom_line(aes(colour=group.names)) + scale_color_manual(values=group.colors[c(1,2,3,4,7,9,8)],name = "admixture group") + xlab("Site_year") +ylab("fitness BLUPs") + scale_x_discrete(expand = c(0.05, 0.05)) + theme_light()

rt.rn <- ggplot(rt.sub, aes(site.complex,fb, group=id)) + geom_line(aes(colour=group.names)) + scale_color_manual(values=group.colors[c(1,2,3,4,7,9,8)],name = "admixture group") + xlab("Site_year") +ylab("fitness BLUPs") + scale_x_discrete(expand = c(0.05, 0.05)) + theme_light()

ar.rn <- ggplot(ar.sub, aes(site.complex,fb, group=id)) + geom_line(aes(colour=group.names)) + scale_color_manual(values=group.colors[c(1,2,3,4,7,9,8)],name = "admixture group") + xlab("Site_year") +ylab("fitness BLUPs") + scale_x_discrete(expand = c(0.05, 0.05)) + theme_light()

ut.rn <- ggplot(ut.sub, aes(site.complex,fb, group=id)) + geom_line(aes(colour=group.names)) + scale_color_manual(values=group.colors[c(1,2,3,4,7,9,8)],name = "admixture group") + xlab("Site_year") +ylab("fitness BLUPs") + scale_x_discrete(expand = c(0.05, 0.05)) + theme_light()

grid.arrange(au.rn, at.rn, ru.rn, rt.rn, ar.rn, ut.rn, nrow = 3)

pdf(file="./figures/27.figures/2012.rxn.norms.pdf", width=9, height=11)
grid.arrange(au.rn, at.rn, ru.rn, rt.rn, ar.rn, ut.rn, nrow = 3)
dev.off()

```

### patterns of significance and effect by line - per site and year
```{r sig pattern}
line.est <- int.mm7[int.mm7$groupFctr=="exp:year:id",] ### 2 of 1600 missing

## does CI contain zero?
line.est$ci.low <- line.est$mean-line.est$sd
line.est$ci.hi <- line.est$mean+line.est$sd
line.est$sig <- apply(line.est,1, function(x){(as.numeric(x[7])<0 & as.numeric(x[8])>0)==FALSE})
### 39% of CIs don't contain zero
line.est$mean.sig <- line.est$mean*line.est$sig

## reshape dataframe
tmp.ids <- do.call(rbind, sapply(line.est$groupID, function(x){strsplit(x,":")}))
tmp.ids <- data.frame(tmp.ids)
colnames(tmp.ids) <- c("site","year", "id")
line.est <- cbind(line.est, tmp.ids)
line.est$exp <- paste(line.est$site, line.est$year, sep="_")

test <- line.est[,c(10,13,14)]
test <- spread(test, exp, mean.sig)

write.table(line.est, file="./data/27.data/line.fitness.estimate.significance.txt")
```

