---
title: "27.fitness.BLUPs"
author: "Daniele Filiault"
date: "12/4/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#setwd("/Volumes/field_experiments/adaptation_sweden/common.gardens/")

library(lme4) # mixed models
library(statmod) 
library(tweedie) # tweedie model specification
library(merTools)
library(MuMIn) # for calculating R2 of mixed models
library(RColorBrewer)
library(pander)
library(dplyr)
library(tidyr)
library(ggplot2)
library(corrplot)
library(heatmaply)
library(gridExtra)
library(superheat) #extended heatmap functions
library(pheatmap)
##for maps
library(maps)
library(mapdata)
library("sf")
library("rnaturalearth")
library("rnaturalearthdata")
library("rgeos")
library("egg")
library("gplots")

### for admixture group plots
ag.names <- c("N.Sweden","Germany","S.Sweden","Iberia","Italy/E.Europe","Russia/Asia","C.Europe","USA","W.Europe")  ### these are my admixture group names based on the table
group.colors <- brewer.pal(9, "Paired")[c(4,7,3,2,6,5,8,1,9)]

```

## Want to estimate BLUPs per accessions (i.e. use accession as a random effect in a mixed effects model)

### 1. prep data

```{r data prep}
d11=readRDS(file="./data/d11_for_analysis.rds")
d12=readRDS(file="./data/d12_for_analysis.rds")

##restrict to fitness
traits=c("fitness")

d11$year=2011
d12$year=2012

d11=d11[, c("year", "exp", "block", "id", "fitness")]
d12=d12[, c("year", "exp", "block", "id", "fitness")]

d=rbind(d11, d12)
d$region[d$exp%in%c("ULL", "RAT")]="S"
d$region[d$exp%in%c("RAM", "ADA")]="N"
d$year=as.factor(d$year)
d=d[, c("year", "region", "exp","block", "id", "fitness")]
d=na.omit(d)

hist(d$fitness)
hist(sqrt(d$fitness))
with(d, boxplot(fitness~exp+year))

pdf(file="./figures/raw.fitness.boxplot.by.year.site.pdf", width=10, height=6)
with(d, boxplot(fitness~year+exp, col=c(rep(brewer.pal(9, "Paired")[4],4), rep(brewer.pal(9, "Paired")[3],4))))
dev.off()

```



### 2. function to get predicted fitness per block/line/exp
```{r test normal linear models}

### do as normal lm first to understand specification, then move to glm so can use tweedie
lm1 <- lm(fitness~id*year*exp, data=d)
lm3 <- lm(fitness~id*year*exp + exp:year/block, data=d)
plot(lm3)
anova(lm1,lm3) ### yes, lm3 is much better.  The block helps. If it is specified correctly???
anova(lm3)

### test model fits by removing some components

lm4 <- lm(fitness~id + year*exp + exp:year/block, data=d)
lm5 <- lm(fitness~id, data=d)
lm6 <- lm(fitness~year*exp + exp:year/block, data=d)
anova(lm3, lm4)
anova(lm3, lm5)
anova(lm3, lm6)

ssq <- anova(lm3)[[2]]
pve <-ssq/sum(ssq)
names(pve) <- rownames(anova(lm3))
print(pve)

#(summary(lm3))
pander(anova(lm3))

```

So the full model with triple interaction plus block is the best one (lm3). Plots aren't the best - you can clearly see the effect of the truncated 0 distribution. Can try this as a tweedie model?

```{r test tweedie models, eval=FALSE}

#profile.all.dat <- tweedie.profile(fitness~id*year*exp + exp:year/block, data = d, p.vec = seq(1.1, 1.9, 0.1),do.plot=TRUE, fit.glm = TRUE)
#print(profile.all.dat$p.max)
#vp.up <- 1.385714
#dat.up <- d

#glm.a11 <- glm(fitness~id*year*exp + exp:year/block, family=tweedie(var.power=vp.up, link.power=0), data=dat.up)

#Analysis of Deviance Table
#Model: Tweedie, link: mu^0
#Response: fitness
#Terms added sequentially (first to last)
#                Df Deviance Resid. Df Resid. Dev
#NULL                            25753     883.49
#id             199   49.732     25554     833.75
#year             1    0.385     25553     833.37
#exp              3   12.883     25550     820.48
#id:year        199   22.459     25351     798.03
#id:exp         597   35.043     24754     762.98
#year:exp         3   75.799     24751     687.18
#id:year:exp    595   29.557     24156     657.63
#year:exp:block  16   22.163     24140     635.46

### so the plots aren't a lot better on this model and the PVE (or PDE) by the variables is about the same as in the normal linear model.

### other thing to try might be a mixed effects model with block as a random effect...
```

Plots for tweedie model aren't much better.  Also, estimates of effect sizes are really similar.  Just stick with the regular linear model despite its shortcomings.

I'm not 100% sure that I'm specifying block correctly in model.  Make unique block names and compare?
Probably better choice is to specify block as a random effect in a mixed effect model.  I'm going that direction to estimate BLUPs anyway...

```{r mixed model specifications}
mm1 <- lmer(fitness ~ 1 + (1 | block), data=d)
mm2 <- lmer(fitness ~ 1 + (1 | id), data=d)
mm3 <- lmer(fitness~exp + (1|exp:block), data=d) # so this is correctly specified (without year effect, at least)
mm4 <- lmer(fitness~exp + year + (1|exp:year:block), data=d)  ## I think this is also correctly specified!
ranef(mm4)
fixef(mm4)
mm5 <- lmer(fitness~exp*year + (1|exp:year:block), data=d)
anova(mm4, mm5)
ranef(mm5)
fixef(mm5)

mm6 <- lmer(fitness~exp*year + (1|exp:year:block) + (1|id), data=d)
anova(mm5,mm6)

mm7 <- lmer(fitness~exp*year + (1|exp:year:block) + (1|exp:year:id), data=d)
anova(mm6,mm7)

mm8 <- lmer(fitness~exp*year  + (1|exp:year:id), data=d)

c2 <- predictInterval(mm7)
int.mm7 <- REsim(mm7)
plotREsim(REsim(mm7))

ranks <- expectedRank(mm7)

VarCorr(mm7)
anova(mm7)
summary(mm7)

# random effect variances
var.mm7 <- VarCorr(mm7)
print(var.mm7,comp=c("Variance","Std.Dev."),digits=4)
re.var <- as.data.frame(var.mm7, row.names = NULL,optional = FALSE)

r.squaredGLMM(mm5)
r.squaredGLMM(mm6)
r.squaredGLMM(mm7)
r.squaredGLMM(mm8)
## R2m is from marginal(fixed effects)
## R2c is from entire model
```

So now let's think about these BLUPs per accession per experiment.
Can we use them to demonstrate local adaptation?
Can they be used for GWAS?

```{r explore blups}
blups <- ranef(mm7)[[1]]
blups <- as.data.frame(blups)
desc <- do.call(rbind,strsplit(rownames(blups),":"))
colnames(desc) <- c("exp","year","id")
blups <- cbind(blups, desc)
colnames(blups)[1] <- "fb"

hist(blups$fb)
pdf(file="./figures/blup.fitness.boxplot.by.year.site.pdf", width=10, height=6)
with(blups, boxplot(fb~year+exp, col=c(rep(brewer.pal(9, "Paired")[4],4), rep(brewer.pal(9, "Paired")[3],4))))
dev.off()

## write BLUPS to file
write.csv(blups, file="./data/marginal.blups.csv", quote=FALSE,row.names=FALSE)

```


```{r load admixture results}
# parsed in 26.parse.redone.admixture.Rmd
a.group <- read.csv("./data/complete.admixture.groups.csv", stringsAsFactors = FALSE)
```


```{r add fitness blups plot rxn norms}
acc.fit.rxn <- blups
acc.fit.rxn <- merge(acc.fit.rxn, a.group[,c(1:3,17)], all.x=TRUE)
acc.fit.rxn$a.group <- as.factor(acc.fit.rxn$a.group)
acc.fit.rxn$site.complex <- paste(acc.fit.rxn$exp, acc.fit.rxn$year, sep="_")
acc.fit.rxn$group.names <- as.factor(acc.fit.rxn$group.names)
acc.fit.rxn$group.names = factor(acc.fit.rxn$group.names,levels(acc.fit.rxn$group.names)[c(5,3,6,4,2,7,1)])

pdf(file="./figures/reaction.norms.blups.pdf", width=8, height=5)
rxn.norms <- ggplot(acc.fit.rxn, aes(site.complex,fb, group=id)) + 
  geom_line(aes(colour=group.names)) + 
  scale_color_manual(values=group.colors[c(1,2,3,4,7,9,8)],name = "admixture group") +
  xlab("Site_year") +
  ylab("fitness BLUPS")+
  guides(color = guide_legend(override.aes = list(size = 3)))
print(rxn.norms)
dev.off()
```

```{r pairwise correlations between blups}
## need to reformat blups to long format
blups <- read.csv(file="./data/marginal.blups.csv", stringsAsFactors=FALSE)
blups.l <- blups %>% pivot_wider(id_cols=id,names_from=c(exp,year),values_from=fb)%>%as.data.frame()
colnames(blups.l) <- c("id","NA_2011","NA_2012","NM_2011", "NM_2012","SR_2011", "SR_2012","SU_2011","SU_2012")

pdf("./figures/correlation.fitness.blups.pdf", width=8, height=8)
corrplot(cor(blups.l[,2:9], use="na.or.complete"), method="ellipse", type="upper", diag=FALSE)
dev.off()

my_cor <- cor(blups.l[,2:9], use="na.or.complete")
colfunc<-colorRampPalette(c("red","white","royalblue"))
heatmaply_cor(my_cor, colors= colfunc(100),cellnote=my_cor,cellnote_textposition="middle center")
### since these are interactive plots in plotly, they are hard to save statically, so I saved this one manually from the window
#file="./figures/correlation.dendrogram.fitness.blups.jpeg")

#try using egg package
##ggheatmap(
  #mtcars,
  #scale = "column",
  #row_side_colors = mtcars[, c("cyl", "gear")]
#)

ns.site <- as.data.frame(c(rep("North",4),rep("South",4)))
ns.palette = c("North" = "green2", "South" = "dodgerblue2")
colnames(ns.site) <- "location"
cor.hm <- ggheatmap(my_cor,row_side_colors = ns.site, row_side_palette=ns.palette,show_dendrogram = c(TRUE, FALSE),margins = c(40, 130))
print(cor.hm)
pdf(file="./figures/blup.correlation.heatmap.pdf", width=8, height=5)
print(cor.hm)
dev.off()
### so this isn't exactly right - missing a bar in the dendrogram, but I can't fix right now
```

```{r blup values by experiment and admixture group}
ga.long.short <- acc.fit.rxn
ga.long.short$exp <- gsub("ADA","N2", ga.long.short$exp)
ga.long.short$exp <- gsub("RAM","N1", ga.long.short$exp)
ga.long.short$exp <- gsub("RAT","S2", ga.long.short$exp)
ga.long.short$exp <- gsub("ULL","S1", ga.long.short$exp)
### remove admixture groups with only one accession
acc.tab <- table(ga.long.short$group.names)
ga.long.short <- ga.long.short[ga.long.short$group.names%in%names(acc.tab[acc.tab>8]),]
ga.long.short$group.names = droplevels(ga.long.short$group.names)
ga.long.short$group.names <- factor(ga.long.short$group.names, level=c("C.Europe","S.Sweden", "N.Sweden","admixed"))

ad.group.short.boxplot <- ggplot(ga.long.short, aes(x=group.names, y=fb, fill=as.factor(group.names))) +
  geom_violin() +
  geom_boxplot(width=0.1, show.legend=FALSE)+
  xlab("admixture group")+ 
  facet_grid(year~exp) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_manual(name = "admixture group", values=brewer.pal(9, "Paired")[c(8,4,3,1)]) +
  ylab("fitness BLUPS")

pdf(file="./figures/fitness.blups.by.admix.group.short.names.pdf", width=10, height=6)
print(ad.group.short.boxplot)
dev.off()

```

### reaction norms between years, same site

```{r ind rxn norms}
#test.sub <- acc.fit.rxn[acc.fit.rxn$year=="2011" & acc.fit.rxn$exp%in%c("RAT","ULL"),]
ada.sub <- acc.fit.rxn[acc.fit.rxn$exp%in%c("ADA"),]
ram.sub <- acc.fit.rxn[acc.fit.rxn$exp%in%c("RAM"),]
ull.sub <- acc.fit.rxn[acc.fit.rxn$exp%in%c("ULL"),]
rat.sub <- acc.fit.rxn[acc.fit.rxn$exp%in%c("RAT"),]


ada.rn <- ggplot(ada.sub, aes(site.complex,fb, group=id)) + geom_line(aes(colour=group.names)) + scale_color_manual(values=group.colors[c(1,2,3,4,7,9,8)],name = "admixture group") + xlab("Site_year") +ylab("fitness BLUPS") + scale_x_discrete(expand = c(0.05, 0.05)) + theme_light()

ram.rn <- ggplot(ram.sub, aes(site.complex,fb, group=id)) + geom_line(aes(colour=group.names)) + scale_color_manual(values=group.colors[c(1,2,3,4,7,9,8)],name = "admixture group") + xlab("Site_year") +ylab("fitness BLUPS") + scale_x_discrete(expand = c(0.05, 0.05)) + theme_light()


ull.rn <- ggplot(ull.sub, aes(site.complex,fb, group=id)) + geom_line(aes(colour=group.names)) + scale_color_manual(values=group.colors[c(1,2,3,4,7,9,8)],name = "admixture group") + xlab("Site_year") +ylab("fitness BLUPS") + scale_x_discrete(expand = c(0.05, 0.05)) + theme_light()


rat.rn <- ggplot(rat.sub, aes(site.complex,fb, group=id)) + geom_line(aes(colour=group.names)) + scale_color_manual(values=group.colors[c(1,2,3,4,7,9,8)],name = "admixture group") + xlab("Site_year") +ylab("fitness BLUPS") + scale_x_discrete(expand = c(0.05, 0.05)) + theme_light()

grid.arrange(ada.rn, ram.rn, ull.rn, rat.rn, nrow = 2)

pdf(file="./figures/site.rxn.norms.pdf", width=11, height=7)
grid.arrange(ada.rn, ram.rn, ull.rn, rat.rn, nrow = 2)
dev.off()
```

### reaction norms between sites, 2011

```{r 2011 rxn norms}
up.year <- "2011"
au.sub <- acc.fit.rxn[acc.fit.rxn$year==up.year & acc.fit.rxn$exp%in%c("ADA","ULL"),]
at.sub <- acc.fit.rxn[acc.fit.rxn$year==up.year & acc.fit.rxn$exp%in%c("ADA","RAT"),]
ru.sub <- acc.fit.rxn[acc.fit.rxn$year==up.year & acc.fit.rxn$exp%in%c("RAM","ULL"),]
rt.sub <- acc.fit.rxn[acc.fit.rxn$year==up.year & acc.fit.rxn$exp%in%c("RAM","RAT"),]
ar.sub <- acc.fit.rxn[acc.fit.rxn$year==up.year & acc.fit.rxn$exp%in%c("ADA","RAM"),]
ut.sub <- acc.fit.rxn[acc.fit.rxn$year==up.year & acc.fit.rxn$exp%in%c("ULL","RAT"),]


au.rn <- ggplot(au.sub, aes(site.complex,fb, group=id)) + geom_line(aes(colour=group.names)) + scale_color_manual(values=group.colors[c(1,2,3,4,7,9,8)],name = "admixture group") + xlab("Site_year") +ylab("fitness BLUPs") + scale_x_discrete(expand = c(0.05, 0.05)) + theme_light()

at.rn <- ggplot(at.sub, aes(site.complex,fb, group=id)) + geom_line(aes(colour=group.names)) + scale_color_manual(values=group.colors[c(1,2,3,4,7,9,8)],name = "admixture group") + xlab("Site_year") +ylab("fitness BLUPs") + scale_x_discrete(expand = c(0.05, 0.05)) + theme_light()

ru.rn <- ggplot(ru.sub, aes(site.complex,fb, group=id)) + geom_line(aes(colour=group.names)) + scale_color_manual(values=group.colors[c(1,2,3,4,7,9,8)],name = "admixture group") + xlab("Site_year") +ylab("fitness BLUPs") + scale_x_discrete(expand = c(0.05, 0.05)) + theme_light()

rt.rn <- ggplot(rt.sub, aes(site.complex,fb, group=id)) + geom_line(aes(colour=group.names)) + scale_color_manual(values=group.colors[c(1,2,3,4,7,9,8)],name = "admixture group") + xlab("Site_year") +ylab("fitness BLUPs") + scale_x_discrete(expand = c(0.05, 0.05)) + theme_light()

ar.rn <- ggplot(ar.sub, aes(site.complex,fb, group=id)) + geom_line(aes(colour=group.names)) + scale_color_manual(values=group.colors[c(1,2,3,4,7,9,8)],name = "admixture group") + xlab("Site_year") +ylab("fitness BLUPs") + scale_x_discrete(expand = c(0.05, 0.05)) + theme_light()

ut.rn <- ggplot(ut.sub, aes(site.complex,fb, group=id)) + geom_line(aes(colour=group.names)) + scale_color_manual(values=group.colors[c(1,2,3,4,7,9,8)],name = "admixture group") + xlab("Site_year") +ylab("fitness BLUPs") + scale_x_discrete(expand = c(0.05, 0.05)) + theme_light()

grid.arrange(au.rn, at.rn, ru.rn, rt.rn, ar.rn, ut.rn, nrow = 3)

pdf(file="./figures/2011.rxn.norms.pdf", width=9, height=11)
grid.arrange(au.rn, at.rn, ru.rn, rt.rn, ar.rn, ut.rn, nrow = 3)
dev.off()

```
### reaction norms between sites, 2012

```{r 2012 rxn norms}
up.year <- "2012"
au.sub <- acc.fit.rxn[acc.fit.rxn$year==up.year & acc.fit.rxn$exp%in%c("ADA","ULL"),]
at.sub <- acc.fit.rxn[acc.fit.rxn$year==up.year & acc.fit.rxn$exp%in%c("ADA","RAT"),]
ru.sub <- acc.fit.rxn[acc.fit.rxn$year==up.year & acc.fit.rxn$exp%in%c("RAM","ULL"),]
rt.sub <- acc.fit.rxn[acc.fit.rxn$year==up.year & acc.fit.rxn$exp%in%c("RAM","RAT"),]
ar.sub <- acc.fit.rxn[acc.fit.rxn$year==up.year & acc.fit.rxn$exp%in%c("ADA","RAM"),]
ut.sub <- acc.fit.rxn[acc.fit.rxn$year==up.year & acc.fit.rxn$exp%in%c("ULL","RAT"),]


au.rn <- ggplot(au.sub, aes(site.complex,fb, group=id)) + geom_line(aes(colour=group.names)) + scale_color_manual(values=group.colors[c(1,2,3,4,7,9,8)],name = "admixture group") + xlab("Site_year") +ylab("fitness BLUPs") + scale_x_discrete(expand = c(0.05, 0.05)) + theme_light()

at.rn <- ggplot(at.sub, aes(site.complex,fb, group=id)) + geom_line(aes(colour=group.names)) + scale_color_manual(values=group.colors[c(1,2,3,4,7,9,8)],name = "admixture group") + xlab("Site_year") +ylab("fitness BLUPs") + scale_x_discrete(expand = c(0.05, 0.05)) + theme_light()

ru.rn <- ggplot(ru.sub, aes(site.complex,fb, group=id)) + geom_line(aes(colour=group.names)) + scale_color_manual(values=group.colors[c(1,2,3,4,7,9,8)],name = "admixture group") + xlab("Site_year") +ylab("fitness BLUPs") + scale_x_discrete(expand = c(0.05, 0.05)) + theme_light()

rt.rn <- ggplot(rt.sub, aes(site.complex,fb, group=id)) + geom_line(aes(colour=group.names)) + scale_color_manual(values=group.colors[c(1,2,3,4,7,9,8)],name = "admixture group") + xlab("Site_year") +ylab("fitness BLUPs") + scale_x_discrete(expand = c(0.05, 0.05)) + theme_light()

ar.rn <- ggplot(ar.sub, aes(site.complex,fb, group=id)) + geom_line(aes(colour=group.names)) + scale_color_manual(values=group.colors[c(1,2,3,4,7,9,8)],name = "admixture group") + xlab("Site_year") +ylab("fitness BLUPs") + scale_x_discrete(expand = c(0.05, 0.05)) + theme_light()

ut.rn <- ggplot(ut.sub, aes(site.complex,fb, group=id)) + geom_line(aes(colour=group.names)) + scale_color_manual(values=group.colors[c(1,2,3,4,7,9,8)],name = "admixture group") + xlab("Site_year") +ylab("fitness BLUPs") + scale_x_discrete(expand = c(0.05, 0.05)) + theme_light()

grid.arrange(au.rn, at.rn, ru.rn, rt.rn, ar.rn, ut.rn, nrow = 3)

pdf(file="./figures/2012.rxn.norms.pdf", width=9, height=11)
grid.arrange(au.rn, at.rn, ru.rn, rt.rn, ar.rn, ut.rn, nrow = 3)
dev.off()

```

### patterns of significance and effect by line - per site and year
```{r sig pattern}
line.est <- int.mm7[int.mm7$groupFctr=="exp:year:id",] ### 2 of 1600 missing

## does CI contain zero?
line.est$ci.low <- line.est$mean-line.est$sd
line.est$ci.hi <- line.est$mean+line.est$sd
line.est$sig <- apply(line.est,1, function(x){(as.numeric(x[7])<0 & as.numeric(x[8])>0)==FALSE})
### 39% of CIs don't contain zero
line.est$mean.sig <- line.est$mean*line.est$sig

## reshape dataframe
tmp.ids <- do.call(rbind, sapply(line.est$groupID, function(x){strsplit(x,":")}))
tmp.ids <- data.frame(tmp.ids)
colnames(tmp.ids) <- c("site","year", "id")
line.est <- cbind(line.est, tmp.ids)
line.est$exp <- paste(line.est$site, line.est$year, sep="_")

test <- line.est[,c(10,13,14)]
test <- spread(test, exp, mean.sig)

write.table(line.est, file="./data/line.fitness.estimate.significance.txt")
```

