---
title: "31.fecundity.BLUPs"
author: "Daniele Filiault"
date: "2/27/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lme4) # mixed models
library(statmod) 
library(merTools)
library(MuMIn) # for calculating R2 of mixed models
library(RColorBrewer)
library(pander)
library(dplyr)
library(tidyr)
library(ggplot2)
library(corrplot)
library(heatmaply)
library(gridExtra)
library(pheatmap)
library(ggpubr)
library(pgirmess)
library(multcompView)
```

## Want to estimate BLUPs per accessions (i.e. use accession as a random effect in a mixed effects model)

### 1. prep data

```{r data prep}
d11=readRDS(file="./data/d11_for_analysis.rds")
d12=readRDS(file="./data/d12_for_analysis.rds")

##restrict to fitness
traits=c("fecundity")

d11$year=2011
d12$year=2012

d11=d11[, c("year", "exp", "block", "id", "fecundity")]
d12=d12[, c("year", "exp", "block", "id", "fecundity")]

d=rbind(d11, d12)
d$region[d$exp%in%c("ULL", "RAT")]="S"
d$region[d$exp%in%c("RAM", "ADA")]="N"
d$year=as.factor(d$year)
d=d[, c("year", "region", "exp","block", "id", "fecundity")]
d=na.omit(d)

hist(d$fecundity)
hist(sqrt(d$fecundity))
with(d, boxplot(fecundity~exp+year))
with(d, boxplot(sqrt(fecundity)~exp+year))


pdf(file="./figures/raw.fecundity.boxplot.by.year.site.pdf", width=10, height=6)
with(d, boxplot(sqrt(fecundity)~year+exp, col=c(rep(brewer.pal(9, "Paired")[4],4), rep(brewer.pal(9, "Paired")[3],4))))
dev.off()
```

### 2. fit "normal" linear models
```{r test normal linear models}

### do as normal lm first to understand specification, then move to glm so can use tweedie
lm1 <- lm(sqrt(fecundity)~id*year*exp, data=d)
lm3 <- lm(sqrt(fecundity)~id*year*exp + exp:year/block, data=d)
plot(lm3)
anova(lm1,lm3) ### yes, lm3 is much better.  The block helps. If it is specified correctly???
anova(lm3)

### test model fits by removing some components

lm4 <- lm(sqrt(fecundity)~id + year*exp + exp:year/block, data=d)
lm5 <- lm(sqrt(fecundity)~id, data=d)
lm6 <- lm(sqrt(fecundity)~year*exp + exp:year/block, data=d)
anova(lm3, lm4)
anova(lm3, lm5)
anova(lm3, lm6)

ssq <- anova(lm3)[[2]]
pve <-ssq/sum(ssq)
names(pve) <- rownames(anova(lm3))
print(pve)

#(summary(lm3))
pander(anova(lm3))

```
So the full model with triple interaction plus block is the best one (lm3). Plots are not the worst.

I'm not 100% sure that I'm specifying block correctly in model.  Make unique block names and compare?
Probably better choice is to specify block as a random effect in a mixed effect model.  I'm going that direction to estimate BLUPs anyway...

```{r mixed model specifications}
mm1 <- lmer(sqrt(fecundity) ~ 1 + (1 | block), data=d)
mm2 <- lmer(sqrt(fecundity) ~ 1 + (1 | id), data=d)
mm3 <- lmer(sqrt(fecundity)~exp + (1|exp:block), data=d) # so this is correctly specified (without year effect, at least)
mm4 <- lmer(sqrt(fecundity)~exp + year + (1|exp:year:block), data=d)  ## I think this is also correctly specified!
#ranef(mm4)
#fixef(mm4)
mm5 <- lmer(sqrt(fecundity)~exp*year + (1|exp:year:block), data=d)
anova(mm4, mm5) ## significant
#ranef(mm5)
#fixef(mm5)

mm6 <- lmer(sqrt(fecundity)~exp*year + (1|exp:year:block) + (1|id), data=d)
anova(mm5,mm6)

mm7 <- lmer(sqrt(fecundity)~exp*year + (1|exp:year:block) + (1|exp:year:id), data=d)
anova(mm6,mm7)

mm8 <- lmer(sqrt(fecundity)~exp*year  + (1|exp:year:id), data=d)


c2 <- predictInterval(mm7)
int.mm7 <- REsim(mm7)
plotREsim(REsim(mm7))

ranks <- expectedRank(mm7)

VarCorr(mm7)
anova(mm7)
summary(mm7)

# random effect variances
var.mm7 <- VarCorr(mm7)
print(var.mm7,comp=c("Variance","Std.Dev."),digits=4)
re.var <- as.data.frame(var.mm7, row.names = NULL,optional = FALSE)

r.squaredGLMM(mm5)
r.squaredGLMM(mm6)
r.squaredGLMM(mm7)
r.squaredGLMM(mm8)
## R2m is from marginal(fixed effects)
## R2c is from entire model
```

So, genotype explains about 10% of variance (as assessed by R2)

```{r explore blups}
blups <- ranef(mm7)[[1]]
blups <- as.data.frame(blups)
desc <- do.call(rbind,strsplit(rownames(blups),":"))
colnames(desc) <- c("exp","year","id")
blups <- cbind(blups, desc)
colnames(blups)[1] <- "fb"

hist(blups$fb)
pdf(file="./figures/blup.fecundity.boxplot.by.year.site.pdf", width=10, height=6)
with(blups, boxplot(fb~year+exp, col=c(rep(brewer.pal(9, "Paired")[4],4), rep(brewer.pal(9, "Paired")[3],4))))
dev.off()

## write BLUPS to file
write.csv(blups, file="./data/marginal.fecundity.blups.csv", quote=FALSE,row.names=FALSE)

```


So all this is the square root of blups.  Should un-transform it?  Or not?

```{r pairwise correlations between blups}
## need to reformat blups to long format
blups.l <- blups %>% pivot_wider(id_cols=id,names_from=c(exp,year),values_from=fb)%>%as.data.frame()

corrplot(cor(blups.l[,2:9], use="na.or.complete"))

my_cor <- cor(blups.l[,2:9], use="na.or.complete")
colfunc<-colorRampPalette(c("red","white","royalblue"))
heatmaply_cor(my_cor, colors= colfunc(100),cellnote=my_cor,cellnote_textposition="middle center")
### since these are interactive plots in plotly, they are hard to save statically, so I saved this one manually from the window
#file="./figures/correlation.dendrogram.fecundity.blups.jpeg")
```


Heatmap blups
```{r normalize blups}
hm.blups <- blups.l[,2:9]
rownames(hm.blups) <- blups.l[,1]

## z-score normalize the blups in each experiment
znorm <- function(x){(x-mean(x, na.rm=TRUE))/sd(x, na.rm=TRUE)}
norm.blups <- apply(hm.blups, 2,znorm)
```

```{r normalized snr fitness by K groups heatmap}
## similar to in 29.BLUP.heatmap.by.K
### K group membership
k.groups <- read.table(file="./data/Kmatrix.6cluster.membership.txt")

## put normalized blups into K clustering order
k.order.blups <- norm.blups[match(rownames(k.groups),rownames(norm.blups)),]

## make in pheatmap so can add breaks and annotations
bk <- c(seq(-max(k.order.blups,na.rm=TRUE),max(k.order.blups,na.rm=TRUE),by=0.5))
#colors (one less than breaks) - centered on 0, using grey as zero so can see breaks
mycols <- c(colorRampPalette(colors = c("red","grey80","blue"))(length(bk)-1))

### where should we place gaps?
gap.coord <- sapply(1:(nrow(k.groups)-1), function(x){
  k.groups[x,1]==k.groups[x+1,1]
  })
gap.coord <- which(gap.coord==FALSE)
gap.co <- c(2)

### add K-matrix group annotation
kcl.order.fact <- k.groups
ann.colors <- list(K.group=c(K1 = "#A6CEE3", K2 = "#1F78B4", K3 = "#B2DF8A", K4 = "#33A02C", K5 = "#FB9A99", K6 = "#E31A1C")) 
## these colors are from brewer.pal(6, "Paired").  Hard to generate this list automatically!

### make heatmap
pdf(file="./figures/heatmap.fecundity.by.Kgroup.pdf")
pheatmap(k.order.blups, cluster_rows=FALSE, color=mycols, breaks=bk, annotation_row=kcl.order.fact, gaps_row=gap.coord, cutree_cols=4, annotation_colors=ann.colors, show_rownames = FALSE)
dev.off()

pheatmap(k.order.blups, cluster_rows=FALSE, color=mycols, breaks=bk, annotation_row=kcl.order.fact, gaps_row=gap.coord, cutree_cols=4, annotation_colors=ann.colors, show_rownames = FALSE)

```
## 5.  Nonparametric tests of fecundity between K groups

```{R nonpara tests K group fecundity}

nb <- cbind(k.order.blups,k.groups)

### get labels for multiple testing
get.tuk.lab <- function(up.kmc){
  tlab <- up.kmc[[3]][,3]
  names(tlab) <- rownames(up.kmc[[3]])
  tlab <- data.frame(multcompLetters(tlab)['Letters'])
  return(tlab)
}

kw.comp <- as.list(rep(NA,8)) 
ph.comp <- as.list(rep(NA,8))

for(up in 1:8){
  up.exp <- colnames(nb)[up]
  up.kt <- kruskal.test(get(up.exp)~K.group, data=nb)
  up.ph <- kruskalmc(get(up.exp)~K.group, data=nb)
  out.ph <- get.tuk.lab(up.ph)
  kw.comp[[up]] <- up.kt
  ph.comp[[up]] <- out.ph
}  

ph.comp <- do.call(cbind,ph.comp)
colnames(ph.comp) <- colnames(nb)[1:8]
ph.comp$K.group <- rownames(ph.comp)
```

###6. nonparametric plots 2011

```{R nonparam plots 2011}
############
### ADA
max.vals <- aggregate(nb$ADA_2011,by=list(nb$K.group), max, na.rm=TRUE)
max.buffer <- data.frame(max.vals[,2] + (0.2*max(max.vals[,2])))
Tukey_test <- cbind(max.buffer,ph.comp$ADA_2011, max.vals[,1])
colnames(Tukey_test) <- c("ADA_2011","Tukey_labels","K.group")
tt.a1 <- Tukey_test

a1.plot <- ggplot(data=nb, aes(x=K.group, y=ADA_2011,fill=K.group)) +
  geom_boxplot() +
  geom_jitter(width=0.1) +
  scale_fill_brewer(palette="Paired") +
  theme_bw() +
  labs(y="normalized fecundity") +
  theme(legend.position = "none") +
  geom_text(data=tt.a1, aes(label=tt.a1$Tukey_labels)) + ggtitle("ADA_2011")

############
### RAM
max.vals <- aggregate(nb$RAM_2011,by=list(nb$K.group), max, na.rm=TRUE)
max.buffer <- data.frame(max.vals[,2] + (0.2*max(max.vals[,2])))
Tukey_test <- cbind(max.buffer,ph.comp$RAM_2011, max.vals[,1])
colnames(Tukey_test) <- c("RAM_2011","Tukey_labels","K.group")
tt.r1 <- Tukey_test

r1.plot <- ggplot(data=nb, aes(x=K.group, y=RAM_2011,fill=K.group)) +
  geom_boxplot() +
  geom_jitter(width=0.1) +
  scale_fill_brewer(palette="Paired") +
  theme_bw() +
  labs(y="normalized fecundity") +
  theme(legend.position = "none") +
  geom_text(data=tt.r1, aes(label=tt.r1$Tukey_labels)) + ggtitle("RAM_2011")

############
### ULL
max.vals <- aggregate(nb$ULL_2011,by=list(nb$K.group), max, na.rm=TRUE)
max.buffer <- data.frame(max.vals[,2] + (0.2*max(max.vals[,2])))
Tukey_test <- cbind(max.buffer,ph.comp$ULL_2011, max.vals[,1])
colnames(Tukey_test) <- c("ULL_2011","Tukey_labels","K.group")
tt.u1 <- Tukey_test

u1.plot <- ggplot(data=nb, aes(x=K.group, y=ULL_2011,fill=K.group)) +
  geom_boxplot() +
  geom_jitter(width=0.1) +
  scale_fill_brewer(palette="Paired") +
  theme_bw() +
  labs(y="normalized fecundity") +
  theme(legend.position = "none") +
  geom_text(data=tt.u1, aes(label=tt.u1$Tukey_labels)) + ggtitle("ULL_2011")

############
### RAT
max.vals <- aggregate(nb$RAT_2011,by=list(nb$K.group), max, na.rm=TRUE)
max.buffer <- data.frame(max.vals[,2] + (0.2*max(max.vals[,2])))
Tukey_test <- cbind(max.buffer,ph.comp$RAT_2011, max.vals[,1])
colnames(Tukey_test) <- c("RAT_2011","Tukey_labels","K.group")
tt.t1 <- Tukey_test

t1.plot <- ggplot(data=nb, aes(x=K.group, y=RAT_2011,fill=K.group)) +
  geom_boxplot() +
  geom_jitter(width=0.1) +
  scale_fill_brewer(palette="Paired") +
  theme_bw() +
  labs(y="normalized fecundity") +
  theme(legend.position = "none") +
  geom_text(data=tt.t1, aes(label=tt.t1$Tukey_labels)) + ggtitle("RAT_2011")
```

```{R nonparam plot 2011 generate}
pdf("./figures/fecundity.2011.nonparametric.by.K.pdf", width=9, height=9)
figure.nonparam <- ggarrange(a1.plot, r1.plot, u1.plot, t1.plot, 
          labels = c("A", "B", "C","D"),
          ncol = 2, nrow = 2)
annotate_figure(figure.nonparam, top = text_grob("normalized fecundity BLUPs - KW tests", face = "bold", size = 20))
dev.off()
```


## 7.  Nonparametric plots - 2012

```{R nonparam plots 2012}
############
### ADA
max.vals <- aggregate(nb$ADA_2012,by=list(nb$K.group), max, na.rm=TRUE)
max.buffer <- data.frame(max.vals[,2] + (0.2*max(max.vals[,2])))
Tukey_test <- cbind(max.buffer,ph.comp$ADA_2012, max.vals[,1])
colnames(Tukey_test) <- c("ADA_2012","Tukey_labels","K.group")
tt.a2 <- Tukey_test

a2.plot <- ggplot(data=nb, aes(x=K.group, y=ADA_2012,fill=K.group)) +
  geom_boxplot() +
  geom_jitter(width=0.1) +
  scale_fill_brewer(palette="Paired") +
  theme_bw() +
  labs(y="normalized fecundity") +
  theme(legend.position = "none") +
  geom_text(data=tt.a2, aes(label=tt.a2$Tukey_labels)) + ggtitle("ADA_2012")

############
### RAM
max.vals <- aggregate(nb$RAM_2012,by=list(nb$K.group), max, na.rm=TRUE)
max.buffer <- data.frame(max.vals[,2] + (0.2*max(max.vals[,2])))
Tukey_test <- cbind(max.buffer,ph.comp$RAM_2012, max.vals[,1])
colnames(Tukey_test) <- c("RAM_2012","Tukey_labels","K.group")
tt.r2 <- Tukey_test

r2.plot <- ggplot(data=nb, aes(x=K.group, y=RAM_2012,fill=K.group)) +
  geom_boxplot() +
  geom_jitter(width=0.1) +
  scale_fill_brewer(palette="Paired") +
  theme_bw() +
  labs(y="normalized fecundity") +
  theme(legend.position = "none") +
  geom_text(data=tt.r2, aes(label=tt.r2$Tukey_labels)) + ggtitle("RAM_2012")

############
### ULL
max.vals <- aggregate(nb$ULL_2012,by=list(nb$K.group), max, na.rm=TRUE)
max.buffer <- data.frame(max.vals[,2] + (0.2*max(max.vals[,2])))
Tukey_test <- cbind(max.buffer,ph.comp$ULL_2012, max.vals[,1])
colnames(Tukey_test) <- c("ULL_2012","Tukey_labels","K.group")
tt.u2 <- Tukey_test

u2.plot <- ggplot(data=nb, aes(x=K.group, y=ULL_2012,fill=K.group)) +
  geom_boxplot() +
  geom_jitter(width=0.1) +
  scale_fill_brewer(palette="Paired") +
  theme_bw() +
  labs(y="normalized fecundity") +
  theme(legend.position = "none") +
  geom_text(data=tt.u2, aes(label=tt.u2$Tukey_labels)) + ggtitle("ULL_2012")

############
### RAT
max.vals <- aggregate(nb$RAT_2012,by=list(nb$K.group), max, na.rm=TRUE)
max.buffer <- data.frame(max.vals[,2] + (0.2*max(max.vals[,2])))
Tukey_test <- cbind(max.buffer,ph.comp$RAT_2012, max.vals[,1])
colnames(Tukey_test) <- c("RAT_2012","Tukey_labels","K.group")
tt.t2 <- Tukey_test

t2.plot <- ggplot(data=nb, aes(x=K.group, y=RAT_2012,fill=K.group)) +
  geom_boxplot() +
  geom_jitter(width=0.1) +
  scale_fill_brewer(palette="Paired") +
  theme_bw() +
  labs(y="normalized fecundity") +
  theme(legend.position = "none") +
  geom_text(data=tt.t2, aes(label=tt.t2$Tukey_labels)) + ggtitle("RAT_2012")
```

```{R nonparam plot 2012 generate}
pdf("./figures/fecundity.2012.nonparametric.by.K.pdf", width=9, height=9)
figure.nonparam <- ggarrange(a2.plot, r2.plot, u2.plot, t2.plot, 
          labels = c("A", "B", "C","D"),
          ncol = 2, nrow = 2)
annotate_figure(figure.nonparam, top = text_grob("normalized fecundity BLUPs - KW tests", face = "bold", size = 20))
dev.off()
```