---
title: "32.ows.BLUPS"
author: "Daniele Filiault"
date: "2/28/2020, revisited 02Dec20"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lme4) # mixed models
library(statmod) 
library(merTools)
library(MuMIn) # for calculating R2 of mixed models
library(RColorBrewer)
library(pander)
library(dplyr)
library(tidyr)
library(ggplot2)
library(corrplot)
library(heatmaply)
library(gridExtra)
library(pheatmap)
library(ggpubr)
library(pgirmess)
library(multcompView)
library(multcomp)

#setwd("/groups/nordborg/projects/field_experiments/adaptation_sweden/common.gardens")
setwd("/Volumes/field_experiments/adaptation_sweden/common.gardens/")

```

## Want to estimate BLUPs per accessions (i.e. use accession as a random effect in a mixed effects model)

### 1. prep data

```{r data prep}
d11=readRDS(file="./data/d11_for_analysis.rds")
d12=readRDS(file="./data/d12_for_analysis.rds")

##restrict to fitness
traits=c("ows")

d11$year=2011
d12$year=2012

d11=d11[, c("year", "exp", "block", "id", "ows")]
d12=d12[, c("year", "exp", "block", "id", "ows")]

d=rbind(d11, d12)
d$region[d$exp%in%c("ULL", "RAT")]="S"
d$region[d$exp%in%c("RAM", "ADA")]="N"
d$year=as.factor(d$year)
d=d[, c("year", "region", "exp","block", "id", "ows")]
d=na.omit(d)
d$ows <- as.factor(d$ows)
#add experiment short names
d$s.name <- d$exp
d$s.name <- gsub("ULL","SU",d$s.name)
d$s.name <- gsub("RAT","SR",d$s.name)
d$s.name <- gsub("RAM","NM",d$s.name)
d$s.name <- gsub("ADA","NA",d$s.name)
d$s.name <- factor(d$s.name, levels=c("NA","NM","SU","SR"))

with(d, table(ows,year, exp))
raw.ows.plot <- ggplot(data=d, aes(x=year, fill=ows)) +
  geom_bar() +
  facet_wrap(~s.name) +
  ylab("number of samples") +
  labs(fill = "winter \nsurvival")

pdf(file="./figures/32.figures/raw.ows.barplot.by.year.site.pdf", width=5, height=6)
print(raw.ows.plot)
dev.off()

# also do means by line
aggdata <-aggregate(ows~exp+block+id+year, FUN=function(x){mean(as.numeric(as.character(x)))},data=d)
ows.m <- aggregate(ows~exp+year+id, FUN=mean, data=aggdata)
ows.m$s.name <- ows.m$exp
ows.m$s.name <- gsub("ULL","SU",ows.m$s.name)
ows.m$s.name <- gsub("RAT","SR",ows.m$s.name)
ows.m$s.name <- gsub("RAM","NM",ows.m$s.name)
ows.m$s.name <- gsub("ADA","NA",ows.m$s.name)
ows.m$s.name <- factor(ows.m$s.name, levels=c("NA","NM","SU","SR"))
ows.m$region[ows.m$exp%in%c("ULL", "RAT")]="S"
ows.m$region[ows.m$exp%in%c("RAM", "ADA")]="N"

line.ows.plot <- ggplot(data=ows.m, aes(x=s.name, y=ows, fill=region)) +
  geom_boxplot() +
  scale_fill_manual(values=c("dodgerblue1","springgreen2")) +
  facet_wrap(~year) +
  ylab("mean overwinter survival per line") +
  xlab("experiment") +
  labs(fill = "experiment \nregion")
  #theme(strip.background=element_rect(fill=c("goldenrod1","magenta3")))

p <- line.ows.plot
g <- ggplot_gtable(ggplot_build(p))
striprt <- which( grepl('strip-r', g$layout$name) | grepl('strip-t', g$layout$name) )
fills <- c("goldenrod1","magenta3")
k <- 1
for (i in striprt) {
  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
  k <- k+1
}
grid::grid.draw(g)

pdf(file="./figures/32.figures/line.ows.barplot.by.year.site.pdf", width=7, height=5)
grid::grid.draw(g)
dev.off()

```

### 2. fit "normal" linear models - this doesn't work since interaction models are so slow
```{r test normal linear models, eval=FALSE}

### do as normal lm first to understand specification

lm1 <- glm(ows~id*year*exp, data=d, family=binomial("logit"))
lm3 <- glm(ows~id*year*exp + exp:year/block, data=d, family=binomial("logit"))
plot(lm3)
anova(lm1,lm3) 
anova(lm3)

### test model fits by removing some components

lm4 <- glm(ows~id + year*exp + exp:year/block, data=d, family="binomial")
lm5 <- glm(ows~id, data=d, family=binomial("logit"))
lm6 <- glm(ows~year*exp + exp:year/block, data=d, family="binomial")
anova(lm3, lm4)
anova(lm3, lm5)
anova(lm3, lm6)

ssq <- anova(lm3)[[2]]
pve <-ssq/sum(ssq)
names(pve) <- rownames(anova(lm3))
print(pve)

#(summary(lm3))
pander(anova(lm3))

```


### 3. Fit mixed models
```{r mixed model specifications}
### was having convergence issues.  Setting to the other optimizer and increasing iterations seems to fix.
mm1 <- glmer(ows ~ 1 + (1 | block), data=d,family=binomial("logit"),control=glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5))) 

mm2 <- glmer(ows ~ 1 + (1 | id), data=d, family=binomial("logit"),control=glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)))

mm3 <- glmer(ows~exp + (1|exp:block), data=d, family=binomial("logit"),control=glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5))) # so this is correctly specified (without year effect, at least)

mm4 <- glmer(ows~exp + year + (1|exp:year:block), data=d, family=binomial("logit"),control=glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)))  ## I think this is also correctly specified!

mm5 <- glmer(ows~exp*year + (1|exp:year:block), data=d, family=binomial("logit"), control=glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)))

mm6 <- glmer(ows~exp*year + (1|exp:year:block) + (1|id), data=d,family=binomial("logit"),control=glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)))

mm7 <- glmer(ows~exp*year + (1|exp:year:block) + (1|exp:year:id), data=d, family=binomial("logit"),control=glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)))

mm8 <- glmer(ows~exp*year  + (1|exp:year:id), data=d, family=binomial("logit"), control=glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)))

# get fit information for all these models
mods <- c("mm1", "mm2", "mm3","mm4", "mm5", "mm6", "mm7", "mm8")
mod.sum <- matrix(NA, ncol=5, nrow=length(mods))
for(up in 1:length(mods)){
  up.m <- mods[up]
  up.lm <- get(up.m)
  up.fit <- summary(up.lm)$AICtab
  mod.sum[up,] <- up.fit
}
rownames(mod.sum) <- mods
colnames(mod.sum) <- names(up.fit)
print(mod.sum)


c2 <- predictInterval(mm7)
int.mm7 <- REsim(mm7)
plotREsim(REsim(mm7))

anova(mm7)
summary(mm7)

# random effect variances
var.mm7 <- VarCorr(mm7)
print(var.mm7,comp=c("Variance","Std.Dev."),digits=4)
re.var <- as.data.frame(var.mm7, row.names = NULL,optional = FALSE)

r.squaredGLMM(mm5)
r.squaredGLMM(mm6)
r.squaredGLMM(mm7)
r.squaredGLMM(mm8)
## R2m is from marginal(fixed effects)
## R2c is from entire model
```

So, comparing model 7 versus 5, genotype explains 11.51327%. (block is only 0.3)  fixed effects are 7.25%


```{r explore blups}
blups <- ranef(mm7)[[1]]
blups <- as.data.frame(blups)
desc <- do.call(rbind,strsplit(rownames(blups),":"))
colnames(desc) <- c("exp","year","id")
blups <- cbind(blups, desc)
colnames(blups)[1] <- "ows.b"
blups$s.name <- blups$exp
blups$s.name <- gsub("ULL","SU",blups$s.name)
blups$s.name <- gsub("RAT","SR",blups$s.name)
blups$s.name <- gsub("RAM","NM",blups$s.name)
blups$s.name <- gsub("ADA","NA",blups$s.name)
blups$s.name <- factor(blups$s.name, levels=c("NA","NM","SU","SR"))
blups$region[blups$exp%in%c("ULL", "RAT")]="S"
blups$region[blups$exp%in%c("RAM", "ADA")]="N"

hist(blups$ows.b)

### BLUPS are the log odds probability of survival.  Make probability

logit2prob <- function(logit){
  odds <- exp(logit)
  prob <- odds / (1 + odds)
  return(prob)
}

blups$ows.p <- logit2prob(blups$ows.b)
with(blups, boxplot(ows.p~year+exp, col=c(rep(brewer.pal(9, "Paired")[4],4), rep(brewer.pal(9, "Paired")[3],4)), ylab="BLUP (logit ows)"))

ows.blup.plot <- ggplot(data=blups, aes(x=s.name, y=ows.b, fill=region)) +
  geom_boxplot() +
  scale_fill_manual(values=c("dodgerblue1","springgreen2")) +
  facet_wrap(~year) +
  ylab("overwinter survival BLUPs (logit)") +
  xlab("experiment") +
  labs(fill = "experiment \nregion")

pdf(file="./figures/32.figures/blups.ows.boxplot.by.year.site.pdf", width=7, height=5)
print(ows.blup.plot)
dev.off()

ows.prob.plot <- ggplot(data=blups, aes(x=s.name, y=ows.p, fill=region)) +
  geom_boxplot() +
  scale_fill_manual(values=c("dodgerblue1","springgreen2")) +
  facet_wrap(~year) +
  ylab("overwinter survival BLUPs (probability)") +
  xlab("experiment") +
  labs(fill = "experiment \nregion")

pdf(file="./figures/32.figures/prob.ows.boxplot.by.year.site.pdf", width=7, height=5)
print(ows.prob.plot)
dev.off()

## write BLUPS to file
write.csv(blups, file="./data/marginal.ows.blups.csv", quote=FALSE,row.names=FALSE)

## look at correlations between actual means and blups
comp <- merge(blups, ows.m, by=c("exp", "year", "id","s.name", "region"), all=TRUE)
comp$year <- as.factor(comp$year)

pdf("./figures/32.figures/blup.vs.mean.ows.pdf", width=6, height=5)
comp.plot <- ggplot(data=comp, aes(x=ows, y=ows.b, color=s.name)) +
  geom_point(aes(shape=year)) +
  xlab("mean overwinter survival") +
  ylab("BLUP overwinter survival")
print(comp.plot)
dev.off()

comp$ows.l <- exp(comp$ows.b)

## write full data to file
write.csv(comp, file="./data/ows.blups.and.means.csv", quote=FALSE, row.names=FALSE)

```


So blups from logistic regression are log odds probability of ows. 
The thing is that this doesn't scale linearly with raw mean ows and it accentuates mean ows differences. 
So what to use when calculating SnR composite fitness???
Question - what if I just modeled 2012 ows?
In the meantime, work through the rest of this with the BLUPs

```{r pairwise correlations between blups}
## need to reformat blups to long format
blups <- read.csv(file="./data/marginal.ows.blups.csv")
blups.l <- blups %>% pivot_wider(id_cols=id,names_from=c(s.name,year),values_from=ows.b)%>%as.data.frame()
#also transform mean values
means.l <- comp %>% pivot_wider(id_cols=id,names_from=c(s.name,year),values_from=ows)%>%as.data.frame()

corrplot(cor(blups.l[,2:9], use="na.or.complete"))

my_cor <- cor(blups.l[,2:9], use="na.or.complete")

reg.annot <- as.data.frame(c(rep("N",4), rep("S",4)))
colnames(reg.annot) <- "region"
rownames(reg.annot) <- rownames(my_cor)
reg.annot$year <- rep(c(2011,2012),4)
reg.annot$year <- as.factor(reg.annot$year)

region.col <- c("dodgerblue1","springgreen2")
year.col <- c("goldenrod1","magenta3")
ann_colors = list(
    region = c(N=region.col[2], S=region.col[1]), 
    year = c("2011"=year.col[1], "2012"=year.col[2])
)

means_cor <- cor(means.l[,2:9], use="na.or.complete")

pdf(file="./figures/32.figures/owsBLUPs.correlations.pdf", width=7, height=6)
pheatmap(my_cor,annotation_col = reg.annot, annotation_colors = ann_colors,color = colorRampPalette((brewer.pal(n = 7, name =
  "BuPu")))(100), main="Correlation of ows BLUPs by experiment/year", display_numbers = T)
dev.off()

pdf(file="./figures/32.figures/ows.means.correlations.pdf", width=7, height=6)
pheatmap(means_cor,annotation_col = reg.annot, annotation_colors = ann_colors,color = colorRampPalette((brewer.pal(n = 7, name =
  "BuPu")))(100), main="Correlation of line means by experiment/year", display_numbers =T)
dev.off()

```


### 4. Heatmap blups
```{r normalize blups}
hm.blups <- blups.l[,2:9]
rownames(hm.blups) <- blups.l[,1]

## z-score normalize the blups in each experiment
znorm <- function(x){(x-mean(x, na.rm=TRUE))/sd(x, na.rm=TRUE)}
norm.blups <- apply(hm.blups, 2,znorm)
```

```{r normalized ows by K groups heatmap}
## similar to in 29.BLUP.heatmap.by.K
### K group membership
k.groups <- read.table(file="./data/Kmatrix.6cluster.membership.txt")

## put normalized blups into K clustering order
k.order.blups <- norm.blups[match(rownames(k.groups),rownames(norm.blups)),]

## make in pheatmap so can add breaks and annotations
bk <- c(seq(min(k.order.blups,na.rm=TRUE),-min(k.order.blups,na.rm=TRUE),by=0.5))
#colors (one less than breaks) - centered on 0, using grey as zero so can see breaks
mycols <- c(colorRampPalette(colors = c("red","grey80","blue"))(length(bk)-1))

### where should we place gaps?
gap.coord <- sapply(1:(nrow(k.groups)-1), function(x){
  k.groups[x,1]==k.groups[x+1,1]
  })
gap.coord <- which(gap.coord==FALSE)
gap.co <- c(2)

### add K-matrix group annotation
kcl.order.fact <- k.groups
ann.colors <- list(K.name=c(B = "#E31A1C", S1 = "#A6CEE3", S2 = "#1F78B4", C = "#FB9A99", N1 = "#B2DF8A", N2 = "#33A02C")) 
## these colors are from brewer.pal(6, "Paired").  Hard to generate this list automatically!

### make heatmap
pdf(file="./figures/32.figures/heatmap.ows.blups.by.Kgroup.pdf")
pheatmap(k.order.blups, cluster_rows=FALSE, color=mycols, breaks=bk, annotation_row=kcl.order.fact, gaps_row=gap.coord, cutree_cols=3, annotation_colors=ann.colors, show_rownames = FALSE)
dev.off()

pheatmap(k.order.blups, cluster_rows=FALSE, color=mycols, breaks=bk, annotation_row=kcl.order.fact, gaps_row=gap.coord, cutree_cols=3, annotation_colors=ann.colors, show_rownames = FALSE)

### also make heatmap for non-normalized mean data
k.order.means <- means.l[match(rownames(k.groups),means.l$id),]
rownames(k.order.means) <- k.order.means$id
k.order.means <- k.order.means[,-1]
## make in pheatmap so can add breaks and annotations
bk <- c(seq(0,1,by=0.01))
#colors (one less than breaks) - centered on 0, using grey as zero so can see breaks
mycols <- c(colorRampPalette(colors = c("purple","grey80"))(length(bk)-1))
pdf(file="./figures/32.figures/heatmap.ows.means.by.Kgroup.pdf")
pheatmap(k.order.means, cluster_rows=FALSE, color=mycols, breaks=bk, annotation_row=kcl.order.fact, gaps_row=gap.coord, annotation_colors=ann.colors, show_rownames = FALSE)
dev.off()

```

### 5. Heatmap means
```{r normalize means}
hm.means <- means.l[,2:9]
rownames(hm.means) <- means.l[,1]

## z-score normalize the blups in each experiment
znorm <- function(x){(x-mean(x, na.rm=TRUE))/sd(x, na.rm=TRUE)}
norm.means <- apply(hm.means, 2,znorm)
```

```{r normalized ows means by K groups heatmap}
## similar to in 29.BLUP.heatmap.by.K
### K group membership
k.groups <- read.table(file="./data/Kmatrix.6cluster.membership.txt")

## put normalized means into K clustering order
k.order.means <- norm.means[match(rownames(k.groups),rownames(norm.means)),]

## make in pheatmap so can add breaks and annotations
bk <- c(seq(-max(abs(k.order.means),na.rm=TRUE),max(abs(k.order.means),na.rm=TRUE),by=0.5))
#colors (one less than breaks) - centered on 0, using grey as zero so can see breaks
mycols <- c(colorRampPalette(colors = c("red","grey80","blue"))(length(bk)-1))

### where should we place gaps?
gap.coord <- sapply(1:(nrow(k.groups)-1), function(x){
  k.groups[x,1]==k.groups[x+1,1]
  })
gap.coord <- which(gap.coord==FALSE)
gap.co <- c(2)

### add K-matrix group annotation
kcl.order.fact <- k.groups
ann.colors <- list(K.name=c(B = "#E31A1C", S1 = "#A6CEE3", S2 = "#1F78B4", C = "#FB9A99", N1 = "#B2DF8A", N2 = "#33A02C")) 
## these colors are from brewer.pal(6, "Paired").  Hard to generate this list automatically!

### make heatmap
pdf(file="./figures/32.figures/heatmap.means.normalized.by.Kgroup.pdf")
pheatmap(k.order.means, cluster_rows=FALSE, color=mycols, breaks=bk, annotation_row=kcl.order.fact, gaps_row=gap.coord, cutree_cols=4, annotation_colors=ann.colors, show_rownames = FALSE)
dev.off()

pheatmap(k.order.means, cluster_rows=FALSE, color=mycols, breaks=bk, annotation_row=kcl.order.fact, gaps_row=gap.coord, cutree_cols=4, annotation_colors=ann.colors, show_rownames = FALSE)

```



## 6.  Nonparametric tests of ows between K groups 2011

```{R nonpara tests K group SnR}
#k.order.sfit

nb <- cbind(k.order.blups,k.groups)
colnames(nb)[9] <- "K.group"
nb$K.group <- factor(nb$K.group, levels=c("B","S1","S2","C","N1", "N2"))

### get labels for multiple testing
get.tuk.lab <- function(up.kmc){
  tlab <- up.kmc[[3]][,3]
  names(tlab) <- rownames(up.kmc[[3]])
  tlab <- data.frame(multcompLetters(tlab)['Letters'])
  return(tlab)
}

kw.comp <- as.list(rep(NA,8)) 
ph.comp <- as.list(rep(NA,8))

for(up in 1:8){
  up.exp <- colnames(nb)[up]
  up.kt <- kruskal.test(get(up.exp)~K.group, data=nb)
  up.ph <- kruskalmc(get(up.exp)~K.group, data=nb)
  out.ph <- get.tuk.lab(up.ph)
  kw.comp[[up]] <- up.kt
  ph.comp[[up]] <- out.ph
}  

ph.comp <- do.call(cbind,ph.comp)
colnames(ph.comp) <- colnames(nb)[1:8]
ph.comp$K.group <- rownames(ph.comp)
```


```{R nonparam plots 2011}
############
### ADA
max.vals <- aggregate(nb$NA_2011,by=list(nb$K.group), max, na.rm=TRUE)
max.buffer <- data.frame(max.vals[,2] + (0.2*max(max.vals[,2])))
Tukey_test <- cbind(max.buffer,ph.comp$NA_2011, max.vals[,1])
colnames(Tukey_test) <- c("NA_2011","Tukey_labels","K.group")
tt.a1 <- Tukey_test

kcol <- brewer.pal(6, "Paired")[c(6,1,2,5,3,4)]

a1.plot <- ggplot(data=nb, aes(x=K.group, y=NA_2011,fill=K.group)) +
  geom_boxplot() +
  geom_jitter(width=0.1) +
  scale_fill_manual(values=kcol) +
  theme_bw() +
  labs(y="normalized ows") +
  theme(legend.position = "none") +
  geom_text(data=tt.a1, aes(label=Tukey_labels)) + ggtitle("NA_2011")

############
### RAM
max.vals <- aggregate(nb$NM_2011,by=list(nb$K.group), max, na.rm=TRUE)
max.buffer <- data.frame(max.vals[,2] + (0.2*max(max.vals[,2])))
Tukey_test <- cbind(max.buffer,ph.comp$NM_2011, max.vals[,1])
colnames(Tukey_test) <- c("NM_2011","Tukey_labels","K.group")
tt.r1 <- Tukey_test

r1.plot <- ggplot(data=nb, aes(x=K.group, y=NM_2011,fill=K.group)) +
  geom_boxplot() +
  geom_jitter(width=0.1) +
  scale_fill_manual(values=kcol) +
  theme_bw() +
  labs(y="normalized ows") +
  theme(legend.position = "none") +
  geom_text(data=tt.r1, aes(label=Tukey_labels)) + ggtitle("NM_2011")

############
### ULL
max.vals <- aggregate(nb$SU_2011,by=list(nb$K.group), max, na.rm=TRUE)
max.buffer <- data.frame(max.vals[,2] + (0.2*max(max.vals[,2])))
Tukey_test <- cbind(max.buffer,ph.comp$SU_2011, max.vals[,1])
colnames(Tukey_test) <- c("SU_2011","Tukey_labels","K.group")
tt.u1 <- Tukey_test

u1.plot <- ggplot(data=nb, aes(x=K.group, y=SU_2011,fill=K.group)) +
  geom_boxplot() +
  geom_jitter(width=0.1) +
  scale_fill_manual(values=kcol) +
  theme_bw() +
  labs(y="normalized ows") +
  theme(legend.position = "none") +
  geom_text(data=tt.u1, aes(label=Tukey_labels)) + ggtitle("SU_2011")

############
### RAT
max.vals <- aggregate(nb$SR_2011,by=list(nb$K.group), max, na.rm=TRUE)
max.buffer <- data.frame(max.vals[,2] + (0.2*max(max.vals[,2])))
Tukey_test <- cbind(max.buffer,ph.comp$SR_2011, max.vals[,1])
colnames(Tukey_test) <- c("SR_2011","Tukey_labels","K.group")
tt.t1 <- Tukey_test

t1.plot <- ggplot(data=nb, aes(x=K.group, y=SR_2011,fill=K.group)) +
  geom_boxplot() +
  geom_jitter(width=0.1) +
  scale_fill_manual(values=kcol) +
  theme_bw() +
  labs(y="normalized ows") +
  theme(legend.position = "none") +
  geom_text(data=tt.t1, aes(label=Tukey_labels)) + ggtitle("SR_2011")
```

```{R nonparam plot 2011 generate}
pdf("./figures/32.figures/ows.blups.2011.nonparametric.by.K.pdf", width=9, height=9)
figure.nonparam <- ggarrange(a1.plot, r1.plot, u1.plot, t1.plot, 
          labels = c("A", "B", "C","D"),
          ncol = 2, nrow = 2)
annotate_figure(figure.nonparam, top = text_grob("normalized overwinter survival BLUPs - KW tests", face = "bold", size = 20))
dev.off()
```
## 7.  Nonparametric tests of ows between K groups - 2012

```{R nonparam plots 2012}
############
### ADA
max.vals <- aggregate(nb$NA_2012,by=list(nb$K.group), max, na.rm=TRUE)
max.buffer <- data.frame(max.vals[,2] + (0.2*max(max.vals[,2])))
Tukey_test <- cbind(max.buffer,ph.comp$NA_2012, max.vals[,1])
colnames(Tukey_test) <- c("NA_2012","Tukey_labels","K.group")
tt.a2 <- Tukey_test

a2.plot <- ggplot(data=nb, aes(x=K.group, y=NA_2012,fill=K.group)) +
  geom_boxplot() +
  geom_jitter(width=0.1) +
  scale_fill_manual(values=kcol) +
  theme_bw() +
  labs(y="normalized ows") +
  theme(legend.position = "none") +
  geom_text(data=tt.a2, aes(label=Tukey_labels)) + ggtitle("NA_2012")

############
### RAM
max.vals <- aggregate(nb$NM_2012,by=list(nb$K.group), max, na.rm=TRUE)
max.buffer <- data.frame(max.vals[,2] + (0.2*max(max.vals[,2])))
Tukey_test <- cbind(max.buffer,ph.comp$NM_2012, max.vals[,1])
colnames(Tukey_test) <- c("NM_2012","Tukey_labels","K.group")
tt.r2 <- Tukey_test

r2.plot <- ggplot(data=nb, aes(x=K.group, y=NM_2012,fill=K.group)) +
  geom_boxplot() +
  geom_jitter(width=0.1) +
  scale_fill_manual(values=kcol) +
  theme_bw() +
  labs(y="normalized ows") +
  theme(legend.position = "none") +
  geom_text(data=tt.r2, aes(label=Tukey_labels)) + ggtitle("NM_2012")

############
### ULL
max.vals <- aggregate(nb$SU_2012,by=list(nb$K.group), max, na.rm=TRUE)
max.buffer <- data.frame(max.vals[,2] + (0.2*max(max.vals[,2])))
Tukey_test <- cbind(max.buffer,ph.comp$SU_2012, max.vals[,1])
colnames(Tukey_test) <- c("SU_2012","Tukey_labels","K.group")
tt.u2 <- Tukey_test

u2.plot <- ggplot(data=nb, aes(x=K.group, y=SU_2012,fill=K.group)) +
  geom_boxplot() +
  geom_jitter(width=0.1) +
  scale_fill_manual(values=kcol) +
  theme_bw() +
  labs(y="normalized ows") +
  theme(legend.position = "none") +
  geom_text(data=tt.u2, aes(label=Tukey_labels)) + ggtitle("SU_2012")

############
### RAT
max.vals <- aggregate(nb$SR_2012,by=list(nb$K.group), max, na.rm=TRUE)
max.buffer <- data.frame(max.vals[,2] + (0.2*max(max.vals[,2])))
Tukey_test <- cbind(max.buffer,ph.comp$SR_2012, max.vals[,1])
colnames(Tukey_test) <- c("SR_2012","Tukey_labels","K.group")
tt.t2 <- Tukey_test

t2.plot <- ggplot(data=nb, aes(x=K.group, y=SR_2012,fill=K.group)) +
  geom_boxplot() +
  geom_jitter(width=0.1) +
  scale_fill_manual(values=kcol) +
  theme_bw() +
  labs(y="normalized ows") +
  theme(legend.position = "none") +
  geom_text(data=tt.t2, aes(label=Tukey_labels)) + ggtitle("SR_2012")
```

```{R nonparam plot 2012 generate}
pdf("./figures/32.figures/ows.blups.2012.nonparametric.by.K.pdf", width=9, height=9)
figure.nonparam <- ggarrange(a2.plot, r2.plot, u2.plot, t2.plot, 
          labels = c("A", "B", "C","D"),
          ncol = 2, nrow = 2)
annotate_figure(figure.nonparam, top = text_grob("normalized overwinter survival BLUPs - KW tests", face = "bold", size = 20))
dev.off()
```

```{R nonparam plot both years generate}
pdf("./figures/32.figures/normalized.ows.blups.nonparametric.both.years.by.K.pdf", width=18, height=12)
figure.nonparam <- ggarrange(a1.plot, r1.plot, u1.plot, t1.plot,a2.plot, r2.plot, u2.plot, t2.plot, 
          labels = c("A", "B", "C","D","E","F","G","H"),
          ncol = 4, nrow = 2)
annotate_figure(figure.nonparam, top = text_grob("normalized ows BLUPs - KW tests", face = "bold", size = 20))
dev.off()
```

## 8.  Nonparametric tests of mean ows between K groups 

```{R nonpara tests K group mean ows}
## don't normalize these means first

k.order.nnm <- hm.means[match(rownames(k.groups),rownames(hm.means)),]

nb <- cbind(k.order.nnm,k.groups)
colnames(nb)[9] <- "K.group"
nb$K.group <- factor(nb$K.group, levels=c("B","S1","S2","C","N1", "N2"))

### get labels for multiple testing
get.tuk.lab <- function(up.kmc){
  tlab <- up.kmc[[3]][,3]
  names(tlab) <- rownames(up.kmc[[3]])
  tlab <- data.frame(multcompLetters(tlab)['Letters'])
  return(tlab)
}

kw.comp <- as.list(rep(NA,8)) 
ph.comp <- as.list(rep(NA,8))

for(up in 1:8){
  up.exp <- colnames(nb)[up]
  up.kt <- kruskal.test(get(up.exp)~K.group, data=nb)
  up.ph <- kruskalmc(get(up.exp)~K.group, data=nb)
  out.ph <- get.tuk.lab(up.ph)
  kw.comp[[up]] <- up.kt
  ph.comp[[up]] <- out.ph
}  

ph.comp <- do.call(cbind,ph.comp)
colnames(ph.comp) <- colnames(nb)[1:8]
ph.comp$K.group <- rownames(ph.comp)
```


```{R nonparam means plots 2011}
############
### ADA
max.vals <- aggregate(nb$NA_2011,by=list(nb$K.group), max, na.rm=TRUE)
max.buffer <- data.frame(max.vals[,2] + (0.2*max(max.vals[,2])))
Tukey_test <- cbind(max.buffer,ph.comp$NA_2011, max.vals[,1])
colnames(Tukey_test) <- c("NA_2011","Tukey_labels","K.group")
tt.a1 <- Tukey_test

a1.plot <- ggplot(data=nb, aes(x=K.group, y=NA_2011,fill=K.group)) +
  geom_boxplot() +
  geom_jitter(width=0.1) +
  scale_fill_manual(values=kcol) +
  theme_bw() +
  labs(y="overwinter survival") +
  theme(legend.position = "none") +
  geom_text(data=tt.a1, aes(label=Tukey_labels)) + ggtitle("NA_2011")

############
### RAM
max.vals <- aggregate(nb$NM_2011,by=list(nb$K.group), max, na.rm=TRUE)
max.buffer <- data.frame(max.vals[,2] + (0.2*max(max.vals[,2])))
Tukey_test <- cbind(max.buffer,ph.comp$NM_2011, max.vals[,1])
colnames(Tukey_test) <- c("NM_2011","Tukey_labels","K.group")
tt.r1 <- Tukey_test

r1.plot <- ggplot(data=nb, aes(x=K.group, y=NM_2011,fill=K.group)) +
  geom_boxplot() +
  geom_jitter(width=0.1) +
  scale_fill_manual(values=kcol) +
  theme_bw() +
  labs(y="overwinter survival") +
  theme(legend.position = "none") +
  geom_text(data=tt.r1, aes(label=Tukey_labels)) + ggtitle("NM_2011")

############
### ULL
max.vals <- aggregate(nb$SU_2011,by=list(nb$K.group), max, na.rm=TRUE)
max.buffer <- data.frame(max.vals[,2] + (0.2*max(max.vals[,2])))
Tukey_test <- cbind(max.buffer,ph.comp$SU_2011, max.vals[,1])
colnames(Tukey_test) <- c("SU_2011","Tukey_labels","K.group")
tt.u1 <- Tukey_test

u1.plot <- ggplot(data=nb, aes(x=K.group, y=SU_2011,fill=K.group)) +
  geom_boxplot() +
  geom_jitter(width=0.1) +
  scale_fill_manual(values=kcol) +
  theme_bw() +
  labs(y="overwinter survival") +
  theme(legend.position = "none") +
  geom_text(data=tt.u1, aes(label=Tukey_labels)) + ggtitle("SU_2011")

############
### RAT
max.vals <- aggregate(nb$SR_2011,by=list(nb$K.group), max, na.rm=TRUE)
max.buffer <- data.frame(max.vals[,2] + (0.2*max(max.vals[,2])))
Tukey_test <- cbind(max.buffer,ph.comp$SR_2011, max.vals[,1])
colnames(Tukey_test) <- c("SR_2011","Tukey_labels","K.group")
tt.t1 <- Tukey_test

t1.plot <- ggplot(data=nb, aes(x=K.group, y=SR_2011,fill=K.group)) +
  geom_boxplot() +
  geom_jitter(width=0.1) +
  scale_fill_manual(values=kcol) +
  theme_bw() +
  labs(y="overwinter survival") +
  theme(legend.position = "none") +
  geom_text(data=tt.t1, aes(label=Tukey_labels)) + ggtitle("SR_2011")
```

```{R nonparam mean plot 2011 generate}
pdf("./figures/32.figures/ows.means.2011.nonparametric.by.K.pdf", width=9, height=9)
figure.nonparam <- ggarrange(a1.plot, r1.plot, u1.plot, t1.plot, 
          labels = c("A", "B", "C","D"),
          ncol = 2, nrow = 2)
annotate_figure(figure.nonparam, top = text_grob("mean overwinter survival - KW tests", face = "bold", size = 20))
dev.off()
```

```{R nonparam means plots 2012}
############
### ADA
max.vals <- aggregate(nb$NA_2012,by=list(nb$K.group), max, na.rm=TRUE)
max.buffer <- data.frame(max.vals[,2] + (0.2*max(max.vals[,2])))
Tukey_test <- cbind(max.buffer,ph.comp$NA_2012, max.vals[,1])
colnames(Tukey_test) <- c("NA_2012","Tukey_labels","K.group")
tt.a1 <- Tukey_test

a2.plot <- ggplot(data=nb, aes(x=K.group, y=NA_2012,fill=K.group)) +
  geom_boxplot() +
  geom_jitter(width=0.1) +
  scale_fill_manual(values=kcol) +
  theme_bw() +
  labs(y="overwinter survival") +
  theme(legend.position = "none") +
  geom_text(data=tt.a1, aes(label=Tukey_labels)) + ggtitle("NA_2012")

############
### RAM
max.vals <- aggregate(nb$NM_2012,by=list(nb$K.group), max, na.rm=TRUE)
max.buffer <- data.frame(max.vals[,2] + (0.2*max(max.vals[,2])))
Tukey_test <- cbind(max.buffer,ph.comp$NM_2012, max.vals[,1])
colnames(Tukey_test) <- c("NM_2012","Tukey_labels","K.group")
tt.r1 <- Tukey_test

r2.plot <- ggplot(data=nb, aes(x=K.group, y=NM_2012,fill=K.group)) +
  geom_boxplot() +
  geom_jitter(width=0.1) +
  scale_fill_manual(values=kcol) +
  theme_bw() +
  labs(y="overwinter survival") +
  theme(legend.position = "none") +
  geom_text(data=tt.r1, aes(label=Tukey_labels)) + ggtitle("NM_2012")

############
### ULL
max.vals <- aggregate(nb$SU_2012,by=list(nb$K.group), max, na.rm=TRUE)
max.buffer <- data.frame(max.vals[,2] + (0.2*max(max.vals[,2])))
Tukey_test <- cbind(max.buffer,ph.comp$SU_2012, max.vals[,1])
colnames(Tukey_test) <- c("SU_2012","Tukey_labels","K.group")
tt.u1 <- Tukey_test

u2.plot <- ggplot(data=nb, aes(x=K.group, y=SU_2012,fill=K.group)) +
  geom_boxplot() +
  geom_jitter(width=0.1) +
  scale_fill_manual(values=kcol) +
  theme_bw() +
  labs(y="overwinter survival") +
  theme(legend.position = "none") +
  geom_text(data=tt.u1, aes(label=Tukey_labels)) + ggtitle("SU_2012")

############
### RAT
max.vals <- aggregate(nb$SR_2012,by=list(nb$K.group), max, na.rm=TRUE)
max.buffer <- data.frame(max.vals[,2] + (0.2*max(max.vals[,2])))
Tukey_test <- cbind(max.buffer,ph.comp$SR_2012, max.vals[,1])
colnames(Tukey_test) <- c("SR_2012","Tukey_labels","K.group")
tt.t1 <- Tukey_test

t2.plot <- ggplot(data=nb, aes(x=K.group, y=SR_2012,fill=K.group)) +
  geom_boxplot() +
  geom_jitter(width=0.1) +
  scale_fill_manual(values=kcol) +
  theme_bw() +
  labs(y="overwinter survival") +
  theme(legend.position = "none") +
  geom_text(data=tt.t1, aes(label=Tukey_labels)) + ggtitle("SR_2012")
```

```{R nonparam mean plot 2012 generate}
pdf("./figures/32.figures/ows.means.2012.nonparametric.by.K.pdf", width=9, height=9)
figure.nonparam <- ggarrange(a2.plot, r2.plot, u2.plot, t2.plot, 
          labels = c("A", "B", "C","D"),
          ncol = 2, nrow = 2)
annotate_figure(figure.nonparam, top = text_grob("mean overwinter survival - KW tests", face = "bold", size = 20))
dev.off()
```

```{R nonparam mean plot both years generate}
pdf("./figures/32.figures/mean.ows.nonparametric.both.years.by.K.pdf", width=18, height=12)
figure.nonparam <- ggarrange(a1.plot, r1.plot, u1.plot, t1.plot,a2.plot, r2.plot, u2.plot, t2.plot, 
          labels = c("A", "B", "C","D","E","F","G","H"),
          ncol = 4, nrow = 2)
annotate_figure(figure.nonparam, top = text_grob("mean overwinter survival - KW tests", face = "bold", size = 20))
dev.off()
```

## 9.  Modeling 2012 data only.  This is what would be relevant for composite fitness (SnR experiments)

```{r mixed models 2012}
dm0 <- glm(ows~exp, data=d12, family=binomial("logit"))
summary(glht(dm0, mcp(exp="Tukey"))) ### ULL/ADA and ULL/RAM.  interesting.

dmid <- glm(ows~id, data=d12, family=binomial("logit"))
summary(dmid)

dm1 <- glmer(ows~exp + (1|id), data=d12,family=binomial("logit"),control=glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)))

dm2 <- glmer(ows~exp + (1|exp:id), data=d12,family=binomial("logit"),control=glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)))

dm3 <- glmer(ows ~ 1 + (1 | id), data=d12, family=binomial("logit"),control=glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)))

dm4 <- glmer(ows~exp + (1|exp:block), data=d12, family=binomial("logit"),control=glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5))) # so this is correctly specified (without year effect, at least)

dm5 <- glmer(ows~exp + (1|exp:block) + (1|id), data=d12,family=binomial("logit"),control=glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)))

## block effects are singular (dm4 and dm5)


# get fit information for all these models
mods <- c("dm1", "dm2", "dm3","dm4", "dm5")
mod.sum <- matrix(NA, ncol=5, nrow=length(mods))
for(up in 1:length(mods)){
  up.m <- mods[up]
  up.lm <- get(up.m)
  up.fit <- summary(up.lm)$AICtab
  mod.sum[up,] <- up.fit
}
rownames(mod.sum) <- mods
colnames(mod.sum) <- names(up.fit)
print(mod.sum)

### looks like dm2 (full model minus block) is best

int2 <- predictInterval(dm2)
int.dm2 <- REsim(dm2)
plotREsim(REsim(dm2))

anova(dm2)
summary(dm2)

# random effect variances
var.dm2 <- VarCorr(dm2)
print(var.dm2,comp=c("Variance","Std.Dev."),digits=4)
re.var.dm2 <- as.data.frame(var.dm2, row.names = NULL,optional = FALSE)

r.squaredGLMM(dm1)
r.squaredGLMM(dm2)
r.squaredGLMM(dm3)
r.squaredGLMM(dm4)
r.squaredGLMM(dm5)
## R2m is from marginal(fixed effects)
## R2c is from entire model

## dm2 only has R2 of 0.03!
## still from comparing between models, there is 
```

## 10 Explore 2012 only blups
```{r explore 2012 blups}
blups12 <- ranef(dm2)[[1]]
blups12 <- as.data.frame(blups12)
desc <- do.call(rbind,strsplit(rownames(blups12),":"))
colnames(desc) <- c("exp","id")
blups12 <- cbind(blups12, desc)
colnames(blups12)[1] <- "ows.b"
blups12$s.name <- blups12$exp
blups12$s.name <- gsub("ULL","SU",blups12$s.name)
blups12$s.name <- gsub("RAT","SR",blups12$s.name)
blups12$s.name <- gsub("RAM","NM",blups12$s.name)
blups12$s.name <- gsub("ADA","NA",blups12$s.name)
blups12$s.name <- factor(blups12$s.name, levels=c("NA","NM","SU","SR"))
blups12$region[blups12$exp%in%c("ULL", "RAT")]="S"
blups12$region[blups12$exp%in%c("RAM", "ADA")]="N"

hist(blups12$ows.b)

### BLUPS are the log odds probability of survival.  Make probability

logit2prob <- function(logit){
  odds <- exp(logit)
  prob <- odds / (1 + odds)
  return(prob)
}

blups12$ows.p <- logit2prob(blups12$ows.b)
with(blups12, boxplot(ows.p~exp, col=c(rep(brewer.pal(9, "Paired")[4],4), rep(brewer.pal(9, "Paired")[3],4)), ylab="BLUP (logit ows)"))

ows.blup2.plot <- ggplot(data=blups12, aes(x=s.name, y=ows.b, fill=region)) +
  geom_boxplot() +
  scale_fill_manual(values=c("dodgerblue1","springgreen2")) +
  ylab("overwinter survival 2012 BLUPs (logit)") +
  xlab("experiment") +
  labs(fill = "experiment \nregion")

pdf(file="./figures/32.figures/blups.ows2012.boxplot.by.year.site.pdf", width=7, height=5)
print(ows.blup2.plot)
dev.off()

ows.prob2.plot <- ggplot(data=blups12, aes(x=s.name, y=ows.p, fill=region)) +
  geom_boxplot() +
  scale_fill_manual(values=c("dodgerblue1","springgreen2")) +
  ylab("overwinter survival 2012 BLUPs (probability)") +
  xlab("experiment") +
  labs(fill = "experiment \nregion")

pdf(file="./figures/32.figures/prob.ows2012.boxplot.by.year.site.pdf", width=7, height=5)
print(ows.prob2.plot)
dev.off()

## look at correlations between actual means and blups
comp2 <- merge(blups12, ows.m[ows.m$year==2012,], by=c("exp", "id","s.name", "region"), all.x=TRUE)

## write BLUPS and means to file
write.csv(comp2, file="./data/32.data/marginal.ows.2012.blups.csv", quote=FALSE,row.names=FALSE)

pdf("./figures/32.figures/blup.vs.mean.ows.2012.pdf", width=6, height=5)
comp.plot <- ggplot(data=comp2, aes(x=ows, y=ows.p, color=s.name)) +
  geom_point() +
  xlab("mean overwinter survival") +
  ylab("probability overwinter survival (from BLUPs)")
print(comp.plot)
dev.off()

```

## 11. Relationship between ows and flowering before winter
```{R ows and flowering}
# load before winter flowering data
# this is an old dataset (from 2012 - from prelim.analysis.fall.sweden.flowering.R)
# if looks interesting, need to re-generate these values.
# but just interested in looking at relationship first
ft <- read.csv("./data/t2.flowering.GWAS.csv")
colnames(ft) <- c("id","fbw")
ft.dat <- merge(comp, ft, all.x=TRUE)
ft.dat <- ft.dat[ft.dat$year=="2011",]

cg.colors <- c("dodgerblue","dodgerblue4","springgreen3","lightgreen")

pdf(file="./figures/32.figures/flowering.ows.2011.pdf", width=8, height=5)
ft.plot <- ggplot(data=ft.dat, aes(x=ows, y=fbw, color=s.name)) +
  geom_point() +
  scale_color_manual(values=cg.colors) +
  xlab("overwinter survival 2011") +
  ylab("flowering before winter 2011") +
  labs(color = "CG site")
print(ft.plot)
dev.off()


```


