---
title: "34.slugs.K.groups"
author: "Daniele Filiault"
date: "3/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library("RColorBrewer")
library("ggplot2")
library(gridExtra)
library(multcompView)
library(corrplot)
library(ggpubr)

```

## Introduction
More investigation of slug damage phenotype
Built off 11.slug.analysis.Rmd, but using new fitness BLUPs and K-matrix groups

### 1.load data
```{r load data}
# slug damage proportions from 01.CG.DLF.intial (mean of fitted values, glm, binomial with log link fxn)
load("./data/slug.phenos.fitted.Rdat")

#K matrix groups
k.groups <- read.table(file="./data/Kmatrix.6cluster.membership.txt")
slug.phenos <- merge(slug.phenos,k.groups, by.x="id", by.y="row.names", all=TRUE)

#fitness blups
fitness <- read.csv("./data/marginal.blups.csv", header=TRUE, stringsAsFactors=FALSE)
fitness <- fitness[fitness$exp=="RAT" & fitness$year=="2011",]
slug.phenos <- merge(slug.phenos, fitness[,c(1,4)], all=TRUE)

#overwinter survival blups
ows <- read.csv("./data/marginal.ows.blups.csv", header=TRUE,stringsAsFactors=FALSE)
ows <- ows[ows$exp=="RAT" & ows$year=="2011",]
slug.phenos <- merge(slug.phenos, ows[,c(1,4)], all=TRUE)

# fecundity
fecund <- read.csv("./data/marginal.fecundity.blups.csv", header=TRUE,stringsAsFactors=FALSE)
fecund <- fecund[fecund$exp=="RAT" & fecund$year=="2011",]
colnames(fecund)[1] <- "fecund"
slug.phenos <- merge(slug.phenos, fecund[,c(1,4)],all=TRUE)

#add descriptive k.group names
ki <- data.frame(matrix(c("K1","S1","K2","S2","K3","C","K4","N1","K5","N2","K6","B"),ncol=2,byrow=TRUE))
colnames(ki) <- c("K.group","K.name")
slug.phenos <- merge(slug.phenos, ki, all.x=TRUE, by="K.group")
slug.phenos$K.name <- factor(slug.phenos$K.name, levels=c("B","S1","S2","C","N1","N2"))

```

### 2. relationships between phenos
```{r pheno relationships}

lm.slug <- lm(ows.b ~ X2011_RAT_severeslug, data=slug.phenos)
summary(lm.slug) #R2=0.431
with(slug.phenos, plot(X2011_RAT_severeslug, ows.b , xlab="proportion of severe slug damage", ylab="overwinter survival"))
abline(lm.slug, col="blue")

lm.fit <- lm(fb ~ X2011_RAT_severeslug, data=slug.phenos)
summary(lm.fit) #r2=0.1807
with(slug.phenos, plot(X2011_RAT_severeslug, fb , xlab="proportion of severe slug damage", ylab="fitness BLUPs RAT 2011"))
abline(lm.fit, col="blue")

lm.fec <- lm(fecund ~ X2011_RAT_severeslug, data=slug.phenos)
summary(lm.fec) #r2=0.06339
with(slug.phenos, plot(X2011_RAT_severeslug, fecund , xlab="proportion of severe slug damage", ylab="fecundity BLUPs RAT 2011"))
abline(lm.fec, col="blue")
```

Proportion severe slug damage can predict ows, fitness, fecundity (R2 in that order)


### 3.  Slug damage by K group
```{r slug by K}
kcol <- brewer.pal(6, "Paired")[c(6,1,2,5,3,4)]

slk <- ggplot(data=slug.phenos, aes(x=K.name, y=X2011_RAT_severeslug,fill=K.name)) +
  geom_boxplot() +
  geom_jitter(width=0.1) +
  theme_bw() +
  labs(y="proportion severe slug damage",x="K group") +
  scale_fill_manual(values = kcol) +
  theme(legend.position = "none")
  
print(slk)


```

```{r anova slug by group}

aov1 <- aov(X2011_RAT_severeslug~K.name, data=slug.phenos)
summary(aov1)
tt1 <- TukeyHSD(aov1)
plot(tt1)

Tukey.levels <- tt1[[1]][, 4]
Tukey.labels <- data.frame(multcompLetters(Tukey.levels)['Letters'])
Tukey.labels <- Tukey.labels[match(levels(slug.phenos$K.name),row.names(Tukey.labels)),]

max.vals <- aggregate(slug.phenos$X2011_RAT_severeslug,by=list(slug.phenos$K.name), max)
max.buffer <- data.frame(max.vals[,2] + (0.05*max(max.vals[,2])))
Tukey_test <- cbind(max.buffer,Tukey.labels, max.vals[,1])
colnames(Tukey_test)[1] <- "X2011_RAT_severeslug"
colnames(Tukey_test)[3] <- "K.name"

slk.anova <- slk +geom_text(data=Tukey_test, aes(label=Tukey.labels))

print(slk.anova)

pdf(file="./figures/severe.slug.by.K.pdf", width=5, height=3.5) 
print(slk.anova)
dev.off()

```


### 4.  Fitness component vs slug damage relationships

```{R slug damage vs fitness metrics by admix group}

##overwinter survival
slug.ows <- ggplot(slug.phenos, aes(x=X2011_RAT_severeslug, y=ows.b, colour=K.name)) + 
  geom_point() + 
  scale_color_manual(values = kcol, name="K group") +
  #scale_color_brewer(palette="Paired") +
  theme_bw() + 
  labs(x="proportion severe slug damage", y="overwinter survival BLUPs") + 
  geom_smooth(method='lm',se=FALSE)
print(slug.ows)

lm.slug.ows <- lm(ows.b~X2011_RAT_severeslug, data=slug.phenos)
summary(lm.slug.ows)


##fecundity
slug.fec <- ggplot(slug.phenos, aes(x=X2011_RAT_severeslug, y=fecund, colour=K.name)) + 
  geom_point() + 
  scale_color_manual(values = kcol, name="K group") +
  theme_bw() + 
  labs(x="proportion severe slug damage", y="fecundity BLUPs") + 
  geom_smooth(method='lm',se=FALSE)

print(slug.fec)

lm.slug.fec <- lm(fecund~X2011_RAT_severeslug, data=slug.phenos)
summary(lm.slug.fec)


##fitness
slug.fit <- ggplot(slug.phenos, aes(x=X2011_RAT_severeslug, y=fb, colour=K.name)) + 
  geom_point() + 
  scale_color_manual(values = kcol, name="K group") +
  theme_bw() + 
  labs(x="proportion severe slug damage", y="fitness BLUPs") + 
  geom_smooth(method='lm',se=FALSE)

print(slug.fit)

lm.slug.fit <- lm(fb~X2011_RAT_severeslug, data=slug.phenos)
summary(lm.slug.fit)


### output these 3 figures
pdf("./figures/slug.fitness.components.pdf", width=10, height=8)
ggarrange(slug.ows, slug.fec, slug.fit, 
          labels = c("A", "B", "C"),
          ncol = 2, nrow = 2)
dev.off()

## output fitness specifically
pdf("./figures/slug.fitness.vs.damage.pdf", width=5, height=3.5)
print(slug.fit)
dev.off()

```
### 5. Fitness component vs slug damage relationships - interaction models
```{R fit vs slug interactions}

### ows
lm.slug.ows.i <- lm(ows.b~X2011_RAT_severeslug*K.group, data=slug.phenos)
summary(lm.slug.ows.i) ## groups 2 and 3 interaction.
anova(lm.slug.ows.i)

### fecundity
lm.slug.fec.i <- lm(fecund~X2011_RAT_severeslug*K.group, data=slug.phenos)
summary(lm.slug.fec.i) ## no interactions, but 4 and 5 (the N Swedish groups) are sig different
anova(lm.slug.fec.i)

### fitness
lm.slug.fit.i <- lm(fb~X2011_RAT_severeslug*K.group, data=slug.phenos)
summary(lm.slug.fit.i) ## no interactions, but 4 and 5 (the N Swedish groups) are sig different
anova(lm.slug.fit.i)

```