---
title: "38.fitness.BLUPS.with.Kgroups.Rmd"
author: "DLF"
date: "5/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lme4) # mixed models
library(statmod) 
library(merTools)
library(MuMIn) # for calculating R2 of mixed models
library(RColorBrewer)
library(pander)
library(dplyr)
library(tidyr)
library(ggplot2)
library(ecodist)
library(gridExtra)
library(pander)
library(knitr)
library(afex)

setwd("/groups/nordborg/projects/field_experiments/adaptation_sweden/common.gardens/")
#setwd("/Volumes/field_experiments/adaptation_sweden/common.gardens/")
```

## Introduction

Expanding some of the linear model analyses from 27.fitness.BLUPS.Rmd, incorporating K.groups from
29.BLUP.heatmap.by.K.Rmd.


### 1. prep data

```{r data prep}

#fitness estimates
d11=readRDS(file="./data/d11_for_analysis.rds")
d12=readRDS(file="./data/d12_for_analysis.rds")

##restrict to fitness
traits=c("fitness")

d11$year=2011
d12$year=2012

d11=d11[, c("year", "exp", "block", "id", "fitness")]
d12=d12[, c("year", "exp", "block", "id", "fitness")]

d=rbind(d11, d12)
d$region[d$exp%in%c("ULL", "RAT")]="S"
d$region[d$exp%in%c("RAM", "ADA")]="N"
d$year=as.factor(d$year)
d=d[, c("year", "region", "exp","block", "id", "fitness")]
d=na.omit(d)

# add K.groups
kg <- read.table('./data/Kmatrix.6cluster.membership.txt',stringsAsFactors=FALSE)
d <- merge(d, kg, by.x="id", by.y="row.names", all.x=TRUE)
d$id <- as.factor(d$id)
d$K.group <- as.factor(d$K.group)
d$ye <- paste(d$exp, d$year, sep="_")

# K.matrix
K <- read.table("./data/K.matrix.200Swedes.txt")
## need accession names
pref="./GWA/snps/sweden_200_MAF10.bimbam"
acclist=scan(paste(pref, ".acclist", sep=""), sep=",")
rownames(K) <- acclist
colnames(K) <- acclist
K <- as.matrix(K)

# BLUPs from 27.fitness.BLUPS.Rmd
blups <- read.csv(file="./data/marginal.blups.csv", stringsAsFactors=FALSE)
blups.l <- blups %>% pivot_wider(id_cols=id,names_from=c(exp,year),values_from=fb)%>%as.data.frame()
blups.l <- blups.l[order(as.character(blups.l$id)),]
blups.l$id <- as.character(blups.l$id)
#blups.l$id==colnames(K) #OK! same order

# add K groups to BLUPS
blups.k <- merge(blups, kg, by.x="id", by.y="row.names", all.x=TRUE)
blups.lk <- merge(blups.l, kg, by.x="id", by.y="row.names", all.x=TRUE)

# write blups for Mayela
write.table(blups.lk, file="./data/fitness.BLUPS.with.Kgroup.txt",quote=FALSE)
```

### 2. Mantel test genetic relatedness versus fitness relatedness
How much does genetic relatedness explain fitness patterns
```{r genetic fitness mantel, eval=FALSE}
k.dist <- dist(K)

m.test <- apply(blups.l[,2:9], 2, function(x){
  up.dist <- dist(x)
  up.mantel <- mantel(up.dist ~ k.dist, mrank=TRUE)
  return(up.mantel)
})

m.test <- t(m.test)
m.test
pdf("./figures/mantel.test.fitness.genetic.relatedness.pdf", height=4, width=8.5)
grid.table(m.test)
dev.off()
```

### 3. Fitness models with K groups instead of individual IDs.
```{r fitness K group normal model, eval=FALSE}
load(file="./data/38.fitness.BLUPS.with.Kgroups.Rdata")

#lm3 <- lm(fitness~id*year*exp + exp:year/block, data=d) is used in 27.fitness.BLUPS.Rmd
# add K groups with id nested inside?

#lm1 <- lm(fitness~K.group/id, data=d)
anova(lm1) ## K.group and id significant

#lm2 <- lm(fitness~year*K.group/id, data=d)
anova(lm2) # all significant

#lm3 <- lm(fitness~year*exp*K.group/id, data=d)
anova(lm3)
pander(anova(lm3))

#lm4 <- lm(fitness~year*exp*K.group/id + exp:year/block, data=d)
#anova(lm4)

lm5 <- lm(fitness~year*exp*K.group + exp:year/block, data=d)
anova(lm5)
pander(anova(lm5))
#save.image(file="./data/38.fitness.BLUPS.with.Kgroups.Rdata")
```


### 4. Models by K.group
Instead of trying to specify this insane model all at once, how about doing one model per K group and then asking for GxE differences among lines of each K group?
```{r fitness modeling by K group}
lm.k1 <- lm(fitness~id*year*exp + exp:year/block, data=d[d$K.group=="K1",])
lm.k2 <- lm(fitness~id*year*exp + exp:year/block, data=d[d$K.group=="K2",])
lm.k3 <- lm(fitness~id*year*exp + exp:year/block, data=d[d$K.group=="K3",])
lm.k4 <- lm(fitness~id*year*exp + exp:year/block, data=d[d$K.group=="K4",])
lm.k5 <- lm(fitness~id*year*exp + exp:year/block, data=d[d$K.group=="K5",])
lm.k6 <- lm(fitness~id*year*exp + exp:year/block, data=d[d$K.group=="K6",])

pander(anova(lm.k1))
pander(anova(lm.k2))
pander(anova(lm.k3))
pander(anova(lm.k4))
pander(anova(lm.k5))
pander(anova(lm.k6))
```

### 5. expand random effects models with K groups and id nesting
```{r raneff with K.groups}
#old best model from 27.fitness.BLUPs.Rmd
mm7 <- lmer(fitness~exp*year + (1|exp:year:block) + (1|exp:year:id), data=d)
#VarCorr(mm7)
#anova(mm7)
#summary(mm7)
print(mm7@call)
r.squaredGLMM(mm7)

# add K.group, with id nested within
mm1 <- lmer(fitness~exp*year*K.group + (1|exp:year:K.group:id) + (1|exp:year:block), data=d)
#VarCorr(mm1)
#anova(mm1)
#summary(mm1)
pander(anova(mm1))
print(mm1@call)
r.squaredGLMM(mm1)
## R2m is from marginal(fixed effects)
## R2c is from entire model

# add K.group, no id effects
mm2 <- lmer(fitness~exp*year*K.group + (1|exp:year:block), data=d)
print(mm2@call)
r.squaredGLMM(mm2)

## mm1 without block effects
mm3 <- lmer(fitness~exp*year*K.group + (1|exp:year:K.group:id), data=d)
print(mm3@call)
r.squaredGLMM(mm3)

anova(mm7, mm1)
anova(mm1, mm2)
anova(mm3, mm1)
```

### 6. extract BLUPs from K.group models
```{r K.group model ID BLUPs}

blups.noK <- ranef(mm1)[[1]]
blups.noK <- as.data.frame(blups.noK)
desc <- do.call(rbind,strsplit(rownames(blups.noK),":"))
colnames(desc) <- c("exp","year","K","id")
blups.noK <- cbind(blups.noK, desc)
colnames(blups.noK)[1] <- "fb"

hist(blups.noK$fb)
pdf(file="./figures/blup.noK.effect.fitness.boxplot.by.year.site.pdf", width=10, height=6)
with(blups.noK, boxplot(fb~year+exp, col=c(rep(brewer.pal(9, "Paired")[4],4), rep(brewer.pal(9, "Paired")[3],4))))
dev.off()

## write BLUPS to file
write.csv(blups.noK, file="./data/marginal.blups.noK.effect.csv", quote=FALSE,row.names=FALSE)

```

### 7. Rank-correlations between experiments
```{r rank cor K.groupmodel blups}
blups.noK.l <- blups.noK %>% pivot_wider(id_cols=id,names_from=c(exp,year),values_from=fb)%>%as.data.frame()
pw <- combn(colnames(blups.noK.l[2:9]),2)

#blups from tests from models with K.groups
ct.noK <- apply(pw,2,function(x){
  up.dat <- blups.noK.l[,colnames(blups.noK.l)%in%x]
  up.test <- cor.test(~get(x[1]) + get(x[2]), data=up.dat, method="spearman")
  out.dat <- cbind(up.test$p.value, up.test$estimate)
  return(out.dat)
})
ct.noK <- matrix(ct.noK, ncol=2,byrow=TRUE)
colnames(ct.noK) <- c("p.Kfixedeff","rho.Kfixedeff")

#blups from tests from models without K.groups
ct <- apply(pw,2,function(x){
  up.dat <- blups.l[,colnames(blups.l)%in%x]
  up.test <- cor.test(~get(x[1]) + get(x[2]), data=up.dat, method="spearman")
  out.dat <- cbind(up.test$p.value, up.test$estimate)
  return(out.dat)
})
ct <- matrix(ct, ncol=2,byrow=TRUE)
colnames(ct) <- c("p.noKmodel","rho.noKmodel")

sp.cor <- as.data.frame(t(pw))
sp.cor <- cbind(sp.cor, ct.noK, ct)
sp.cor <- sp.cor[,c(1:2,5,3,6,4)]
sp.cor$rho.diff <- sp.cor$rho.noKmodel-sp.cor$rho.Kfixedeff
sp.cor <- sp.cor[order(sp.cor$rho.diff),]

wilcox.test(sp.cor$rho.Kfixedeff,sp.cor$rho.noKmodel,paired=TRUE)

id.model <- sp.cor$p.noKmodel<0.05
K.model <- sp.cor$p.Kfixedeff<0.05
sig.tests <- cbind(id.model, K.model)
sp.cor <- cbind(sp.cor, sig.tests)
```


