---
title: "50.compare.scan.nonscan.subsets.analyses.Rmd"
author: "Daniele Filiault"
date: "4/23/2021"
output: html_document
---

```{r setup, include=FALSE, purl=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '/Volumes/field_experiments/adaptation_sweden/common.gardens/')

library(ggpmisc)
library(ggplot2)
library(gridExtra)
library(dplyr)
library(tidyr)
library(ggpubr)
library(cowplot)
library(viridis)
library(rcompanion)
library(multcompView)
library(egg)
source("./50.compare.scan.nonscan.subsets.functions.R") ### functions for analysis
```


## Introduction

Script 48 looks at patterns of beta versus home allele, home allele frequency, and difference between home and away allele frequency for selection scan datasets.  

Here I would like to coplot these distributions and test whether there are significant differences in between scan subsets and the rest of the genome.
Running analyses here, functions developed in 50.compare.scan.nonscan.subsets.functions.Rmd

## 1. Load data

```{r load data, purl=FALSE}
### allele frequency data from script 46.allele.freq.differences.Kgroups
load("./data/pop.af.dat.Rdat")
load("./data/allele.freq.GWAS.snps.Rdata") #afg
```

```{r prep AFD data, purl=FALSE}
### combine AFD datasets
colnames(pop.af.dat)[1:2] <- c("chrom", "pos")
pos <- do.call(rbind,strsplit(rownames(afg),"_"))
colnames(pos) <- c("chrom", "pos")
pos <- as.data.frame(pos)
afg <- cbind(afg, pos)
af.dat <- merge(pop.af.dat, afg, by=c("chrom", "pos"), all=TRUE)
```

```{r set GWAS results files, purl=FALSE}
gwas.res.files.n <- c("./res/gemma_marginal/gemma_lmm_blup_RAM_2011.rds", "./res/gemma_marginal/gemma_lmm_blup_RAM_2012.rds", "./res/gemma_marginal/gemma_lmm_blup_ADA_2011.rds","./res/gemma_marginal/gemma_lmm_blup_ADA_2012.rds")
gwas.res.files.s <- c("./res/gemma_marginal/gemma_lmm_blup_RAT_2011.rds", "./res/gemma_marginal/gemma_lmm_blup_RAT_2012.rds", "./res/gemma_marginal/gemma_lmm_blup_ULL_2011.rds","./res/gemma_marginal/gemma_lmm_blup_ULL_2012.rds")
```

## 2. Swedish genomes paper sweep scan

```{r swedishgenomes}
# prep data
up.scan.file <- "./data/003.selection.scans/Swedish.genomes.paper.sweeps.simple.csv"
up.scan.name <- "swedishgenomes"
win.size <- 10000
sg.dat <- read.csv(up.scan.file, stringsAsFactors=FALSE)
sg.dat <- sg.dat[,1:2]
colnames(sg.dat) <- c("chr", "pos")

# run analysis
compare.by.scan.nonscan(af.dat=af.dat, ss.dat=sg.dat, up.scan.name="swedishgenomes", win.size=10000, gwas.res.files.n=gwas.res.files.n, gwas.res.files.s=gwas.res.files.s)
```

## 3. Fournier-Level - all experiments
```{r fournierlevelall data}
# prep data
up.scan.file <- "./data/003.selection.scans/Fournier_Level_GWAs_Clim_Data.csv"
up.scan.name <- "fournier.level"
win.size <- 10000

fl.dat <- read.csv(up.scan.file, stringsAsFactors=FALSE)
fl.dat <- fl.dat[,1:2]
colnames(fl.dat) <- c("chr", "pos")

compare.by.scan.nonscan(af.dat=af.dat, ss.dat=fl.dat, up.scan.name="fournierlevel", win.size=10000, gwas.res.files.n=gwas.res.files.n, gwas.res.files.s=gwas.res.files.s)
```


## 4a. Fournier-Level - Oulu (FIN) experiments only
```{r fournierlevelFIN data}
fl.dat <- read.csv(up.scan.file, stringsAsFactors=FALSE)
FIN.dat <- fl.dat[fl.dat$Location=="FIN",]
FIN.dat <- FIN.dat[,1:2]
colnames(FIN.dat) <- c("chr", "pos")
# subset for FIN experiment

compare.by.scan.nonscan(af.dat=af.dat, ss.dat=FIN.dat, up.scan.name="fournierlevelFIN", win.size=10000, gwas.res.files.n=gwas.res.files.n, gwas.res.files.s=gwas.res.files.s)
```

## 4b. Fournier-Level - Spain (SP) experiments only
```{r fournierlevelSP data}
fl.dat <- read.csv(up.scan.file, stringsAsFactors=FALSE)
SP.dat <- fl.dat[fl.dat$Location=="SP",]
SP.dat <- SP.dat[,1:2]
colnames(SP.dat) <- c("chr", "pos")
# subset for SP experiment

compare.by.scan.nonscan(af.dat=af.dat, ss.dat=SP.dat, up.scan.name="fournierlevelSP", win.size=10000, gwas.res.files.n=gwas.res.files.n, gwas.res.files.s=gwas.res.files.s)
```

## 4c. Fournier-Level - German (GER) experiments only
```{r fournierlevelGER data}
fl.dat <- read.csv(up.scan.file, stringsAsFactors=FALSE)
GER.dat <- fl.dat[fl.dat$Location=="GER",]
GER.dat <- GER.dat[,1:2]
colnames(GER.dat) <- c("chr", "pos")
# subset for GER experiment

compare.by.scan.nonscan(af.dat=af.dat, ss.dat=GER.dat, up.scan.name="fournierlevelGER", win.size=10000, gwas.res.files.n=gwas.res.files.n, gwas.res.files.s=gwas.res.files.s)
```

## 4d. Fournier-Level - UK (UK) experiments only
```{r fournierlevelSP data}
fl.dat <- read.csv(up.scan.file, stringsAsFactors=FALSE)
UK.dat <- fl.dat[fl.dat$Location=="UK",]
UK.dat <- UK.dat[,1:2]
colnames(UK.dat) <- c("chr", "pos")
# subset for UK experiment

compare.by.scan.nonscan(af.dat=af.dat, ss.dat=UK.dat, up.scan.name="fournierlevelUK", win.size=10000, gwas.res.files.n=gwas.res.files.n, gwas.res.files.s=gwas.res.files.s)
```


## 5. Hancock GEA 
```{r hancock data}
# prep data
up.scan.file <- "./data/003.selection.scans/Hancock.tophits.txt"
ha.dat <- read.table(up.scan.file, stringsAsFactors=FALSE, header=TRUE)
ha.dat <- ha.dat[,2:3]
colnames(ha.dat) <- c("chr", "pos")

compare.by.scan.nonscan(af.dat=af.dat, ss.dat=ha.dat, up.scan.name="hancock", win.size=10000, gwas.res.files.n=gwas.res.files.n, gwas.res.files.s=gwas.res.files.s)
```

## 6. Huber sweeps
```{r christian huber sweeps N/S}
# prep data
up.scan.file <- "./data/003.selection.scans/christian.intervals.csv"
ns.dat <- read.csv(up.scan.file, stringsAsFactors=FALSE)
ns.dat <- ns.dat[,1:2]
colnames(ns.dat) <- c("chr", "pos")
## input is 2 columns giving chr and pos

compare.by.scan.nonscan(af.dat=af.dat, ss.dat=ns.dat, up.scan.name="huber", win.size=10000, gwas.res.files.n=gwas.res.files.n, gwas.res.files.s=gwas.res.files.s)
```

## 7. Horton sweeps
```{r horton FST}
up.scan.file <- "./data/003.selection.scans/selection_scans_horton/global.all.fsts"
fst.dat <- read.table(up.scan.file, stringsAsFactors=FALSE, header=TRUE)
fst.dat <- fst.dat[order(fst.dat$fst, decreasing=TRUE),]
#take top 1%
fst.dat <- fst.dat[1:(nrow(fst.dat)*0.01),]
fst.dat <- fst.dat[,2:3]
colnames(fst.dat) <- c("chr", "pos")

compare.by.scan.nonscan(af.dat=af.dat, ss.dat=fst.dat, up.scan.name="hortonFST", win.size=10000, gwas.res.files.n=gwas.res.files.n, gwas.res.files.s=gwas.res.files.s)

```

```{r horton CLR}
up.scan.file <- "./data/003.selection.scans/selection_scans_horton/allCLR.txt"
clr.dat <- read.table(up.scan.file, stringsAsFactors=FALSE, header=TRUE)
clr.dat <- clr.dat[order(clr.dat$meanCLR, decreasing=TRUE),]
#take top 1%
clr.dat <- clr.dat[1:(nrow(clr.dat)*0.01),]
clr.dat <- clr.dat[,1:2]
colnames(clr.dat) <- c("chr", "pos")

compare.by.scan.nonscan(af.dat=af.dat, ss.dat=clr.dat, up.scan.name="hortonCLR", win.size=10000, gwas.res.files.n=gwas.res.files.n, gwas.res.files.s=gwas.res.files.s)

```


```{r horton PHS}
up.scan.file <- "./data/003.selection.scans/selection_scans_horton/phsscores.pos.txt"
phs.dat <- read.table(up.scan.file, stringsAsFactors=FALSE, header=TRUE)
phs.dat <- phs.dat[order(phs.dat$PHS, decreasing=TRUE),]
#take top 1%
phs.dat <- phs.dat[1:(nrow(phs.dat)*0.01),]
phs.dat <- phs.dat[,1:2]
colnames(phs.dat) <- c("chr", "pos")


compare.by.scan.nonscan(af.dat=af.dat, ss.dat=phs.dat, up.scan.name="hortonPHS", win.size=10000, gwas.res.files.n=gwas.res.files.n, gwas.res.files.s=gwas.res.files.s)
```

## 8. Exposito-Alonso data (GWA, AGWA)
```{r exposito GWA}
up.scan.file <- "./data/003.selection.scans/exposito_2018/S3_gwa.csv"
egwas.dat <- read.csv(up.scan.file, stringsAsFactors=FALSE, header=TRUE)
egwas.dat <- egwas.dat[,1:2]
colnames(egwas.dat) <- c("chr", "pos")


compare.by.scan.nonscan(af.dat=af.dat, ss.dat=egwas.dat, up.scan.name="exposGWAS", win.size=10000, gwas.res.files.n=gwas.res.files.n, gwas.res.files.s=gwas.res.files.s)
```

```{r exposito AGWA}
up.scan.file <- "./data/003.selection.scans/exposito_2018/S4_agwa.csv"
agwas.dat <- read.csv(up.scan.file, stringsAsFactors=FALSE, header=TRUE)
agwas.dat <- agwas.dat[,1:2]
colnames(agwas.dat) <- c("chr", "pos")
### this is only 5 loci

compare.by.scan.nonscan(af.dat=af.dat, ss.dat=agwas.dat, up.scan.name="exposAGWAS", win.size=10000, gwas.res.files.n=gwas.res.files.n, gwas.res.files.s=gwas.res.files.s)
```

## 9. Price data (LFMM, GEMMA, BAYESCAN, AFD.LD)
```{r price LFMM}
up.scan.file <- "./data/003.selection.scans/price_2020/LFMM_MinTmpCldM_K_8_q_values"
plfmm.dat <- read.table(up.scan.file, stringsAsFactors=FALSE, header=TRUE)
plfmm.dat <- plfmm.dat[plfmm.dat$q_value<0.1,]
plfmm.dat <- plfmm.dat[,1:2]
colnames(plfmm.dat) <- c("chr", "pos")


compare.by.scan.nonscan(af.dat=af.dat, ss.dat=plfmm.dat, up.scan.name="priceLFMM", win.size=10000, gwas.res.files.n=gwas.res.files.n, gwas.res.files.s=gwas.res.files.s)
```

```{r price GEMMA}
up.scan.file <- "./data/003.selection.scans/price_2020/GEMMA_MinTmpCldM_q_values"
pgemma.dat <- read.table(up.scan.file, stringsAsFactors=FALSE, header=TRUE)
pgemma.dat <- pgemma.dat[pgemma.dat$q_value<0.1,]
pgemma.dat <- pgemma.dat[,1:2]
colnames(pgemma.dat) <- c("chr", "pos")


compare.by.scan.nonscan(af.dat=af.dat, ss.dat=pgemma.dat, up.scan.name="priceGEMMA", win.size=10000, gwas.res.files.n=gwas.res.files.n, gwas.res.files.s=gwas.res.files.s)
```

```{r price BAYESCAN}
up.scan.file <- "./data/003.selection.scans/price_2020/BAYESCAN_FDR_FINAL"
pbayes.dat <- read.table(up.scan.file, stringsAsFactors=FALSE, header=TRUE)
pbayes.dat <- pbayes.dat[pbayes.dat$FDR<0.1,]
pbayes.dat <- pbayes.dat[,1:2]
colnames(pbayes.dat) <- c("chr", "pos")


compare.by.scan.nonscan(af.dat=af.dat, ss.dat=pbayes.dat, up.scan.name="priceBAYES", win.size=10000, gwas.res.files.n=gwas.res.files.n, gwas.res.files.s=gwas.res.files.s)
```

```{r price AFD.LD}
up.scan.file <- "./data/003.selection.scans/price_2020/AFD_LD_Italy_Sweden_pops"
pal.dat <- read.table(up.scan.file, stringsAsFactors=FALSE, header=TRUE)
pal.dat <- pal.dat[pal.dat$AFD>0.7 & pal.dat$LD>0.19,] #cutoff -> AFD>0.7 and LD>0.19 (95th percentiles)
pal.dat <- pal.dat[,1:2]
colnames(pal.dat) <- c("chr", "pos")


compare.by.scan.nonscan(af.dat=af.dat, ss.dat=pal.dat, up.scan.name="priceAFDLD", win.size=10000, gwas.res.files.n=gwas.res.files.n, gwas.res.files.s=gwas.res.files.s)
```
## 10. Agren cross data - different script, need to define windows manually rather than around bp

