---
title: "F2.figure2"
author: "Daniele Filiault"
date: "7/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(glue)
library(ggplot2)
library(viridis)
library(RColorBrewer)
library(grid)
library(ggpubr)
library(ggh4x)
library(dplyr)
library(tidyr)
library(pgirmess)
library(ape)
library(multcompView)
```

## Introduction

Generates figure 2 of manuscript.
Panel A - Fitness BLUPs by background group (original analysis in script 29)
Panel B - Slug damage by background group (original analysis in script 34)
Panel C - Relationship between slug damage and fitness BLUPs (original analysis in script 34)

## 01. load data

```{r load fitness BLUP data}
#zscore normalized fitness BLUPs
fit.blups <- read.table(file="./data/normalized.BLUPs.txt")

## background group ID data
kcl <- read.table(file="./data/K.groups.txt", stringsAsFactors=FALSE)

fit.blups <- merge(fit.blups, kcl, by="row.names")
fit.blups <- fit.blups[,colnames(fit.blups)%in%c("Row.names","K.group")==FALSE]
```

```{r load slug BLUP data}
# slug damage proportions from 01.CG.DLF.intial (mean of fitted values, glm, binomial with log link fxn)
load("./data/slug.phenos.fitted.Rdat")

#K matrix groups
slug.phenos <- merge(slug.phenos, kcl, by.x="id", by.y="row.names", all=TRUE)

#fitness blups
fitness <- read.csv("./data/marginal.blups.csv", header=TRUE, stringsAsFactors=FALSE)
fitness <- fitness[fitness$exp=="RAT" & fitness$year=="2011",]
slug.phenos <- merge(slug.phenos, fitness[,c(1,4)], all=TRUE)

#overwinter survival blups
ows <- read.csv("./data/marginal.ows.blups.csv", header=TRUE,stringsAsFactors=FALSE)
ows <- ows[ows$exp=="RAT" & ows$year=="2011",]
slug.phenos <- merge(slug.phenos, ows[,c(1,4)], all=TRUE)

# fecundity
fecund <- read.csv("./data/marginal.fecundity.blups.csv", header=TRUE,stringsAsFactors=FALSE)
fecund <- fecund[fecund$exp=="RAT" & fecund$year=="2011",]
colnames(fecund)[1] <- "fecund"
slug.phenos <- merge(slug.phenos, fecund[,c(1,4)],all=TRUE)

#add descriptive k.group names
slug.phenos$K.name <- factor(slug.phenos$K.name, levels=c("B","S1","S2","C","N1","N2"))
```


## 02. do KW tests of fitness BLUP differences between groups
```{r KW tests sig letters}

### get labels for multiple testing
get.tuk.lab <- function(up.kmc){
  tlab <- up.kmc[[3]][,3]
  names(tlab) <- rownames(up.kmc[[3]])
  tlab <- data.frame(multcompLetters(tlab)['Letters'])
  return(tlab)
}

tukey.normfit <- as.list(rep(NA,8)) 

for(up in 1:8){
  up.exp <- colnames(fit.blups[up])
  up.kt <- kruskal.test(get(up.exp)~K.name, data=fit.blups)
  up.ph <- kruskalmc(get(up.exp)~K.name, data=fit.blups)
  out.ph <- get.tuk.lab(up.ph)

  max.vals <- aggregate(fit.blups[,up],by=list(fit.blups$K.name), max, na.rm=TRUE)
  total.max <- max(fit.blups[,1:8], na.rm=TRUE)
  max.buffer <- data.frame(max.vals[,2] + (0.15*total.max))
  Tukey_test <- cbind(max.buffer,out.ph, max.vals[,1])
  colnames(Tukey_test) <- c("blup","Tukey_labels","K.name") # blup isn't really a good description of what this variable is, but i need to keep the name consistant with the blup data so it can be coplotted in ggplot2
  Tukey_test$exp <- up.exp
  tukey.normfit[[up]] <- Tukey_test
}  

tukey.normfit <- do.call(rbind, tukey.normfit)
tuk.loc <- do.call(rbind, strsplit(tukey.normfit$exp,"_"))
tukey.normfit$exp.site <- tuk.loc[,1]
tukey.normfit$year <- tuk.loc[,2]
tukey.normfit <- cbind(tukey.normfit, geo = ifelse(tukey.normfit$exp.site%in%c("NA","NM"), "North Experiments","South Experiments"))
#tukey.normfit$exp.site <- factor(tukey.normfit$exp.site, levels=c("SR", "SU", "NA", "NM"))
tukey.normfit$exp.site <- factor(tukey.normfit$exp.site, levels=c("NA", "NM", "SR", "SU"))
```

## 03. do KW tests of fitness BLUP differences between groups - panel A

```{r plot fitness BLUP differences - panel A}
fb.l <- fit.blups %>% pivot_longer(cols = 1:8,names_to = "exp", values_to = "blup") %>%as.data.frame()
fb.loc <- do.call(rbind, strsplit(fb.l$exp,"_"))
fb.l$exp.site <- fb.loc[,1]
fb.l$year <- fb.loc[,2]
fb.l <- cbind(fb.l, geo = ifelse(fb.l$exp.site%in%c("NA","NM"), "North Experiments","South Experiments"))
#fb.l$geo <- factor(fb.l$geo, levels=c("South Experiments", "North Experiments"))
fb.l$geo <- factor(fb.l$geo, levels=c("North Experiments", "South Experiments"))
#fb.l$exp.site <- factor(fb.l$exp.site, levels=c("SR", "SU", "NA", "NM"))
fb.l$exp.site <- factor(fb.l$exp.site, levels=c("NA", "NM", "SR", "SU"))

fb.l$K.name <- factor(fb.l$K.name, levels=c("B","S1","S2","C","N1","N2"))
kcol <- brewer.pal(6, "Paired")[c(6,1,2,5,3,4)]

# make variable for "home" annotation
#ann <- unique(tukey.normfit[,c(4,5,6,7)])
#ann$blup <- -3
#ann$pos <- 5.5
#ann$K.name <- "B"  ## dummy value so this works with ggplot

fit.byk.plot <- ggplot(data=fb.l, aes(x=K.name, y=blup, fill=K.name)) +
  geom_boxplot() +
  geom_jitter(width=0.1, size=0.5) +
  scale_fill_manual(values=kcol) +
  theme_bw() +
  labs(y="normalized fitness") +
  theme(legend.position = "none") +
  xlab("genetic group") +
  geom_text(data=tukey.normfit, aes(label=Tukey_labels), size=3) +
  #geom_text(data=ann, aes(label="home\ngroups", x=pos)) +
  #facet_grid(year~exp.site)
  facet_nested(year~geo+exp.site)
  
```


## 04. Plot slug damage by K group - panel B

```{r slug by k Kruskal Wallace}
up.exp <- "X2011_RAT_severeslug"
up.ph <- kruskalmc(get(up.exp)~K.name, data=slug.phenos)
out.ph <- get.tuk.lab(up.ph)
max.vals <- aggregate(slug.phenos[,up.exp],by=list(slug.phenos$K.name), max, na.rm=TRUE)
total.max <- max(slug.phenos[,up.exp], na.rm=TRUE)
max.buffer <- data.frame(max.vals[,2] + (0.1*total.max))
Tukey_test <- cbind(max.buffer,out.ph, max.vals[,1])
colnames(Tukey_test) <- c(up.exp,"Tukey_labels","K.name") # blup isn't really a good description of what this variable is, but i need to keep the name consistant with the blup data so it can be coplotted in ggpl
```

```{r plot slug damage by K}

slug.byk.plot <- ggplot(data=slug.phenos, aes(x=K.name, y=X2011_RAT_severeslug,fill=K.name)) +
  geom_boxplot() +
  geom_jitter(width=0.1) +
  theme_bw() +
  labs(y="proportion severe\nslug damage",x="genetic group") +
  scale_fill_manual(values = kcol) +
  theme(legend.position = "none") +
  geom_text(data=Tukey_test, aes(label=Tukey_labels), size=3)
  
```

## 05. Plot slug damage vs fitness - panel C
``` {r plot slug vs fitness}

slug.fit <- ggplot(slug.phenos, aes(x=X2011_RAT_severeslug, y=fb, colour=K.name)) + 
  geom_point() + 
  scale_color_manual(values = kcol, name="genetic\ngroup") +
  theme_bw() + 
  labs(x="proportion severe slug damage", y="fitness BLUPs") + 
  geom_smooth(method='lm',se=FALSE)

```

## 06. Combine plots
```{r combine plots}

lower.plots <- plot_grid(slug.byk.plot, NULL, slug.fit, labels = c('B','', 'C'), label_size = 12, ncol=3, rel_widths=c(0.44,0.02, 0.54))
fig2 <- plot_grid(fit.byk.plot, NULL, lower.plots, labels = c('A','', ''), label_size = 12, ncol = 1, rel_heights = c(0.64, 0.02,0.34))

ggsave(fig2, file="./manuscript.figures/Figure2.jpg", width=10, height=8)
```
