---
title: "F3.figure4"
author: "Daniele Filiault"
date: "7/29/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(RColorBrewer)
library(grid)
library(gridExtra)
library(ggpubr)
library(cowplot)
library(ggh4x)
library(viridis)
```


## Introduction

Generates figure 4 of manuscript. 
panel A is cumulative beta distribution by allele frequency difference
panel B is  allele frequency difference by beta bin - numbers
panel C is  allele frequency difference by beta bin - proportions

## 01.  Get data
Data and analysis is from script 51

```{r get data}
# for panel A
load(file="./data/51.data/n.afd.sum.Rdat")
load(file="./data/51.data/s.afd.sum.Rdat")

# for panels B and C
load(file="./data/51.data/nbeta.sum.afd")
load(file="./data/51.data/sbeta.sum.afd")

```

  
## 02. make cumulative beta plots
```{r cumul beta data prep}
afd.sum <- rbind(n.afd.sum, s.afd.sum)
rm(n.afd.sum)
rm(s.afd.sum)
af.loc <- do.call(rbind, strsplit(afd.sum$exp,"_"))
afd.sum$exp.site <- af.loc[,1]
afd.sum$year <- af.loc[,2]

## also need borders of fixed for vertical lines
afd.fixed <- aggregate(index~exp, data=afd.sum[afd.sum$ha.afd>=0.85,], FUN=min)
af.loc <- do.call(rbind, strsplit(afd.fixed$exp,"_"))
afd.fixed$exp.site <- af.loc[,1]
afd.fixed$year <- af.loc[,2]

```

```{r cumul beta plot}

### plot weighed cumulative beta
panel.a <- ggplot(data=afd.sum, aes(x=index,y=csum.wb.norm)) +
  geom_point(aes(color=ha.afd)) +
  scale_color_viridis(direction=-1,name="allele\nfrequency\ndifference") +
  facet_nested(exp.site + year~.)  +
  #facet_wrap(~exp, ncol=2) +
  geom_hline(yintercept=0,linetype="dashed", color = "red") +
  xlab("number of SNPs") +
  ylab("cumulative weighted beta (normalized)") + 
  geom_vline(data=afd.fixed, aes(xintercept=index), linetype="dashed", color="purple") +
  theme(plot.margin=unit(c(1,0.5,0.5,0.5), "lines"))

```

## 03. make barplots (panels B and C)
```{r AF comp by beta North}

afd.beta.abs.n <- as.list(1:4)
afd.beta.prop.n <- as.list(1:4)
up.exp <- levels(nbeta.sum.afd$exp)

for(up in 1:4){
  up.dat <- nbeta.sum.afd[nbeta.sum.afd$exp==up.exp[up],]
  up.dat$beta.bin <- cut(up.dat$home.beta,10)
  up.dat$nb.bin <- as.numeric(up.dat$beta.bin)
  #get counts
  up.tab <- with(up.dat, table(ha.bins, nb.bin))
  up.df <- as.data.frame(up.tab)
  up.df$exp <- up.exp[up]
  afd.beta.abs.n[[up]] <- up.df
  #get as proportions
  up.ptab <- prop.table(up.tab, margin=2)
  up.pdf <- as.data.frame(up.ptab)
  up.pdf$exp <- up.exp[up]
  afd.beta.prop.n[[up]] <- up.pdf
}
afd.beta.abs.n <- do.call(rbind, afd.beta.abs.n)
afd.beta.prop.n <- do.call(rbind, afd.beta.prop.n)
```

```{r afd comp by beta South}

afd.beta.abs.s <- as.list(1:4)
afd.beta.prop.s <- as.list(1:4)
up.exp <- levels(sbeta.sum.afd$exp)

for(up in 1:4){
  up.dat <- sbeta.sum.afd[sbeta.sum.afd$exp==up.exp[up],]
  up.dat$beta.bin <- cut(up.dat$home.beta,10)
  up.dat$nb.bin <- as.numeric(up.dat$beta.bin)
  #get counts
  up.tab <- with(up.dat, table(ha.bins, nb.bin))
  up.df <- as.data.frame(up.tab)
  up.df$exp <- up.exp[up]
  afd.beta.abs.s[[up]] <- up.df
  #get as proportions
  up.ptab <- prop.table(up.tab, margin=2)
  up.pdf <- as.data.frame(up.ptab)
  up.pdf$exp <- up.exp[up]
  afd.beta.prop.s[[up]] <- up.pdf
}
afd.beta.abs.s <- do.call(rbind, afd.beta.abs.s)
afd.beta.prop.s <- do.call(rbind, afd.beta.prop.s)
```

```{r combine n and s data}
afd.beta.abs <- rbind(afd.beta.abs.n, afd.beta.abs.s)
afd.beta.prop <- rbind(afd.beta.prop.n, afd.beta.prop.s)
afd.loc <- do.call(rbind, strsplit(afd.beta.abs$exp,"_"))
afd.beta.abs$exp.site <- afd.loc[,1]
afd.beta.abs$year <- afd.loc[,2]
afd.beta.prop$exp.site <- afd.loc[,1]
afd.beta.prop$year <- afd.loc[,2]

afd.beta.abs$ha.bins <- factor(afd.beta.abs$ha.bins, levels=rev(levels(afd.beta.abs$ha.bins)))
afd.beta.prop$ha.bins <- factor(afd.beta.prop$ha.bins, levels=rev(levels(afd.beta.prop$ha.bins)))
```

```{r make absolute numbers bin plot}

panel.b <- ggplot(data=afd.beta.abs, aes(x=nb.bin,y=Freq,fill=ha.bins)) +
  geom_bar(position="stack",stat="identity") + 
  coord_flip()+
  ylab("number of SNPs") +
  xlab("home beta bin") +
  scale_fill_brewer(palette = "Paired", name="allele\nfrequency\ndifference\nbin") +
  #facet_grid(exp.site + year~.)
  #facet_nested(exp.site + year~.)+
  facet_grid(exp ~.) +
  theme(strip.background=element_blank(), strip.text=element_blank(), plot.margin=unit(c(1,0.5,0.5,1), "lines"), legend.position = "none")
```

```{r make proportions bin plot}

panel.c <- ggplot(data=afd.beta.prop, aes(x=nb.bin,y=Freq,fill=ha.bins)) +
  geom_bar(position="stack",stat="identity") + 
  coord_flip()+
  ylab("proportion of SNPs") +
  xlab("home beta bin") +
  scale_fill_brewer(palette = "Paired", name="allele\nfrequency\ndifference\nbin", guide=guide_legend(reverse=T)) +
  facet_nested(exp.site + year~.)  +
  theme(axis.text.y=element_blank(), plot.margin=unit(c(1,1,0.5,0), "lines"))
 
```


## 04. combine panels to final figure

```{r generate final figure}

fig4 <- ggarrange(panel.a, panel.b, panel.c + rremove("ylab"),
                    ncol = 3,
                    labels=c("A","B","C"),
                    font.label = list(size = 10, color = "black", face = "bold", family = NULL, position = "top"),
                    widths=c(2,1,1.5))

ggsave(fig4, file="./manuscript.figures/Figure4.jpg", width=10, height=9)

```

