---
title: "F5.figure5"
author: "Daniele Filiault"
date: "8/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(RColorBrewer)
library(grid)
library(gridExtra)
library(ggpubr)
library(cowplot)
library(ggh4x)
library(viridis)
```


## Introduction

Generates figure 5 of manuscript. 
panel A -> a cartoon of the four beta vs beta quadrants
panel B -> beta vs beta for SU2011 vs NM2011
panel C -> beta vs beta for SU2011 vs SU2012
panel D -> beta by beta by AF for SU2011 vs NM2011
panel E -> beta by beta by AF for SU2011 vs SU2012


## 01.  Get data
Data and analysis is from script 47.
Home betas of all SNPS and experiments with AF bins.

```{r load home beta and afd data}
load(file="./data/47.data/afb.n.Rdat")
load(file="./data/47.data/afb.s.Rdat")

```

```{r functions to extract data}
#up.exp <- "SU_2011"
#comp.exp <- "NM_2011"
#pol.dat <- afb.s

g.quad.counts <- function(up.exp, comp.exp, pol.dat){
  ue.col <- which(colnames(pol.dat)==up.exp)
  ce.col <- which(colnames(pol.dat)==comp.exp)
  g.table <- as.numeric(table(pol.dat[,ue.col]>0, pol.dat[,ce.col]>0))
  g.table <- cbind(g.table, c("NN","PN","NP","PP") )  ## N=negative, P=positive, first is home, second is away betas
  g.table <- as.data.frame(g.table, stringsAsFactors=FALSE)
  colnames(g.table) <- c("snp.no","quadrant")
  g.table$snp.no <- as.numeric(g.table$snp.no)
  g.table$snp.p <- g.table$snp.no/sum(g.table$snp.no)
  return(g.table)
  }

bin.quad.counts <- function(up.exp, comp.exp, pol.dat){
  ue.col <- which(colnames(pol.dat)==up.exp)
  ce.col <- which(colnames(pol.dat)==comp.exp)
  combos <- expand.grid(levels(pol.dat$ahome.bins), levels(pol.dat$aaway.bins))
  c.table <- apply(combos,1, function(x){
      up.d <- pol.dat[pol.dat$ahome.bins==x[1] & pol.dat$aaway.bins==x[2],]
      up.t <- table(up.d[,ue.col]>0, up.d[,ce.col]>0)
      up.u <- as.numeric(up.t)
      return(up.u)
      })
  c.table <- t(c.table)
  c.table <- as.data.frame(c.table)
  colnames(c.table) <- c("NN","PN","NP","PP") ## N=negative, P=positive, first is home, second is away betas
  c.table <- cbind(c.table,combos)
  colnames(c.table)[5:6] <- c("ahome.bins", "aaway.bins")

  c.table$totals <- apply(c.table[,1:4],1, sum)
  c.table.l <- gather(c.table, quadrant, snp.no, NN:PP, factor_key=TRUE)
  c.table.l$snp.p <- c.table.l$snp.no/c.table.l$totals
  return(c.table.l)
}

```


## 02. panel A - cartoon beta vs beta plot
```{r panel A}
mock.d <- data.frame(cbind(rep(1,4), c("NN","PN","NP","PP"), rep(0.25,4)), stringsAsFactors = FALSE)
colnames(mock.d) <- c("snp.no", "quadrant", "snp.p")
mock.d$snp.p <- as.numeric(mock.d$snp.p)

gd <- mock.d %>% 
    mutate(xmin = if_else(quadrant %in% c("NN","NP"), -sqrt(snp.p), 0),
         xmax = if_else(quadrant %in% c("PP", "PN"), sqrt(snp.p), 0),
         ymin = if_else(quadrant %in% c("NN","PN"), -sqrt(snp.p), 0),
         ymax = if_else(quadrant %in% c("PP","NP"), sqrt(snp.p), 0))

gd$quadrant <- factor(gd$quadrant, levels=c("NP","NN","PN","PP"))

panel.a <- ggplot(gd) + 
    geom_rect(aes(xmin=xmin, xmax=xmax, ymin=ymin, ymax=ymax, fill=quadrant)) +
    geom_text(aes(x = xmin + 0.5*sqrt(snp.p), y = ymin +0.5*sqrt(snp.p), label = c("common\nnegative","adaptive\ntradeoff",  "maladaptive\ntradeoff","common\npositive" )), size=4, fontface ="bold", color=c("white","black", "black", "white")) +
    #facet_grid(aaway.bins~ahome.bins,switch="both") +
    coord_equal() +
    scale_fill_brewer(palette="Paired") +
    geom_hline(yintercept = 0) +
    geom_vline(xintercept = 0) +
    labs(x="effect of home allele\nin home experiment", y="effect of home allele\nin away experiment") +
    #geom_segment(aes(x = 0.17, y = 0.17, xend = -0.17, yend = -0.17), color="white", arrow = arrow(length = unit(0.5, "cm"), ends="both"), size=2, lineend="round", linejoin="round") +
    #geom_segment(aes(x = -0.17, y = 0.17, xend = 0.17, yend = -0.17), color="black", arrow = arrow(length = unit(0.5, "cm"), ends="both"), size=2, lineend="round", linejoin="round") +
    theme_minimal() +
    theme(axis.text = element_blank(),
        panel.grid = element_blank(),
        legend.position="none")

```



## 03. format data for other panels


```{r generate SU_2011 vs NM_2011 datasets}
sn.g.dat <- g.quad.counts(up.exp="SU_2011", "NM_2011", pol.dat=afb.s)
sn.bin.dat <- bin.quad.counts(up.exp="SU_2011", "NM_2011", pol.dat=afb.s)

```

```{r generate SU_2011 vs SU_2012 datasets}
ss.g.dat <- g.quad.counts(up.exp="SU_2011", "SU_2012", pol.dat=afb.s)
ss.bin.dat <- bin.quad.counts(up.exp="SU_2011", "SU_2012", pol.dat=afb.s)

```

## 04. panel B
```{r panel B}
gd <- sn.g.dat %>% 
    mutate(xmin = if_else(quadrant %in% c("NN","NP"), -sqrt(snp.p), 0),
         xmax = if_else(quadrant %in% c("PP", "PN"), sqrt(snp.p), 0),
         ymin = if_else(quadrant %in% c("NN","PN"), -sqrt(snp.p), 0),
         ymax = if_else(quadrant %in% c("PP","NP"), sqrt(snp.p), 0))

gd$quadrant <- factor(gd$quadrant, levels=c("NP","NN","PN","PP"))

panel.b <- ggplot(gd) + 
    geom_rect(aes(xmin=xmin, xmax=xmax, ymin=ymin, ymax=ymax, fill=quadrant)) +
    geom_text(aes(x = xmin + 0.5*sqrt(snp.p), y = ymin +0.5*sqrt(snp.p), label = formatC(round(snp.p*100,2), digits=0,format="f")), size=5, fontface ="bold", color=c("white","black", "black", "white")) +
    #facet_grid(aaway.bins~ahome.bins,switch="both") +
    coord_equal() +
    scale_fill_brewer(palette="Paired") +
    geom_hline(yintercept = 0) +
    geom_vline(xintercept = 0) +
    labs(x="effect of Southern allele\nin SU 2011", y="effect of Southern allele\nin NM 2011") +
    theme_minimal() +
    theme(axis.text = element_blank(),
        panel.grid = element_blank(),
        legend.position = "none")
```

## 06. panel c - 

```{r panel c plot}

gd <- ss.g.dat %>% 
    mutate(xmin = if_else(quadrant %in% c("NN","NP"), -sqrt(snp.p), 0),
         xmax = if_else(quadrant %in% c("PP", "PN"), sqrt(snp.p), 0),
         ymin = if_else(quadrant %in% c("NN","PN"), -sqrt(snp.p), 0),
         ymax = if_else(quadrant %in% c("PP","NP"), sqrt(snp.p), 0))

gd$quadrant <- factor(gd$quadrant, levels=c("NP","NN","PN","PP"))

panel.c <- ggplot(gd) + 
    geom_rect(aes(xmin=xmin, xmax=xmax, ymin=ymin, ymax=ymax, fill=quadrant)) +
    geom_text(aes(x = xmin + 0.5*sqrt(snp.p), y = ymin +0.5*sqrt(snp.p), label = formatC(round(snp.p*100,2), digits=0,format="f")), size=5, fontface ="bold", color=c("white","black", "black", "white")) +
    #facet_grid(aaway.bins~ahome.bins,switch="both") +
    coord_equal() +
    scale_fill_brewer(palette="Paired") +
    geom_hline(yintercept = 0) +
    geom_vline(xintercept = 0) +
    labs(x="effect of Southern allele\nin SU 2011", y="effect of Southern allele\nin SU 2012") +
    theme_minimal() +
    theme(axis.text = element_blank(),
        panel.grid = element_blank(),
        legend.position = "none")
```

## 07. panel d 
```{r panel d}
pd <- sn.bin.dat %>% 
    mutate(xmin = if_else(quadrant %in% c("NN","NP"), -sqrt(snp.p), 0),
         xmax = if_else(quadrant %in% c("PP", "PN"), sqrt(snp.p), 0),
         ymin = if_else(quadrant %in% c("NN","PN"), -sqrt(snp.p), 0),
         ymax = if_else(quadrant %in% c("PP","NP"), sqrt(snp.p), 0))

pd$quadrant <- factor(pd$quadrant, levels=c("NP","NN","PN","PP"))
pd <-  cbind(pd, text.col = ifelse(pd$quadrant%in%c("NN","PP"), "white", "black"))
pd$ahome.bins <- factor(pd$ahome.bins, levels=rev(levels(pd$ahome.bins)))

text.color=c("black", "white")

panel.d <- ggplot(pd) + 
    geom_rect(aes(xmin=xmin, xmax=xmax, ymin=ymin, ymax=ymax, fill=quadrant)) +
    geom_text(aes(x = xmin + 0.5*sqrt(snp.p), y = ymin +0.5*sqrt(snp.p), color=text.col, label = formatC(round(snp.p*100,2), digits=0,format="f")), size=2.5, fontface ="bold") +
    facet_grid(ahome.bins ~ aaway.bins) +
    coord_equal() +
    scale_fill_brewer(palette="Paired") +
    scale_color_manual(values=text.color) +
    geom_hline(yintercept = 0) +
    geom_vline(xintercept = 0) +
    labs(x="effect of Southern allele in SU 2011", y="effect of Southern allele in NM 2011") +
    theme_minimal() +
    theme(axis.text = element_blank(),
        panel.grid = element_blank(),
        legend.position = "none",
        strip.background = element_rect(fill="grey80"),
        strip.placement = "outside")


```

## 08. panel e 
```{r panel e}
pd <- ss.bin.dat %>% 
    mutate(xmin = if_else(quadrant %in% c("NN","NP"), -sqrt(snp.p), 0),
         xmax = if_else(quadrant %in% c("PP", "PN"), sqrt(snp.p), 0),
         ymin = if_else(quadrant %in% c("NN","PN"), -sqrt(snp.p), 0),
         ymax = if_else(quadrant %in% c("PP","NP"), sqrt(snp.p), 0))

pd$quadrant <- factor(pd$quadrant, levels=c("NP","NN","PN","PP"))
pd <-  cbind(pd, text.col = ifelse(pd$quadrant%in%c("NN","PP"), "white", "black"))
pd$ahome.bins <- factor(pd$ahome.bins, levels=rev(levels(pd$ahome.bins)))

text.color=c("black", "white")

panel.e <- ggplot(pd) + 
    geom_rect(aes(xmin=xmin, xmax=xmax, ymin=ymin, ymax=ymax, fill=quadrant)) +
    geom_text(aes(x = xmin + 0.5*sqrt(snp.p), y = ymin +0.5*sqrt(snp.p), color=text.col, label = formatC(round(snp.p*100,2), digits=0,format="f")), size=2.5, fontface ="bold") +
    facet_grid(ahome.bins ~ aaway.bins) +
    coord_equal() +
    scale_fill_brewer(palette="Paired") +
    scale_color_manual(values=text.color) +
    geom_hline(yintercept = 0) +
    geom_vline(xintercept = 0) +
    labs(x="effect of Southern allele in SU 2011", y="effect of Southern allele in SU 2012") +
    theme_minimal() +
    theme(axis.text = element_blank(),
        panel.grid = element_blank(),
        legend.position = "none",
        strip.background = element_rect(fill="grey80"),
        strip.placement = "outside")


```

## 09. assemble and annotate panels d and e
```{r lower panels}

#lower.panel <- annotate_figure(panel.d, right = text_grob("Southern (home) allele frequency", rot=-90), top = text_grob("frequency of Southern (home) allele in Northern population"))

panel.d <- annotate_figure(panel.d, right = text_grob("Southern (home) allele frequency", rot=-90), top = text_grob("frequency of Southern (home) allele in Northern population"))

panel.e <- annotate_figure(panel.e, right = text_grob("Southern (home) allele frequency", rot=-90), top = text_grob("frequency of Southern (home) allele in Northern population"))

```


## 10. assemble figure
```{r assemble figure}

fig5 <- ggdraw() +
  draw_plot(panel.a, x = 0, y = .65, width = .33, height = .33) +
  draw_plot(panel.b, x = .33, y = .65, width = .33, height = .33) +
  draw_plot(panel.c, x = .66, y = .65, width = .33, height = .33) +
  draw_plot(lower.panel , x = 0, y = 0, width = 1, height = 0.6) +
  draw_plot_label(label = c("A", "B", "C", "D"), size = 15,
                  x = c(0, 0.33, 0.66, 0), y = c(1, 1, 1, 0.65))

ggsave(fig5, file="./manuscript.figures/Figure5.jpg", width=9.5, height=8.8)

```

```{r assemble figure extended}

fig5e <- ggdraw() +
  draw_plot(panel.a, x = 0, y = .78, width = .33, height = .22) +
  draw_plot(panel.b, x = .33, y = .78, width = .33, height = .22) +
  draw_plot(panel.c, x = .66, y = .78, width = .33, height = .22) +
  draw_plot(panel.d , x = 0, y = 0.39, width = 1, height = 0.38) +
  draw_plot(panel.e , x = 0, y = 0, width = 1, height = 0.38) +
  draw_plot_label(label = c("A", "B", "C", "D", "E"), size = 15,
                  x = c(0, 0.33, 0.66, 0, 0), y = c(1, 1, 1, 0.77, 0.38))

ggsave(fig5e, file="./manuscript.figures/Figuree5.jpg", width=9.0, height=13)

```



